### 2021-02-26 README.Rmd

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Something about MPAs and Bycatch

Repo for manuscript exploring the ways in which habitat overlap, fleet spillover, and MPA design strategies can interact to affect conservation and fishery outcomes of MPAs, focused on WCPO-ish system.