### 2021-06-10 documents/03_marlin_paper.Rmd
first draft of paper. 01_ is completely broken to some accidental deleting and undoing, so waiting until home to revert to backup version from 2021-06-09 at 6:28AM

---
title: "Solutions to the potential for unintended conservation consequences of MPAs"
author:
    - Daniel Ovando
    - Darcy Bradley
    - Lennon Thomas
    - Echelle Burns
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_document2: default
  bookdown::html_document2: default
bibliography: references.bib
params:
  run_name: ["v1.0"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, include = TRUE)

source(here::here("scripts", "00_setup.R"))

theme_set(marlin::theme_marlin())

```

```{r load-results}

results_files <- list.files(file.path(results_path))

rdses <- results_files[str_detect(results_files,'.rds')]

assign(str_remove(rdses[8],".rds"), read_rds(file.path(results_path,rdses[2])))

walk(rdses,~assign(str_remove(.x,".rds"), read_rds(file.path(results_path,.x)), envir = .GlobalEnv)) #load and assign all .rds objects in the appropriate results folder

rdatas <- results_files[str_detect(results_files,'.Rdata')]

walk(rdatas, ~load(file.path(results_path,.x), envir = .GlobalEnv))

# a bit of housekeeping to avoid saving a massive object
tmp <- raw_experiment_results  %>%
  map( ~ .x$obj)

group_var <- "placement_strategy"

total_results <- total_experiment_results

rm(total_experiment_results)

results <- placement_experiment %>%
  mutate(temp = raw_experiment_results) %>%
  group_by_at(colnames(.)[!colnames(.) %in% c("temp", "prop_mpa")]) %>%
  nest() %>%
  ungroup() %>%
  mutate(placement_id = 1:nrow(.)) %>%
  unnest(cols = data) %>%
  mutate(xid = map(raw_experiment_results, names)) %>%
  unnest(cols = c(temp, xid)) %>%
  mutate(obj = map(temp, "obj"),
         xid = as.numeric(xid)) %>%
  unnest(cols = obj)

if (any(results$econ[results$prop_mpa == 1] > 0)) {
  stop("something has gone wrong: yields present with 100% mpa")
}

warning("move placement experiement id back above run_experiments")

# total_experiment <- placement_experiment %>%
#   left_join(experiments, by = 'xid')

# results <- results %>%
#   left_join(placement_experiment)

# compare to BAU

results <- results %>%
  group_by(placement_id, xid, critter) %>%
  mutate(bau_biodiv = biodiv[prop_mpa == 0],
         bau_econ = econ[prop_mpa == 0])

excols <-
  experiments %>%
  select(xid, data) %>%
  unnest(cols = data) %>%
  select(-habitat,-critter, -scientific_name, -centroid) %>%
  ungroup() %>%
  unique()

```

# Introduction

Marine protected areas (MPAs) are a growing part of the marine
conservation toolbox, with 7.4% of the world's oceans currently
protected and, and bodies such as the IUCN calling for 30% protection by
2030. Many of the largest of these MPAs are being implemented in the
form of "blue-water" MPAs largely covering pelagic ecosystems,
characterized by highly mobile species and fishing fleets using a
diverse mix of static and dynamic ocean features. These systems operate
very differently from the coastal areas where modern MPA design
practices have been developed. Pelagic systems are also characterized by
combinations of relatively abundant species targeted by fishing, such as
most tuna stocks, and highly threatened species caught as incidental
"bycatch" in targeted fishing operations, such as oceanic whitetip
sharks. Blue-water MPAs are increasingly being proposed specifically to
benefit these by-catch populations while minimizing losses or where
possible providing benefits to fisheries.

While the effect of MPAs on fisheries yields is a continuously debated
topic, the conservation impacts of MPAs have received less attention of
late, due in part perhaps to the perceived obvious nature of the
problem: if you prevent fishing of a species within an area, there will
be more of that species in that area, and so the conservation benefits
of MPAs will increase with MPA size. \@hastings2017 clearly illustrates
this idea. Using a simple two-species two-patch analytical model, they
derive conditions under which MPAs can solve the "weak-stock" problem in
which a less productive species is incidentally exploited by a fleet
targeting a more valuable and productive stock.

However, as @hastings2017 point out, their results depend on a very
specific set of assumptions, namely that 1) adults are perfectly
sedentary 2) the target and bycatch species overlap perfectly in terms
of habitat xx....The authors suggest that the impact of these
assumptions should be tested in scenarios under which they are unlikely
to apply. We build on \@hastings2017 to consider the role of MPAs in reducing bycatch reductino while managing tradeoffs with fisheries under a more realistic model reflecting the dynamics of pelagic ecosystems and fisheries.

The core models underpinning MPA science have largely been based on
stylized versions of near-shore reef-style xx what to call that, like
coral / rocky / kelp forest, etc xx systems, where multiple species
often share common habitat, and adults have reasonable site fidelity
while larvae are broadly dispersed. Pelagic ecosystems have vastly
different dynamics than coastal seas [@game2009; @lubchenco2003],
characterized by highly migratory species specialized to take advantage
of highly dynamic ocean features such as eddies currents while also
utilizing static features such as seamounts. Pelagic ecosystems also
support both coastal domestic fishing industries along with highly
mobile international high-seas fishing operations. Given these dynamics,
it remains unclear how many of the lessons from coastal MPA science are
applicable to pelagic ecosystems [\@ovando2021XX].

Pelagic ecosystems present both opportunities and challenges for MPAs.
The highly productive and mobile nature of targeted fisheries in these
systems may mean that highly targeted MPAs can be designed that provide
effective conservation at minimal fishery costs. Strategic protection of
key aggregation locations may enhance depleted pelagic fisheries [xx
that Indian ocean study I can't remember right now]. xx big question
here: how focused do we want this paper to be? is it just about
conservation tradeoffs or lack thereof with bycatch species, or is about
the net conservation and fishery effects of BWMPAs. I'm inclined the
former to tee-up the later xx.

However, the dynamic nature of pelagic systems also creates a risk of unintended conservation consequences of MPAs. When an MPA is implemented on former fishing grounds, fishing fleets have two choices: exit the fishery or move to new fishing grounds. When MPAs redistribute effort away from one area and into another, they can accidentally increase fishing effort on species found within the new fishing grounds [@abbott2012]. This problem of policy "leakage" is well known in many environmental fields [@citations] but the general properties of this problem have received relatively little attention in the theoretical or empirical MPA literature [see @abbott2012;@savina2008;fulton2011a;@baum2003 for notable exceptions]. We fill this gap by

- Assessing the conditions under which this may be a problem using a generalized simulation environment
  - General insights about design strategy, effect of varying degrees of habitat overlap, etc, tradeoffs with fisheries

- Given that "it depends" nature of the problem, assess the specific potential for this problem in a case study
  - specific insights about design strategy, effect of varying degrees of habitat overlap, etc, tradeoffs with fisheries

Our goal here is not to provide tactical recommendations on the design
of blue-water MPAs, but rather to provide a strategic evaluation of the
sensitivity of bycatch populations and targeted fisheries to key
unknowns and design choices in MPA planning.

<!-- The outcome of MPAs for fishing and conservation are highly sensitive to -->

<!-- a wide range of economic and ecological factors. In order to constrain -->

<!-- the possible parameter space, we tune a simulation model based around -->

<!-- best available knowledge of the species and fleets operating in the -->

<!-- areas of the western central pacific ocean (WCPO) while allowing key -->

<!-- unknowns to vary, in particular the nature of adult and larval movement, -->

<!-- the degree of habitat overlap between different species, the economic -->

<!-- incentives facing fishing fleets, and the design strategy used in MPA -->

<!-- placement XX. Most conventional MPA design models are either static, -->

<!-- single species, or both. We present a dynamic, multi-species simulation -->

<!-- model that allows us to cohesively assess the impact of MPAs on a suite -->

<!-- of by-catch and targeted species collectively affected by multiple -->

<!-- fishing fleets.  -->

<!-- We show - something about how when species are homogeneous in space, -->

<!-- conservation gains are basically guaranteed except for constant catch -->

<!-- scenearios, which need to be considered in tuna-type scenarios - When we -->

<!-- introduce heterogeneity in stock structure roughly reflective of -->

<!-- distributions in the WCPO we do / do not begin to see tradeoffs in -->

<!-- conservation outcomes among bycatch species - quantify if there is a -->

<!-- concave function to this - Specific design paradigms are / are not more -->

<!-- / less resilient to this issue - Compare tradeoffs between MPA designs -->

<!-- that minimze internal bycatch tradeoffs (there have to be none if MPA = -->

<!-- 100%) and fishery outcomes -->

<!-- We show that MPAs increased biomass of by-catch species in XX% of -->

<!-- simulations, and decreased them in XX% of simulations. On average a 1% -->

<!-- increase in by-catch species resulted in a XX% change in fishery -->

<!-- catches... Outcomes were most sensitive to XX and XX, and least -->

<!-- sensitive to XX and XX. We do not suggest that these same results would -->

<!-- hold for other systems, but rather demonstrate the importance / -->

<!-- resilience of MPAs to key uncertain variables. In particular for pelagic -->

<!-- ecosystems, our work demonstrates the importance of considering the -->

<!-- collective impacts of space, vulnerability, life history, and fleet -->

<!-- dynamics in MPA design for conservation. -->

# Methods

@hastings2017 employed a simply model to derive conditions under which
MPAs would benefit bycatch species. However they also pointed out that
even small increases in model complexity quickly make analytical
solutions to the effects of MPAs intractable. As such, we based our numerical
analysis on a spatially explicit, multi-species, multi-fleet,
bio-economic model we call
[`marlin`](https://github.com/DanOvando/marlin). Models of MPAs span a
wide range of complexity, from simple analytical tools to end-to-end
ecosystem models [@fulton2015]. `marlin` follows a "models of
intermediate complexity" (MICE) philosophy [@plagányi2014] targeted
specifically around the role of technical interactions across species.
Many MPA models simulate the impacts of MPAs on a single-species level [@ovando2021XX]. This can be useful for for example examining the ways in which species-specific life history traits interact with fleet dynamics to influence MPA outcomes. However, bycatch problems are driven by technical interactions, in which fishing effort for a target species incidentally captures another species xx how do we want to talk about this in here? Like usually incidentally, but clearly sometimes people try and catch sharks too xx. `marlin` accounts for this by allowing the user to simulate an arbitrary number of fleets capable of catching an arbitrary number of species in a spatial area. Different fleets can place different economic value on different species, while also having species-specific selectivity functions. This allows us to simulate bycatch dynamics of MPAs in pelagic ecosystems.

We first ran a series of simulation exeriments designed to explore broad sensitivities of bycatch and fisheries to key factors.

We then ran a focused case study intended to reflect some of the traits of the tuna fisheries of the WCPO, holding constant...

## `marlin`

Lots of equations. Pretty standard age-structured population models, extra bells and whistles that need explaining

- Design of movement and habitat
  - Movement is a product of distance and habitat, such that critters will all else equal move to the closest patches with the best habitat

- Ontogenetic shifts

- Nature of fleet tuning (shared effort different q per species)

-

## Simulation Experiments

Given the complexity of the operating model used here, a complete sensitivity analysis across all parameters is not practical. Instead, we focus on the role of XX key attributes of pelagic ecosystems: the degree of spatial overlap between populations, the presence of seasonal migrations, and the potential for ontogenetic shifts in habitat use, and degree of fishing pressure on each species.

## Species & Fleets

Each simulation considers `n_distinct(marlin_inputs$scientific_name)` species comprised of the top 12 species by landings reported in the tuna fisheries of the Western and Central Pacific Ocean (WCPO). Rough Life history parameters are pulled from `FishLife` [@thorson2020]. For teleost fishes, we assumed that fecundity was proportional to weight-at-age times the maturity ogive. For cartilaginous species, we adopt a fixed fecundity relationship of XX pups per mature individual XX check on what I actually implemented here XX. The unfished spawning stock biomass for each of the included species was tuned in the model to reflect best available estimates from stock assessments.

Each simulation includes a longline and a purse-seine fleet. Each species can have a separate price (USD/kg) and selectivity ogive for each fleet, as dictated by the literature.

Fleets exert a given amount of effort in a patch each time step. The catch resulting from that effort is given by the Baranov [@baranov18XX] equatino

$$C_{f,s,p,t} = \frac{q_{f,s}E_{f,p,t}}{q_{f,s}E_{f,p,t} + m_s} \times \sum_{a=0}^{A}B_{s,p,t,a} \times (1 - e^{-(q_{f,s}E_{f,p,t}S_a + m_s)}) $$

Where $C$ is catch by fleet $f$ of species $s$ in patch $p$ in timestep *t*. $E$ is total effort by that fleet in a patch and time step, $m$ is natrual mortality for a given species, $B$ is biomass at age *a*, and $S$ is selectivity.

$q$ is the catchability coefficient for fleet $f$ and species $s$.

Fleets distribute themselves in space according to revenue per unit effort

$$R_{f,p,t} = \frac{\sum_{s=1}^{S}v_{f,s}C_{f,s,p,t}}{E_{f,p,t}}$$
Where $v$ is the price per kg of each species. Bycatch in the model is defined through the price. An unregulated species with no commercial value would have $v = 0$, whereas a regulated and prohibited species might actually have a negative price.

Effort per fleet is set a a constant of 1 per patch, and the model then tunes $q_{f,s}$ such that given all the other parameters of the model the fishing mortality rate for each species matches the best available level for that species. This is what allows the model to capture bycatch dynamics: A fleet can distribute itself purely based on revenues from target species, but through the $q$ parameter still exert fishing mortality on other species, even if that species has a price of zero, if those bycatch species inhabit the same patches as the more valuable target species.

### Species Overlap

We simulate degrees of species overlap using two parameters, $\sigma_c$ and $\sigma_h$. All species start with the assumption that the centroid of their preferred habitat occurs in the center-most patch of teh two dimensional system simulated by marlin. For our simulations, we then generated a random deviation from that default centroid for each species using $\sigma_c$ in the manner

$$centroid_s \sim normal(center,\sigma_c)$$
where *s* denotes species, and *center* is the coordinate of the central cell. In this way, the larger $\sigma_c$ grows, the more variation there is betwen the centroid of the preferred habitat for each species.

We then control the spread of a creatures preferred habitat around that centroid per a Gaussian dispersal kernal in the manner

$$habitat_{p,s} = \frac{1}{\sqrt{2\pi\sigma_h^2}}e^{-\frac{d_p^2}{2\sigma_h^2}}$$

Where $d$ is the euclidian distance from the $centroid_s$ to patch $p$ and $\sigma_h$ controls the dispersal of the habitat around that range.

### Ontogenetic Shifts

`marlin` allows for different habitats to be used by adults vs. recruits. For scenarios with ontogenetic shifts ,we assign recruits to a habitat grid that is the inverse of $habitat_{p,s}$. In this way organisms enter the model living on one area, and then transition to adult habitat at older life stages.

XX this isn't very interesting at this point since movement is so high so I vote we get rid of it for now XX.

### Seasonal Migration

For this paper `marlin` is set to run with two seasons per year. When seasonal movement is turned on, animals use one habitat matrix in season one, and the inverse of that habitat matrix in season 2.

### MPA Design Strategies

MPA design is governed by four general parameters: the number of cells to place inside an MPAs, and the design strategy used to place the MPAs, the species considered in the MPA planning process, and the accuracy of the information used in the planning process.

We consider `r n_distinct(experiment_obj$placement_strategy)` different MPA placement strategies.

For the "depletion" strategy, MPA locations are prioritized based in proportion to the spawning stock biomass (SSB) of the most depleted species. So, patches with high amounts of highly depleted species are prioritized over patches with high biomass of lightly depleted species.

For the "rate" strategy, MPAs are placed in based on the depletion-weighted catch relative to the total catch in a patch. So, patches where with high rates of catch of depleted species relative to totcal catch are prioritized.

For the "avoid_fishing" strategy, MPAs locations are prioritized inversely to fishery catches. So, patches with low catches are prioritized over patches with high catches.

For the "target_fishing" strategy, MPAs locations are prioritized proportionally to fishery catches. So, patches with high total catches are prioritized over patches with high catches.

For the "area" strategy, we simply assign patches in sequential order until the desired total MPA size is reached.

The information used to inform each MPA design strategy (except area) can include a multiplicative observation error term $\sigma_{obs}$ where

$$information \sim lnorm(data,\sigma_{obs}) $$

For example, for the depletion scenario, $\sigma_{obs}$ controls the accuracy of the spatial depletion data used to prioritize MPA placement.

## Case Study

In order to constrain the parameter space somewhat, we also ran a series of case study experiments. Rathere than interating over different habitat configurations and movement dynamics, we fix habitat for each species based on the results of @bradley2021XX, and set ontogenetic movement and seasonal migrations to FALSE.

### MPA Network Optimization

For the case study section, as a point of comparison we calculate an optimized MPA network along with the "rule of thumb" networks based on the design strategies outlined in section XX. The optimization routine is based on an objective function of the form

$$obj = \alpha \times scale(\sum_{s=1}^{S}\frac{ssb_s}{ssb0_s}) + (1 - \alpha) \times scale(\sum_{s=1}^{S}R_s) $$

where $\alpha$ controls the relative weight assigned to a "conservation" objective based on depletion, relative to an "economic" objective based on revenues. The 'scale' function rescales both objective components to the range of 0 to 1 to facilitate interpretation of $\alpha$ xx come up with a better version of this XX

The optimization routine itself works through a sampling-importance-resampling algorithm. In the first iteration of the algorithm, the marginal contribution of each individual patch to the objective function is calculated. The patch with the highest objective function value is then placed in an MPA. For the remaining iterations, we sample *X* cells from the remaining non-MPA cells with probability proportional to their marginal value to the objective function in the last iteration. We then calculate the marginal value of each of these cells when individually added to the MPA network begun in the prior step. The cell with the higheset marginal value is then added to the MPA network, and the marginal values of the non-selected but sampled cells are updated based on the values in that iteration. The process then continues until the desired MPA size is reached.

## Assessing Performance

Each model run started at equilibrium conditions and was then run for `r years` years (`r seasons * years` seasons. We assessed performance by comparing relevant metrics (e.g. depletion and revenue) under a given scenario with MPAs relative to the values of those same metrics in the simulated counterfactual world identical in every way save for the presencee of MPAs.

# Results

## Experiments: How severe and undere what conditions are unintended conservation consequences? (clearer name for these)

Make clear that a goal of all this complexity was to see if there were any clear emergent properties (this thing matters way more than this thing). We highlight some of those, but given the complexity we found we present the complexity of interacting ways in which different things can effect the conservation outcomes of MPAs

XX after looking through these should widdle it down to only things you actually examine, annoying / no point to having a bunch of unexplained minor variation in the mix relatede to things like ontogenetic shifts) xx

xx the problem here. These things all interact in so many complex ways that it's hard to look at either graphically or statistically, and maybe for the best. i.e. the answer isn't "if habitat overlap is less than XX% you might have a problem, otherwise you're fine", but rather that a bunch of things can interact to cause potential problems in complex ways. but that's hard to explain xx.

Big message: even constrained by the same initial depletion and core biology (e.g. movement rates), a wide range of outcomes are possible for each species for a given MPA size, dependent on the design strategy, habitat overlap, migrations, quality of information, etc. Across the board as MPA size increases the ratio of positive to negative outcomes increases, until all species achieve their maximum conservation potential in every iteration at 100% MPA. HOwever, every species experienced a set of simulations, roughyl xx%, that resulted in net losses in SSB/SSB0 relative to BAU. Losses were smaller than potential gaines, but reached meaningful levels in some species, and interestingly peaked in many cases right before max MPA (reflecting concentration of fishing in small amounts of habitat in habited by the species).Important to stress in here that there are no probabilities in this, no way of knowing which state is more probable than others, just given the range of things simulated more ways for a bigger positive effect. Also make clear though, biggish losses (10% or more) possible, but rare.

```{r exp-results, fig.cap="Results of simulation experiments on change in SSB/SSB0 (y-axis) as a function of MPA size (x-axis). Color indicates number of simulations in log-10 space that fell into a particular effect size at a given MPA size"}
tmp <- results %>%
  left_join(excols, by = "xid") %>%
  ggplot(aes(prop_mpa, biodiv - bau_biodiv)) +
  geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
  geom_bin2d() +
  facet_wrap( ~ critter) +
  scale_x_continuous(labels = scales::percent, name = "% MPA") +
  scale_y_continuous(name = "Change in SSB/SSB0") + 
  scale_fill_viridis_c(trans = "log10")

tmp

```

Moving to expressing things are percentage of species with some / significant loss

key result: Assuming perfect information, degree of habitat matters buts its messy. Both the depletion and target_fishing strategies are capable of completely avoiding losses if all species share the same habitat. As habitat non-overlap (xx better phrase) inceases, this gets harder to do, and both incur some losses. All of the other placement strategies exhhibit some losses, even with perfect overlap, at some MPA size.

```{r hab-effect, fig.cap="Precent of species with some degree of loss in SSB as a function of MPA size, MPA design strategy, and degree of habitat non-overlap (sigma_c), come up with a better name for this. Perfect information, no seasonal movement, all species considered in planning."}
fig_3 + 
  scale_y_continuous(labels = scales::percent, name = "% of Species with Loss in SSB/SSB0") +
  scale_color_discrete(name = "Habitat Non-Overlap") + 
  theme(legend.position = "top")
```

moving away from perfect information though, things get messier: now just looking at the same plot but with all the other options turned on things get messy quickly.

```{r mess-effect, fig.cap="Precent of species with some degree of loss in SSB as a function of MPA size, MPA design strategy, and degree of habitat non-overlap (sigma_c), come up with a better name for this. Rows indicate presence or absence of seasonal movement."}

tmp <- total_results %>%
  ggplot(aes(prop_mpa, loss, color = factor(sigma_centroid))) +
  geom_jitter(alpha = 0.1) +
  geom_smooth() +
  facet_grid(seasonal_movement ~ placement_strategy) + 
    scale_color_discrete(name = "Habitat Non-Overlap") + 
    scale_y_continuous(labels = scales::percent, name = "% of Species with Loss in SSB/SSB0")+
  theme(legend.position = "top")

tmp
```

same thing but now for SSB loss greater than or equal to 10%. Maybe should make all figure this? Basically, pretty common to cause some technical loss, but it's usually pretty marginal and effectively zero, so no benefit but not really much loss either. Significant losses can happen but less common, though throughout need to be really careful about being clear that this nothign about the probability of any one of these states of nature.

```{r signif-mess-effect, fig.cap="Percent of species with more than 10% loss in SSB/SSB0 as a function of MPA size, MPA design strategy, and degree of habitat non-overlap, come up with a better name for this. Rows indicate presence or absence of seasonal movement."}

tmp <- total_results %>%
  ggplot(aes(prop_mpa, signif_loss, color = factor(sigma_centroid))) +
  geom_jitter(alpha = 0.1) +
  geom_smooth() +
  facet_grid(seasonal_movement ~ placement_strategy) + 
    scale_color_discrete(name = "Habitat Non-Overlap") + 
      scale_y_continuous(labels = scales::percent, name = "% of Species with more than 10% Loss in SSB/SSB0")+

  theme(legend.position = "top")

tmp
```

Huh the seasonal migration thing is counterintuitive, but now that I think about it maybe makes sense, since with migrations every species gets some protection and some fishing, rather than some getting a lot of protection and others a lot more fishing.

My prior: that the magnitude of the benefits will be smaller with seasonal migration. So, you get a sort of tradeoff where on the plus side seasonal migrations means less change of unitended consequences, but less overall benefits as well

Yep that mostly checks.

```{r signif-mess-biodiv-effect, fig.cap="Percent of species with more than 10% loss in SSB/SSB0 as a function of MPA size, MPA design strategy, and degree of habitat non-overlap, come up with a better name for this. Rows indicate presence or absence of seasonal movement."}

tmp <- total_results %>%
  ggplot(aes(prop_mpa, delta_biodiv, color = factor(sigma_centroid))) +
  geom_jitter(alpha = 0.1) +
  geom_smooth() +
  facet_grid(seasonal_movement ~ placement_strategy) + 
    scale_color_discrete(name = "Habitat Non-Overlap") + 
      scale_y_continuous(name = "Total change in SSB/SSB0")+

  theme(legend.position = "top")

tmp

```

The other thing to consider is quality of information... weird results so far.

Poorer information decreases benefits in some cases...

```{r mess-biodiv-voi-effect, fig.cap="Precent of species with some degree of loss in SSB as a function of MPA size, MPA design strategy, degree of observation error. Rows indicate presence or absence of seasonal movement."}

tmp <- total_results %>%
  ggplot(aes(prop_mpa, delta_biodiv, color = factor(placement_error))) +
  geom_jitter(alpha = 0.1) +
  geom_smooth() +
  facet_grid(seasonal_movement ~ placement_strategy) + 
    scale_color_discrete(name = "Placement Error") + 
    scale_y_continuous(name = "Change in SSB/SSB0")+
  theme(legend.position = "top")

tmp
```

But has a confusing effect on unintended consequences. I guess makes some sense if unintended consequences are a consequence of very targeted interventions, and the greater observation error becomes, the close MPA placement gets to random, and the closer it gets to random, the less likely you are to start hammering one species while benefiting another, rather all just get some weaker protection? Kind of interesting if so.

```{r mess-loss-voi-effect, fig.cap="Precent of species with some degree of loss in SSB as a function of MPA size, MPA design strategy, degree of observation error. Rows indicate presence or absence of seasonal movement."}

tmp <- total_results %>%
  ggplot(aes(prop_mpa, loss, color = factor(placement_error))) +
  geom_jitter(alpha = 0.1) +
  geom_smooth() +
  facet_grid(seasonal_movement ~ placement_strategy) + 
    scale_color_discrete(name = "Placement Error") + 
    scale_y_continuous(labels = scales::percent, name = "% with loss in SSB/SSB0")+
  theme(legend.position = "top")

tmp
```

## Case Study

The general experiments conducted show two general insights

1. Even conditional on fishing exploitation and general life history strategy, the effects of MPAs on a group of species with technical interactions is highly variable and dependent on habitat distribution, design decisions, and data quality. xx make this more explicit.

2. Some degree of unintended conservation consequences occured across nearly all of our simulations below MPA coverage near 100%, though small losses were much more common than large. Different design strategies are better suited to avoiding unintended consequences.

There two general results beg some important questions. First, are unintended consequences inevitable given the kinds of technical interactions explored here, or can optimized MPA networks reduce or eliminate these unintended consequences. Second, one clear solution to the problem of unintended consequences is to simply make the MPAs large: if MPAs cover 100% of the area there can be no direct unintended conservation consequences. The potential for negative impacts on economic fishery outcomes increases as MPAs reach large size though. What then are the tradeoffs between reducing the potential for unintended conservation consequences and fishery revenues?

To address these questions, we ran a more limited "case study" set of simulations, based on the dynamics of the WCPO tuna fisheries. Rather than iterating over different movement dynamics and habitat configurations, we fix habitat layers on the values predicted by @bradley2021, and xx other things we fix xx.

For this subset of simulations then, along with the various "rule of thumb" MPA design strategies we ran an optimization routine (see section XX) across a range of objective function weightings, in the manner of @rassweiler2014. This allowed us to assess 1) the degree to which unintended conservation consequences of MPAs can be reduced by optimized design and 2) the magnitude of tradeoffs between conservation outcomes and fishery revenues under both rule-of-thumb and optimized MPA networks.

key results: optimization helps (Nobel prize please). Can reduce percent of species with conservation loss down to 25% pretty quickly if you prioritize conservation. Rule of thumb designs like targeting fishing can get there with perfect information and when considering all animals, but can be pretty far from optimal with imperfect information or when only considering a subset of animals in the planning process (that's what all the different points represent). Some rules of thumb are more sensitive to these decision than others. E.g. rate can get you optimal, but goes off the rails pretty quickly with bad information.

There is also a pretty clear breakpoint in results between basically trying for conservation or trying for economics, and not a lot of wiggle room in between (the two general bands of lines). xx check to what extent this depends on the way you scale the two objective function components xx.

```{r, fig.cap="Percent of species with some conservation loss as a function of MPA size. Points are draws from rule-of-thumb scenarios, lines are results of optimization routine across a gradient of conservation weighting."}
total_experiment_obj %>%
  filter(prop_mpa > 0) %>%
  ggplot(aes(prop_mpa, loss)) +
    geom_point() +
  geom_smooth(data = total_opt_obj, aes(p_protected / 100, loss, color = alpha, group = alpha), se = FALSE) +
  scale_x_continuous(labels = scales::percent, name = "Area in MPA") +
  scale_y_continuous(labels = scales::percent, name = "% Experiments With Conservation Loss") +
  facet_wrap(~placement_strategy) +
  scale_color_viridis_c(name = "Conservation Weighting") + 
  theme(legend.position = "top")
```

Key results: Even with optimization, you can basically either be in a world with 50% or more of species having some level of unintended conservation consequence, but no economic loss / some economic gain, or you can be in a world with lower unintended consequences, but economic losses. But, even rule of thumb approaches can get you to some places with relatively little unintended consequences at relatively minimal economic loss.

```{r, fig.cap="Tradeoff between unitendedn consequences and fisheries revenue. Filled polygons in background represent rule-of-thumb results, with redder colors indicating higher density of points. Points are results of optimization, with color indicating conservation weighting in optimization. Black diagonal line shows 1:1 relationship"}

cs_fig_11

```

Key results: Presence of and magnitude of unintended consequences pretty sensitive to design descisions. On a macro scale though far less so. In terms of change in total SSB/SSB0 (metric of biodiversity, i.e. maxed out when all species are unfished) and change in revenue, there is basically a pretty simple linear tradeoff between biodiversity and revenue, with a slope less than negative 1, meaning that 1% increase in biodiversity comes at a greater than 1% loss in fishery revenue. This stacks up pretty nicely with @rassweiler2014., where value of optimization was low when overfishing of valuable species isn't prevalent, as is the case here. So, design and optimization can help a lot with the more "micro" level outcomes (e.g. making sure that biodiversity gaines aren't coming from massive gains in one species to detriment of other), but on a macro level pretty insensitive in this case study.

```{r, fig.cap="Tradeoff frontier between revenue and biodiversity xx across different rule odf thumb design strategies. Points indicate role of thumb results, lines are results of optimized networks across a gradient of conservation weighting."}

cs_fig_10 + 
  theme(legend.position = "top")
```

# Discussion

**what is the story**

- Unintended conservation consequences of MPAs resulting from technical interactions have been discussed in the literature, but almost exclusively in the context of highly tactical modeling or localed empirics. We conduct one of the first large-scale assessments of the potential of this problem across a range of scenarios

- We find that these unintended consequences can be common outside of "canonical" MPA models *a la* hastings etc.

- But, good news is magnitude of unintended consequences were usually small, and total benefits almost always outweigh individual species losses (again no probability of states though)

- Some design strategies are better at avoiding unintended consequences than others, but sensitive to data quality. Optimization can help, but even with optimization unintended consequences can happen.

- Under the levels of fishing mortality in the WCPO, optimization and design can help reduce tradeoffs between fisheries revenue and unintended consequences than others, but macro-level tradeoffs between biodiversity and fishery revenue are very insensitive to design choices, with a pretty clear "increasing biodiversity through MPAs reduces fishery revenues". 
  -  xx This is clearly hyper-sensitive to the one thing we don't iterate over much here: how overfished are the target species. At the moment "they're not". But, this runs fast enough, is it worth running a hypothetical case study where they are just to illustrate how things change?

# Conclusions

- Unintended conservation outcomes of MPAs totally possible outside of "consider a homogenous fish blob" models

- But, usually small, and net conservation benefits outweigh conservation costs

- Should pay particular attention to potential though when species near tipping points are of concern: hard to imagine those not being part of the prioritization but something to consider. 
- Other interesting things.

# References

::: {#refs}
:::

# Scraps

Thinking through the “naive” MPA design strategies to compare against the optimized outcomes. Thoughts at the moment below. These can be run in a few seconds each, so we cab be be pretty liberal in coming up with combinations if there are other ideas people want to try, but also if some of these seem pointless can drop as well!
6:11
Design strategies
1. Assign protection by SSB / SSB0 weighted abundance in space (protect the highest concentrations of the most threatened species)
2. Assign protection by bycatch rates (distribution of biomass bycatch / biomass target)
3. Assign protection by targeting fishing effort
4. Assign protection by avoiding fishing effort
Knowledge states
1. Degrees of accuracy over the thing being measured (i.e. error in the design strategy)
2. Number of species considered (e.g. design aronnd only 2 bycatch species instead of all)
3. über naive: just close off random patches up to a specified size
We then plot these points in objective function space and compare to the efficiency frontier / consider frequency and magnitude of conservation and economic tradeoofs
The question all this is answering: how sensitive are conservation and economic outcomes to design strategy / uncertainty when there are technical interactions and heterogeneity in habitat? The first stage is teeing up the idea that beyond conservation / yield tradeoffs, as you increase habitat heterogeneity with technical interactions you increase the potential for conservation tradeoffs. But, the degree is going to be super context dependent, so to dive in a bit we then focus on this case study of the particular species / habitat overlap in the WCPO. Further work might reveal more generalities in this, but htis is basically getting a sense for the potential magnitude of the challenge.

XX stress a key difference: may other papers on this are static, in that
they just some up displacement on different layers, rather than
addressing dynamic respones over tim xx

We show....
  - sensitivity of both conservation and fishery outcomes to pelagic realism and design choices
  - compared to targeted / dynamic intervention... 
  - value of information would be great...

Contrast to coastal MPA theoretical work
  - @hastings2017
  - @ovando2021
  - @white2010
  - not quite sure what to do about points raised in @fulton2011a
    - to what extent has this been done...

The paragraph from @fulton2011a that broke me...

"
Spatial management was insufficient as the dominant management lever because it proves quite difficult to strike the balance between protecting areas that are of a size sufficient to produce any conservation benefit and providing sufficient fishing grounds to sustain an econom- ically viable industry (which is not possible in many fisheries if closures on the order of 70–80% or more are imposed, Fulton et al. 2007). More- over, Savina etal. (2008) found that effort displaced by closures potentially undermines any benefits coming from the zoning. For instance, strong improvements in diversity (by nearly 20%) under heavy fishing pressure occurred if effort outside a marine protected area (MPA) was controlled. However, without effort control, there was a significant (>20%) drop in overall diversity. Fulton et al. (2007), Savina et al. (2008), Smith et al. (2010) and unpublished work associated with Horne et al. (2010) all found that while spatial management appears to be an effective means of satisfying conservation objectives con- cerning habitat and restoring age-structures of less mobile species, benefits from closures for individual species depend on its trophic position, degree of mobility, geographical distribution and extent of ontogenetic shifts in habitat use and depth (highly mobile or prey species do not benefit in the way that more site attached or predatory species do)."

Lots of studies have shown that lots of things matter, can we provide guidance on which things are priorities?

Where does this fit in context of @smith2021? One important difference is that they point out differences in effectiveness, where effectiveness is still almost always positive, since they just measure leatherbacks. So their fig 5, there are a few cases where leatherback catches actually go up, but it's rare. So, extending to multiple species and keeping track off all of them shows it's even harder

How does this contrast with the the large and growing dynamic ocean management lit?
  - By definition, with perfect information dynamic $\geq$ static spatial controls
  - DOM literature is mostly static, tracks displacement onto static abundance / risk layers
  - Doesn't allow for dynamic impacts, for example fleet adjustments based on updated economic surface, or short-term costs for long-term benefits (can overstate costs?)
  - Dynamic model allows for delta between dynamic policies, e.g. Fmsy vs DOM
    - analytical models a la hastings have value but do / don't stand up to realism

Presentation of the model as a MICE model for MSP
  - reasonable parameterizable
  - flexible
  - fast

Focus of debate has been on fisheries and yields. As MPAs move to pelagic systems though, we show that conservation outcomes also depend on correctly identifying ecological state / design decisions etc.

What are the considerations of protection you need to make good
decisions. It's not always obvious how to design things to get the
outcome you want. If we could make the case that something like closing
the high seas is insufficient conservation management without
conservation management inside EEZs / what is the

Need to do an optimization here, calculate the marginal value of added
cells.

What's this about? Pointing out a problem, which to make impactful would
require making the case that there is a problem with current design
strategies. e.g.. "MPAs are often designed like this (simple dispersive
models with constant habitat, or MARXSAN style things), and doing that
does/does not create problems A tactical evaluation. Picking a handful
of high profile MPAs, parameterizing all the knowns as well as possible,
iterating over unknowns, and see what the range of outcomes are.
Cleaner, more impactful, but also opens it up (depending on the results)
to a details fight about very specific modeling choices How plausible
are losses? Maybe we drill down on this a bit. From a purely bycatch
perspective, the usual questions of fisheries tradeoffs aren't on the
table (but can add to the discussion considerations of how much target
fishing you give up). Given that then, while it's possible that
"targeted" interventions would work better (ignoring costs/enforcement,
etc), there's really no reason not to do MPAs. Unless, you actually end
up concentrating fishing pressure outside on populations. So, we
consider how plausible is it for that to really happen. Issue here is
that seems likely to be pretty hand-wavey. Most likely that the vast
majority of simulations will show positive impacts, but have no way of
assigning probabilities to any one state of nature.

The Role of Blue-Water Marine Protected Areas in Bycatch Reduction

What's the question: How effective are Blue-Water MPAs in reducing
fishing mortality of important pelagic bycatch species?

Another angle. There's going to be a lot of unknowns in designing BWMPA.
How much does not fully understanding a few key life history (e.g.
movement) matter, and how much does the role that the species plays in
the broader fleets portfolio matter

Could simulate habitat from a multivariate normal, where you use the
sigma to control how similar rasters are across species

Marine protected areas are increasingly proposed to protect biodiversity
in pelagic systems. However, maby paradigms of mpa design come from
coastal systems. Here we ask, how likely are blue-water MPAs to provide
conservation benefits to bycatch species, and what are the key
uncertainties underlying these outcomes? We begin with a general
discussion of the ways in which traits of open-ocean systems might
influence the conservation performance of BWMPAs. Given the massive
parameter space, we then constrain ourselves to a styulized version of
the western central pacitic ocean to ask, how feasible is it to provide
broad protection of bycatch species through common MPA design paradigms
in the context of this system. Our results demonstrate the critical
importance of XX and XX in designing MPAs in pelagic systems, and
highlight the potential risks of not accounting for XX. Introductionhere

Bycatch is a significant driver of population declines for many pelagic
species Sharks, seabirds, mammals, etc MPAs increasingly turned to as a
solution for conservation problems, etc. MPA science is largely based on
dynamics of coastal systems. Open-ocean ecosystems have very different
dynamics that require dedicated research Burgess et al. (2018) estimated
the reductions in fishing mortality needed to halt declines for a number
of bycatch species Here we ask, what are the key drivers of bycatch
conservation outcomes from MPAs? There are all of these species; you
reduce fishing mortality. You can either do fisheries management, or you
can do MPAs, let's focus in on the MPA side. This could reduce fishing
mortality, or it could do, and it depends on the dynamics of the species
and the fleets. When can MPAs reduce fishing mortality enough for
bycatch species? And where can you go wrong?

It doesn't make any sense to ask "do MPAs help bycatch": the question is
obviously it depends. To quickly make this case, we'll assume that a
species is in need of recovery. We then do the simple mako example to
show that effects of spatial closure depend on the overlap of habitat
and the assumption of fleet dynamics, and provide some empirical
evidence that in particular effort redistribution is not unlikely.

Determining the effect of an MPA network on bycatch then will also
depend on trying to get close to a tactical model. But, this creates a
problem for blue water systems in particular. Can make the argument that
given the centrality of relatively static habitat features in coastal
MPAs "protect the key habitats" might not be a bad strategy from a
conservation strategy. But in highly mobile pelagic systems there (may
be) a lot more variables in play. Where should researchers direct their
attention?

To reduce the parameter space, we constrain ourselves to a system where
1) we know the species involved 2) we have some sense of the relative
depletion and value of the different species and 3) some basic life
history of the species. We then vary commonly unknown parameters,
particularly preferred habitats, spatial catchability, movement rates
etc, and pair these simulations with different MPA designs, and examine
the conservation outcomes for bycatch species.

What i don't like about this

There isn't an "answer" to this: baring something like "changing the
type of density dependence never had any effect greater than XX%", for
anything that on occasion has an impact the answer of "how much" will
almost always depend on what else is going on.

So, it seems like there are two real angles of attack here

Pointing out a problem, which to make impactful would require making the
case that there is a problem with current design strategies. I.e. "MPAs
are often designed like this (simple dispersive models with constant
habitat, or MARXSAN style things), and doing that does/does not create
problems A tactical evaluation. Picking a handful of high profile MPAs,
parameterizing all the knownns as well as possible, iterating over
unknowns, and see what the range of outcomes are. Cleaner, more
impactful, but also opens it up (depending on the results) to a details
fight about very specific modeling choices

Methods

Collect life history for key bycatch species and target species, cite
Bradley et al. pulling together species distribution and threats etc.
Introduce marlin Parameterize marlin runs with "knowns" from life
history of bycatch and target species Highlight key "unknowns" that we
will iterate over, e.g. fleet models, density dependent movement, timing
of density dependence etc Simulate effects of MPAs given reasonably
realistic situations but iterating over key unknowns Evaluate the range
of outcomes and key drivers of MPA effects I'm thinking we try and avoid
a really specific case study but hard not to accidentally do one without
making everything up....

Results Range of likely changes in fishing-induced mortality for the
species selected Key drivers of uncertainty in those outcomes, focusing
on things that switch states from reduction to increase in mortality How
common are conservation tradeoffs between bycatch species? Fishery
impacts

Discussion What does this say about BWMPAs and bycatch? Not going to get
into design decisions here, depending how common tradeoffs are I think
that will be the main focus Highlight any species that frequently
induced tradeoffs Tee up future research MPAs in context of other
management MPAs as risk reduction strategy Etc.