### 2021-10-12 documents/03_marlin_paper.Rmd
finally back to main

---
title: "Opportunities and precautions for the use of MPAs in bycatch reduction"
author:
    - Daniel Ovando
    - Darcy Bradley
    - Lennon Thomas
    - Echelle Burns
date: "`r Sys.Date()`"
output: 
  bookdown::word_document2:
    reference_docx: template.docx
  bookdown::html_document2: default
  bookdown::pdf_document2: default
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
bibliography: references.bib
params:
  run_name: ["v1.1"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, include = TRUE)

source(here::here("scripts", "00_setup.R"))

theme_set(marlin::theme_marlin())

sci_to_com <- taxize::sci2comm(unique(marlin_inputs$scientific_name)) %>% 
  bind_rows(.id = "critter") %>% 
  pivot_longer(everything(),names_to = "critter",values_to = "common_name") %>% 
  mutate(common_name = stringr::str_to_title(common_name))

capitalize <- function(string) {
  substr(string, 1, 1) <- toupper(substr(string, 1, 1))
  string
}

```

```{r load-results}

results_files <- list.files(file.path(results_path))

rdses <- results_files[str_detect(results_files,'.rds')]

walk(rdses,~assign(str_remove(.x,".rds"), read_rds(file.path(results_path,.x)), envir = .GlobalEnv)) #load and assign all .rds objects in the appropriate results folder

rdatas <- results_files[str_detect(results_files,'.Rdata')]

walk(rdatas, ~load(file.path(results_path,.x), envir = .GlobalEnv))

```

<!-- \*\* something to point out: increased catches are a paradox, either -->
<!-- come from rebuilding stock or concentrated fishing, but trajectory can -->
<!-- tell you something. not possible for MPAs to increase catches without -->
<!-- decreasing catches first if they are rebuilding? \*\* -->

<!-- Seasonal movement is kind of interesting, basically just dilutes the -->
<!-- extremes -->

<!-- less resilient species are both more in need of protection and easier to -->
<!-- accidentally fish harder -->

# Introduction

<!-- MAKE IT MORE ABOUT BYCATCH-->

<!-- Empirically, how do you differentiate increase in catch from fishing harder vs increases in catch from rebuilding: rebuilding has to have a decrease at first -->

Marine protected areas (MPAs) are a growing part of the marine
conservation toolbox, with 5.9% of the world's oceans currently in
implemented MPAs [@grorud-colvert2021], and bodies such as the IUCN
calling for 30% protection by 2030. One of the potential benefits of
MPAs is their ability to provide protection to broad segments of an
ecosystem at once, rather than requiring targeted management based on
individual species that may be unfeasible in many parts of the world
lacking in data or fisheries management capacity. Their ability to
provide protection across a range of species with one policy has led them to be suggested as a tool to reduce problems around weak stock or "bycatch"
species [@hastings2017]. We use the term "bycatch" here to generally refer to the
capture of species within a fishery primarily targeting a different set
of species.

While the effect of MPAs on fisheries yields is a continuously debated
topic, the conservation impacts of MPAs are seemingly simpler, albeit
challenging to detect in many cases [@ovando2021a], due in part perhaps
to the perceived obvious nature of their benefits: if you prevent
fishing of a species within an area, there will be more of that species
in that area, and so the conservation benefits of MPAs will increase
with MPA size. @hastings2017 clearly illustrates this idea. Using a two-species two-patch analytical model, they derive conditions
under which MPAs can solve the "weak-stock" problem in which a less
productive species is caught as bycatch by a fleet targeting a more
valuable and productive stock.

However, as @hastings2017 point out, their results depend on a very
specific set of assumptions, namely that 1) adults are perfectly
sedentary 2) the target and bycatch species overlap perfectly in terms
of habitat and 3) recruits are perfectly mixed throughout the system. In
this paper We build on @hastings2017 to consider the role of MPAs in
reducing bycatch while managing tradeoffs with fisheries under a more
resolved model.

Specifically, the potential benefits of MPAs as a bycatch management are
complicated by two key issues: effort displacement and technical
interactions. One of the primary questions around MPAs is to what extent
do MPAs remove versus displace fishing effort. If MPAs remove fishing
effort, then their direct effects on target and bycatch species can only
be positive. If however MPAs displace fishing effort evaluating their
potential impacts becomes more complicated. The net conservation impact
of MPAs becomes a balancing act between the effects of reduced fishing
inside protected borders and increased fishing intensity outside MPAs.
The outcomes of this displacement can be further complicated by
technical interactions, i.e. the degree to which fishing effort for one
species affects another, particularly when species have markedly
different habitat distributions. For example, @abbott2012 demonstrated
that an MPA designed to protect red king crab (*Paralithodes
camtschaticus*) caused an unintended increase in bycatch of Pacific
halibut (*Hippoglossus stenolepis*) by the displaced bottom-trawl
fishery. Policy leakage is not a unique feature to MPAs, e.g. XX, but
the potential impacts of MPAs on target and bycatch species taking into
account the potential for these leakage effects has received relativelly
little generalized attention in the literature [see @abbott2012;
@savina2008; @fulton2011a; @baum2003 for notable exceptions].

This study addresses this gap using a novel bio-economic simulation
model that allows for efficient evaluation of the effects of MPAs on
target and bycatch species across a range of ecological and economic
dynamics, critically accounting for the potential impacts of effort
displacement and technical interactions. We use this model first to
explore general conditions that lead to positive and negative outcomes
for target and bycatch species from MPAs. We then present a case study
tuned to reflect the dynamics of the tuna, billfish, and shark
populations and fisheries of the Western and Central Pacific Ocean
(WCPO) to explore the value of optimal MPA design in managing tradeoffs
between maximizing bycatch conservation and minimizing economic impacts.
We chose the WCPO as a case study as the modern science of MPAs was
largely developed around coastal ecosystems, often characterized by
relatively static features such as coral reefs [citation]. However,
large MPAs are increasingly being implemented and considered across
pelagic ecosystems. These pelagic systems are characterized by highly
mobile species and fishing fleets using a diverse mix of static and
dynamic ocean features [@game2009; @lubchenco2003]. These ecosystems are
also home to combinations of relatively abundant species targeted by
fishing, such as many tuna stocks [@pons2016], and highly threatened
species caught as incidental "bycatch" in targeted fishing operations,
such as oceanic whitetip sharks [@pacoureau2021]. These dynamics,
combined with the increased policy interest in large MPAs covering
expanses of blue-water ecosystems, make pelagic systems such as the WCPO
an ideal case study in bycatch management through MPAs. The goal of this
paper is to provide a strategic evaluation of the sensitivity of bycatch
populations and targeted fisheries to key unknowns and design choices in
MPA planning.

# Methods

@hastings2017 employed an analytically tractable model to derive
conditions under which MPAs would benefit bycatch species. However they
also pointed out that even small increases in model complexity quickly
make analytical solutions to the effects of MPAs intractable. As such,
we employ a numerical simulation strategy in the form of a spatially
explicit, multi-species, multi-fleet, bio-economic model we call
[`marlin`](https://github.com/DanOvando/marlin). Models of MPAs span a
wide range of complexity, from simple analytical tools to end-to-end
ecosystem models [@fulton2015]. `marlin` follows a "models of
intermediate complexity" (MICE) philosophy [@plagányi2014] targeted
specifically around the role of technical interactions across species.
Many MPA models simulate the impacts of MPAs on a single-species level
[@ovando2021a]. This can be useful for for example examining the ways in
which species-specific life history traits interact with fleet dynamics
to influence MPA outcomes. However, bycatch problems are driven by
technical interactions, in which fishing effort for a target species
incidentally captures another species. `marlin` accounts for this by
allowing the user to simulate an arbitrary number of fleets capable of
catching an arbitrary number of species in a spatial area. Different
fleets can place different economic value on different species, while
also having species-specific selectivity functions. This allows us to
simulate bycatch dynamics of MPAs in pelagic ecosystems.

Below we provide a summary of the `marlin` model, as well as details of
the generalized simulation exercises and case study analyses.

## `marlin`

#### Fleet Dynamics

The key feature of `marlin` for the purposes of this project is the
ability to capture technical interactions across species. While marlin
does not simulate species interactions, fishing fleets in `marlin` are
capable of targeting and affecting multiple species simultaneously. This
allows us to capture for example the potential impacts of displacement
of effort from one area into another across multiple species.

A fishing fleet or *metier* is defined in `marlin` by a total pool of
effort that can have different catchability, prices, and selectivity for
individual species within a simulation. For example, both a longline and
purse seine fleet may capture bigeye (*Thunnus obesus*) and skipjack
(*Katswamus pelamus*), but the longline fleet is much more likely to
catch bigeye than skipjack, and *vice versa*. Each fishing fleet
generates catches, revenues, and profits from individual species that it
targets, and makes decisions based on its total objective function
across all species.

Revenues *R* for fleet *f* in time step *t* are given by

$$
 R_{t,p,f} = \sum_s^S \sum_a^A \Pi_{s,f}\frac{\alpha_{a,s,f}q_{s,f}E_{t,p,f}}{F_{t,p,s,a}} \times Y_{t,p,s,a}
$$

where *p* is patch, *s* is species, *a* is age, $\alpha$ is selectivity
at age, *q* is catchability , *F* is total fishing mortality, $\Pi$ is
price per unit weight, and *Y* is total catches.

$$
Y_{t,p,s,a} = \frac{F_{t,p,s,a}}{z_{t,p,s,a}} \times b_{t,p,s,a} \times (1 - e^{-z_{t,p,s,a}}) 
$$

where *F* is total fishing mortality, given by

$$
F_{t,p,s,a} = \sum_f^F\alpha_{a,s,f}q_{s,f}E_{t,p,f}
$$

Selectivity $\alpha$ is modeled through a logistic form per

```{=tex}
\begin{equation}
\alpha_{a,s,f}=\frac{1}{(1 + e^{-log(19)\times\frac{l_{a,s} - l^{sel}_{s,f}}{\delta^{sel}_{s,f}}})}
  (\#eq:sel)
\end{equation}
```
where $l$ is length at age, $l^{sel}$ is the length at 50% selectivity,
and $\delta$ is the difference between the length at 50% selectivity and
the length at 95% selectivity.

*z* is total mortality given by

$$
 z_{t,p,s,a} = F_{t,p,s,a} + m_{s,a}
$$

Where *m* is natural mortality.

This means that each species experiences a total mortality at age in a
time step in a patch, individual fractions of which are portioned off as
catches and subsequently revenues for each fleet. Given revenues, we
then calculate profits $\phi$ per

$$
\phi_{t,p,f} = R_{t,p,f} - c_f(E_{t,p,f} + 1)_{t,p,f}
^{1.3}
$$

where *c* is cost per unit effort and, the 1.3 exponent is in the manner
of @costello2016a . The addition of one to the total effort is to
constrain the value to be greater than 1, preventing undesired behavior
where fishing effort actually gets less expensive if total effort is
less than 1.

The important feature here then is that the revenues and profits for a
given fleet are a function of the prices and catchability of individual
species, but that mortality itself is independent of price. So, a fleet
that has selectivity and catchability greater than zero can exert
fishing mortality on a species even if it obtains no value from that
species.

Each fishing fleet then decides how to allocate its effort in space
based on one of four possible spatial allocation strategies. The ideal
free distribution (IFD) is the standard method for distributing fishing
fleets in spaces (see @gillis2003 and references therein). However, the
IFD presents a number of complications for our study. First, in order to
simplify the assessment of the impacts of different fishing mortality
rates we simulate scenarios with a fixed total pool of effort. As MPAs
change the spatial distribution of profits though, the total amount of
effort required to satisfy the IFD may not equal the total amount of
effort required by the model. Second, even if we were to implement the
IFD for a single fleet, the model is capable of tracking multiple fleets
at once. This means that the IFD for a single fleet would be a
game-theoretic outcome of the actions of each of the individual fleet
conditional on the actions of all other fleets. While possible to
implement this would slow down our model runs to the point of
practically preventing large-scale evaluation of the effects of MPAs on
bycatch.

As such we explored a series of "next best" fleet distribution
algorithms. While not the IFD in any individual time step, over time
they start to approximate the IFD, as each assumes that fleets base
their decisions for the current time step on the outcomes in the prior
timestep, meaning that the impacts of the actions of other fishing
fleets are somewhat internalized. The four possible fleet distribution
algorithms are

1.  Revenue per unit effort (RPUE): The fleet distributes itself in
    space based on revenue per unit effort
2.  Revenue: The fleet distributes itself in space based on revenue
3.  Profit per unit effort (PPUE): The fleet distributes itself in space
    based on profit per unit effort
4.  Profit: The fleet distributes itself in space based on profit

Revenue based spatial distribution is not likely to be very realistic .
However, due to the complexity of parameterizing cost functions fleet
dynamics are often evaluated based on yield or revenue alone, and so we
include those scenarios here to evaluate the potential implications of
this choice. The decision on whether to allocate the fleet based on
absolute or relative (per unit effort) is more complex. When effort
represents the actions of separate and individual fishing actors (e.g.
independent fishing vessels), a per-unit-effort strategy may be more
realistic, in which fishers distribute themselves based on the expected
catch of their individual efforts [@hilborn1987]. Conversely, a system
defined by a sole owner seeking to maximize total profits might be
better represented by a fleet model based on total profits

#### Population Model

The underlying population model used is an age structured single-species
model in the manner of @ovando2021a.

For the population model, numbers *n* at time *t* for age *a* are given
by

```{=tex}
\begin{equation}
n_{t,p,s,a}=\begin{cases}
      = BH(ssb_{t-1,p,s,a}) & \text{if $a = 1$}\\
     = n_{t-1,p,s,a-1}e^{-(z_{t-1,p,s,a-1}}, & \text{if $1< a < max(age)$}\\
     =  n_{t-1,p,s,a}e^{-(z_{t-1,p,s,a})} + n_{t-1,a-1}e^{-(z_{t-1,p,s,a-1})}, & \text{if $a = max(a)$}
  \end{cases}
  (\#eq:pop)
\end{equation}
```
where *BH* is the Beverton-Holt recruitment function, *ssb* is
spawning-stock-biomass.

where $l_a$ is the mean length at at age, $l_sel$ is the length at which
on average 50% of individuals are selected by the fishery, and
$\delta_{sel}$ are the additional units of length at which on average
95% of fish are selected by the fishery.

*ssb* is calculated by converting age to mean length, calculating weight
at age, maturity at age, and then calculating spawning stock biomass as
the sum of spawning potential at age in a given time step.

```{=tex}
\begin{equation}
  l_{a,s} = l\infty_s\left(1 - e^{-k_s(a - a0_{s})}\right)
  (\#eq:len)
\end{equation}
```
Biomass *b* at age is then given by

```{=tex}
\begin{equation}
  b_{a,s} = wa_s \times l_{a,s}^{wb_s}
  (\#eq:weight)
\end{equation}
```
and maturity *mat* is calculated as

```{=tex}
\begin{equation}
  mat_{a,s} = \frac{1}{(1 + e^{-log(19)\times\frac{l_{a,s} - lmat_s}{{\delta}mat_s}}}
  (\#eq:mat)
\end{equation}
```
where $l_{mat}$ is the length at which on average 50% of individuals are
sexual maturity, and $\delta_{mat}$ is the units of length beyond
$l_{mat}$ at which on average 95% of fish are sexually mature.

Spawning stock biomass at time *t* is then calculated as

```{=tex}
\begin{equation}
  ssb_{t,p,s} = \sum_{a=1}^Aw_{a,s,t}^{\gamma_s} mat_{t,a} n_{t,p,s,a}
    (\#eq:ssb)
\end{equation}
```
##### Recruitment {.unnumbered}

Recruitment follows Beverton-Holt dynamics. We do however allow for
three variants in the timing of density dependence:

1.  Global density dependence: Density dependence is a function of the
    sum of spawning biomass across all patches, and recruits are then
    distributed according to habitat quality

```{=tex}
\begin{equation}
n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p}}\times{h}\times\sum_{p=1}^P{ssb_{t-1,p}}}{0.2\times{\sum_{p=1}^Pssb0_p}\times(1 - h)+(h - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p}}}\right)\times {hab_p}
  (\#eq:dd2)
\end{equation}
```
where *hab* is a vector of recruitment habitat quality by patch degined
by

$$hab_p = r0_p / \sum_p^P{r0_p}$$ 2. Local density dependence: Density
dependence occurs independently in each patch and recruits are retained
in their home patch.

```{=tex}
\begin{equation}
  n_{t,p,s,a = 1,p} = \left(\frac{0.8{\times}r0_{s,p}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p}}}\right)
  (\#eq:dd1)
\end{equation}
```
3.  Local density dependence then disperse: Density dependence occurs
    independently in each patch and recruits are then dispersed.

```{=tex}
\begin{equation}
  n_{t,p,s,a = 1,p} = \left(\frac{0.8{\times}r0_{s,p}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p}}}\right)\times \boldsymbol{d^l} \times \epsilon_t
  (\#eq:dd3)
\end{equation}
```
where **d^l^** is the recruitment movement matrix

4.  Post-dispersal density dependence: Larvae are distributed throughout
    the system, and then density dependence occurs based on the density
    of adult biomass at the destination patch.

```{=tex}
\begin{equation}
  larv_{t,p} = ssb_{t-1} \times\boldsymbol{d^l}
  (\#eq:larvmove)
\end{equation}
```
```{=tex}
\begin{equation}
n_{t,a = 1,p} = \left(\frac{0.8{\times}r0_{p}\times{h}\times{larv_{t,p}}}{0.2\times{ssb0_p}\times(1 - h)+(h - 0.2)\times{larv_{t,p}}}\right) \times \epsilon_t
  (\#eq:dd4)
\end{equation}
```
##### Dispersal {.unnumbered}

Dispersal in the model is broken into two components: adults and
recruits Both assume a Gaussian dispersal kernel of the form

```{=tex}
\begin{equation}
m_{s,p_i,p_j}=\frac{1}{\sqrt{2\pi\sigma_{s,p_i}^2}}e^{-\frac{d_{p_i,p_j}^2}{2\sigma_{s,p_i}^2}}
  (\#eq:move)
\end{equation}
```
where *i* is the source patch, *j* is the destination patch, *d* is the
distance between patches *i* and *j* , and $\sigma_s*$is the movement
rate, in units of patches.

Along with distance, net movement is defined by habitat quality. Habitat
quality defines the quailty for adults or larvae of a given patch. The
net probability of movement from patch *i* to patch *j* is defined then
by the product of distance and habitat quality, normalized to sum to one
from each patch, such that all else being equal organisms will move the
most to the closest patch with the best habitat. We allow for seasonal
movement in the model, where years are broken into two seasons. Movement
in season two is defined as the opposite of movement in season one,
intended to reflect the behavior of seasonal migrations.

## Rule-of-thumb MPA Design Strategies

MPA design is governed by two general parameters: the number of cells to
place inside an MPAs, and the design strategy used to place the MPAs.

We consider `r n_distinct(experiment_obj$placement_strategy)` different
MPA placement strategies.

1.  Depletion: MPA locations are prioritized in proportion to the
    spawning stock biomass of the most depleted species pre-MPA. So,
    patches with high spawning biomass of highly depleted species are
    prioritized over patches with high spawning biomass of lightly
    depleted species.

2.  Rate: MPAs are placed in based on the pre-MPA depletion-weighted
    catch relative to the total catch in a patch. So, patches with high
    rates of catch of depleted species relative to total catch are
    prioritized.

3.  Avoid_fishing: MPAs locations are prioritized inversely to fishery
    catches. So, patches with low catches are prioritized over patches
    with high catches.

4.  Target_fishing: MPAs locations are prioritized proportionally to
    fishery catches. So, patches with high total catches are prioritized
    over patches with low catches.

5.  Area: For the "area" strategy, we simply assign patches in
    sequential order until the desired total MPA size is reached.

<!-- The information used to inform each MPA design strategy (except area) -->

<!-- can include a multiplicative observation error term $\sigma_{obs}$ where -->

<!-- $$information \sim lnorm(data,\sigma_{obs}) $$ -->

<!-- For example, for the depletion scenario, $\sigma_{obs}$ controls the -->

<!-- accuracy of the spatial depletion data used to prioritize MPA placement. -->

## MPA Optimization

Our MPA optimization algorithm seems to create an MPA network that
maximizes an objective function, given by

$$obj_p = \Phi \times scale(\sum_f^F \boldsymbol{\phi_{f,p}}) + (1 - \Phi)  scale(\boldsymbol{\sum_s^S\frac{ssb_s}{ssb0_{s,p}}})$$

Where *scale* rescales the values to be between 0 and 1 such that the
weight parameter is comparable between vectors. The algorithm works by
first assigning every patch in the system to an MPA. For the first
round, the algorithm then calculates the equilibrium objective function
value resulting from sequentially removing each individual patch from
the MPA network. The cell whose removal from the MPA produces the
highest objective function value is then removed from the MPA network,
and the marginal value of the objective function each of the remaining
cells from the last steps are then stored. For the next steps, we
sampled 25% of the remaining MPA cells with probability proportional to
their marginal contribution to the objective function in the prior
iteration, selecting the cell whose removal produces the highest total
objective function as the cell to remove from the network, and updateing
the marginal value of the sampled-but-not-selected cells. This in
essence is a sampling-importance-resampling algorithm. The algorithm
continues until no more cells are left in the MPA. This then provides
for a given objective weighting value $\Phi$ the near-optimal MPA
network at every MPA size from 0 to 100%. We say "near" as they is some
path dependency here, in that for example the network that maximizes the
objective function with 40% of the MPAs does not necessarily contain the
optimal network at 30%. However, the reversed nature of the algorithm,
starting at 100% MPA coverage and working better, produces a more stable
network with regards to this problem than starting at 0% and growing to
100%.

## Generalized Experiments

Given the complexity of the operating model used here, a complete
sensitivity analysis across all parameters is not practical. Instead, we
focus on the role of life history, fishing pressure, and MPA design.

Each iteration uses one of three species: A stylized tuna modeled after
skipjack tuna (), a stylized marlin modeled after (), and a stylized
shark modeled after (). Each species has all life history parameters
pulled from FishLife (@thorson2020), with the exception of steepness, set
to XX, and fecundity. For the tuna and marlin species, fecundity is
assumed to scale with weight modified by the hyperallometric exponent.
For the shark species, we assume a constant value of six pups per year
at all ages greater than the age at maturity. We assume that tunas have
an unfished biomass of XX, marlins XX, and sharks XX, to reflect a
large, medium, and small population.

For the generalized experiments we only include one fishing fleet. We
asign a price per KG of \$XX for tuna, intended to reflect a low value
high volume fishery, \$XX for marlin for a higher value low volume
fishery, and \$0 for sharks to reflect at true bycatch species. We
assign a selectivity of 60% of length at maturity for skipjack to allow
for some overfishing when F/M is high enough, 70% for mariln, and 90%
for shark.

We ran `r nrow(state_experiments)` state experiments, varing the state
of nature experienced by a simulation. The candidate states include

| State                 | Distribution          | Description                                     |
|-----------------------|-----------------------|-------------------------------------------------|
| $\sigma_{centroid}$   |                       | Standard deviation of habitat centroid          |
| $\sigma_{habitat}$    |                       | Standard deviation around centroid              |
| Spatial catchability  |                       | Catchability varies in sapce                    |
| Spatial allocation    |                       | Fleet allocation strategy                       |
| F/M                   |                       | Fishing mortality relative to natural mortality |
| Adult movement rate   |                       |                                                 |
| Recruit movement rate |                       |                                                 |
| Hyperallometry        |                       | Presence of hyperallometric fec                 |
| DD timing             |                       | Timing of density dependence                    |
| Seasonal Movement     | $\subset[TRUE,FALSE]$ | Seasonal Movement Present                       |

For each of these `r nrow(state_experiments)` state experiments, we then
apply `r nrow(placement_experiments)` placement experiments defined by
an MPA size, an MPA placement strategy, and ...XX

We simulate degrees of species overlap using two parameters, $\sigma_c$
and $\sigma_h$. All species start with the assumption that the centroid
of their preferred habitat occurs in the center-most patch of teh two
dimensional system simulated by marlin. For our simulations, we then
generated a random deviation from that default centroid for each species
using $\sigma_c$ in the manner

$$centroid_s \sim normal(center,\sigma_c)$$ where *s* denotes species,
and *center* is the coordinate of the central cell. In this way, the
larger $\sigma_c$ grows, the more variation there is betwen the centroid
of the preferred habitat for each species.

We then control the spread of a creatures preferred habitat around that
centroid per a Gaussian dispersal kernal in the manner

$$habitat_{p,s} = \frac{1}{\sqrt{2\pi\sigma_h^2}}e^{-\frac{d_p^2}{2\sigma_h^2}}$$

Where $d$ is the euclidian distance from the $centroid_s$ to patch $p$
and $\sigma_h$ controls the dispersal of the habitat around that range.

## WCPO Case Study Application

The purpose of the case study is to assess the value of optimization.
Given the time constraints of optimization, rather than iterating over
the `r nrow(state_experiments)`, we fix the values from that table to
values that best reflect the state of
`r n_distinct(experiment_obj$critter)` species reported as caught in the
WCPO.

Sigh. lots of details to add in here. This is the worst.

```{r, echo = FALSE}
knitr::kable(sci_to_com %>%
               rename("Scientific Name" = "critter", "Common Name" = common_name))
```

## Assessing Performance

Each model run started at equilibrium conditions and was then run for
`r years` years (`r seasons * years` seasons. We assessed performance by
comparing relevant metrics (e.g. depletion and profits) under a given
scenario with MPAs relative to the values of those same metrics in the
simulated counterfactual world identical in every way save for the
presence of MPAs.

Degree of habitat difference was calculated as

xx will work out the notation for this xx

$$
 Habitat Difference = sd(sqrt(hab))
$$

In order to weight habitat difference by cells where animals occur, not
where they are not.

# Results

Our simulation experiments allow us to examine the wide range of effects
that MPAs can have on groups of target and bycatch species depending on
combinations of ecological, economic, and design parameters. The goal of
this exercise is not to assign any probability to the potential impacts
of MPAs; we have not provided any explicit probabilities to the
alternative states modeled here, and so it may be that a state that was
only modeled on a few occasions in our simulations is in fact highly
common in the real world.

What these experiments are designed to do is illustrate the range of
outcomes that are possible from MPAs under the types of conditions
evaluated here, and evaluate which drivers are potentially the most
influential in affecting particular outcomes with regards to bycatch.
Our case study of a WCPO-esque system is intended then to reduce the
parameter space enough that we can examine the relative tradeoffs and
value of optimization for economic and conservation objectives in a
pelagic system.

One clear caveat across all of these results is that we have made two
very strong assumptions around the behavior of the fishing fleet. The
first is that we assume that effort is displaced, not removed, by the
creation of MPAs. This is a widely debated question around MPAs, and the
real impacts of MPAs on fishing effort will clearly be setting
dependent. For example, large-scale pelagic fishing fleets already have
high rates of vessel movement built into their economic models, and so
may be more likely to move rather than exit the fishery in response to
MPAs. In contrast, smaller scale coastal fishing operations may simply
exit the fishery if their local fishing grounds are placed in MPAs.
There is clear evidence of effort displacement in many fisheries though
(citation), and moreover a motivation for this study was to evaluate the
performance of MPAs in reducing bycatch given the potential for effort
displacement and technical interactions. If MPAs do not produce
increases in fishing effort outside their borders, then the direct
conservation effects of MPAs have to be positive for all species
concerned.

The second is that we do not account for evolution of total effort. Our
assumption is that while fleets can make economic decisions based around
where to allocate their effort in space, the total pool of effort they
exert is fixed (we would note that the underling `marlin` model does
have capacity for more realistic fleet models such as open-access
dynamics). This is clearly an incorrect assumption over the long term,
but allows us to focus on the impacts of the other drivers considered
here. A more through consideration of the impacts of fleet model
structure on MPA outcomes is generally needed, and is critical for any
tactical efforts at MPA modeling.

<!-- what we found was a series of neccesary but not sufficient conditions  -->

<!-- make the tautology clear: illustrating a point, not proving a thing since clearly depends on if you assume displacement, and then degree of non-overlap etc.  -->

## Generalized Experiments

We focused first on the range of impacts that MPAs can have on mature
biomass of target and bycatch species across the range of variables
explored. To illustrate, we pulled out two example simulations, one
where MPAs provided relatively large benefits across all three species,
and a second where MPAs actually produced substantial losses in some
species. In the "best case" scenario, each of the three species shares a
common habitat distribution, and so protection of the central habitat of
one species results in protection for all species. In the "worst case"
scenario, an MPA designed primarily around the tuna fishery displaces
effort onto the marlin and shark habitats, increasing the tuna
population while driving the other two species down
(Fig.\@ref(fig:fig-info)).

```{r fig-info, fig.cap="Illustrative example demonstrating how degrees of overlap can influence "}

experiment_summary <- total_experiment_results %>%
  filter(prop_mpa < 0.5,
         !placement_strategy %in% c("area", "avoid_fishing"),
         placement_strategy == "rate",
         placement_error == min(placement_error)) %>%
  group_by(placement_id, state_id) %>%
  summarise(
    max_loss = max(signif_loss),
    min_loss = max(n_loss),
    total_gain = sum(abs_delta_biodiv)
  ) %>%
  ungroup()

worst_case <- experiment_summary %>%
  # filter(max_loss == max(max_loss)) %>%
  arrange(total_gain) %>%
  # filter(total_gain == min(total_gain)) %>%
  slice(2)

worst_state <- state_experiments %>%
  filter(state_id == worst_case$state_id)

worst_result <- compact_experiment_results %>%
  filter(placement_id == worst_case$placement_id,
         state_id == worst_case$state_id,
         prop_mpa <0.5)

worst_result_plot <- worst_result %>%
  ggplot() +
  geom_text(data = tibble(
    x = .1,
    y = max(worst_result$biodiv - worst_result$bau_biodiv)
  ),
  aes(x = x, y = y, label = "Worst Case"),
  size = 4) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_line(aes(prop_mpa, biodiv - bau_biodiv, color = critter),
            show.legend = FALSE, size = 1.25) +
  scale_x_continuous(labels = scales::percent, name = "MPA Size") +
  scale_y_continuous(
    name = ''
  )

worst_habitat <- worst_state %>%
  select(state_id, data) %>%
  unnest(cols = data) %>%
  select(scientific_name, contains("sigma_"), habitat) %>%
  unnest(cols = habitat) %>%
  mutate(
    scientific_name = fct_recode(
      scientific_name,
      "Shark" = "carcharhinus longimanus",
      "Tuna" = "katsuwonus pelamis",
      "Marlin" = "kajikia audax"
    )
  )

worst_habitat_plot <- worst_habitat %>%
  ggplot(aes(x, y, z = (habitat), color = scientific_name)) +
  geom_contour(show.legend = FALSE) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_blank(),
    panel.border = element_blank()
  )

best_case <- experiment_summary %>%
  filter(min_loss == min(min_loss, na.rm = TRUE)) %>%
  filter(total_gain == max(total_gain, na.rm = TRUE)) %>%
  slice(1)

best_state <- state_experiments %>%
  filter(state_id == best_case$state_id)

best_result <- compact_experiment_results %>%
  filter(placement_id == best_case$placement_id,
         state_id == best_case$state_id,
         prop_mpa < 0.5)

best_result_plot <- best_result %>%
  ggplot() +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_line(aes(prop_mpa, biodiv - bau_biodiv, color = critter), size = 1.25) +
  geom_text(data = tibble(
    x = .1,
    y = max(best_result$biodiv - best_result$bau_biodiv)
  ),
  aes(x = x, y = y, label = "Best Case"),
  size = 4) +
  scale_x_continuous(labels = scales::percent, name = "") +
  scale_y_continuous("Change in SSB/SSB0") +
  theme(legend.position = "top") +
  scale_color_discrete(name = element_blank()) +
  labs(title = "A) ")

best_habitat <- best_state %>%
  select(state_id, data) %>%
  unnest(cols = data) %>%
  select(scientific_name, contains("sigma_"), habitat) %>%
  unnest(cols = habitat) %>%
  mutate(
    scientific_name = fct_recode(
      scientific_name,
      "Shark" = "carcharhinus longimanus",
      "Tuna" = "katsuwonus pelamis",
      "Marlin" = "kajikia audax"
    )
  )

best_habitat_plot <- best_habitat %>%
  ggplot(aes(x, y, z = (habitat), color = scientific_name)) +
  geom_contour(show.legend = FALSE) +
  labs(title=  "B)", subtitle = "Habitat Distribution") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_blank(),
    panel.border = element_blank()
  ) # +
  # facet_wrap(~scientific_name)

fig_info <- (best_result_plot / worst_result_plot) |
  (best_habitat_plot / worst_habitat_plot)

fig_info

```

```{r}

critter_results <- compact_experiment_results %>%
  filter(bau_biodiv > 0, prop_mpa > 0) %>% 
  group_by(state_id, placement_id) %>%
  mutate(real_sigma_centroid = sd(centroid)) %>%
  mutate(delta_biodiv = pmin(1,biodiv - bau_biodiv),
         loss = biodiv < bau_biodiv,
         real_loss = biodiv < (0.9 * bau_biodiv),
         rec_form = as.factor(rec_form)) %>% 
  mutate(critter = fct_relevel(critter,c("Tuna", "Marlin","Shark")),
         rec_form = fct_recode(rec_form, "Global" = "0", "Local" = "1", "Local Then Disperse" = "2", "Disperse Then Local" = "3"))

```

Looking at the full
`r n_distinct(total_experiment_results$placement_id) * n_distinct(total_experiment_results$state_id)`
experiments run, the results were highly variable. Across our
simulations, gains were much more common than losses, with losses of any
magnitude occurring in `r scales::percent(mean(critter_results$loss))`
of simulations with MPAs, but losses of 10% or greater of SSB in only
`r scales::percent(mean(critter_results$real_loss))` of simulations.
However we stress again that we can assign no probabilities to any of
these outcomes in the real world; this simply tells us that under the
range of conditions, selected to represent plausible states of nature,
there are many more ways for MPAs to provide a conservation effect
greater than or equal to 0 than there are for them to provide a net loss
for some species.

The question of interest then are which of the considered drivers appear
to be the most influential in affecting the magnitude and direction of
conservation effects of MPAs. Fishing mortality is the most obvious
answer; if fishing mortality is zero in the absence of an MPA then
clearly a no-take MPA will have no effect. As such we do not focus on
that metric here, focusing instead on other key drivers identified by
our simulation. Four visually prominent factors were MPA size, species,
design strategy, and degree of habitat difference. The degree to which
MPAs of a given size impact spawning biomass depends on a host of
factors. Any potential for conservation losses are intuitively convex in
MPA size: effects of any kind are low when MPA size is low, and
conservation losses must be zero when MPA size is 100% (which this graph
does not quite reach), with peak losses in our simulations occurring
between MPA sizes of XX and XX. Life history plays a clear role our
stylized with shark and marlin populations being much more likely to
experience net conservation losses as a function of MPA placement than
the more resilient tunas. High degree of habitat differences can lead to
both to the potential for large conservation gains or losses depending
on whether a more isolated species is located primarily inside or
outside of MPAs (Fig.\@ref(fig:fig-exp-biodiv)).

```{r fig-exp-biodiv, fig.cap = "Trajectories of changes in SSB/SSB0 relative to BAU for each species across rule-of-thumb MPA design scenarios. Color indicates degree of habitat difference."}

fig_exp_biodiv <- compact_experiment_results %>%
  filter(prop_mpa < 1) %>%
  # filter(f_v_m > 1.5) %>%
  group_by(state_id, placement_id) %>%
  ggplot(aes(
    prop_mpa,
    pmin(1,biodiv - bau_biodiv),
    color =  sigma_habitat,
    group = interaction(placement_id, state_id)
  )) +
  geom_line(alpha = 0.5) +
  geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
  facet_grid(critter ~ placement_strategy, labeller = labeller(placement_strategy = capitalize)) +
  scale_color_viridis_c(name = "Habitat Difference", guide = guide_colorbar(frame.colour = "black",ticks.colour = "black", barwidth = unit ("15", "lines"))) +
  scale_y_continuous("Change in SSB/SSB0 relative to BAU") +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1), name = "MPA Size") +
  theme(legend.position = "top")

fig_exp_biodiv
```

We can somewhat quantitatively evaluate the relative impact of these
drivers using a regression, given that by construction variation in each
of evaluated drivers are independent in our simulations. We fit a
binomial regression to the presence or absence of of a net conservation
loss per species as a function of the simulated experimental variables.
The factors that according to this model most increase the probability
of a net conservation loss occurring include habitat differences, a
placement strategy of avoiding fishing areas, and the life history of
the species (shark or marlin relative to tuna). The factors that most
decrease the probability of a conservation loss from an MPA are MPA
size, placement strategy based on local depletion, and somewhat
interestingly fishing mortality rate
(Fig.\@ref(fig:fig-loss-drivers)-A).

However, this result is somewhat misleading (xx so should we just
remove?xx). Fig.\@ref(fig:fig-exp-biodiv) shows the clearly bimodal
nature of for example habitat differences. This analysis helps inform us
that habitat differences are a necessary but not sufficient condition
for conservation losses induced by effort displacement. The regression
based results require assuming that each of the effects are independent
and linear in nature. In fact each of these factors can interact in
complex and non-linear ways that may not be apparent visually or through
this type of regression analysis. To address this, we ran a random
forest using the `ranger` package (citation) in R (citation) and
extracted the impurity corrected importance scores for each of the
variables. While this does not tell us the expected marginal effect of
individual drivers, it does tell us which variables are important if our
goal is to accurately predict the potential for unintended consequences
or magnitude of conservation gains in MPAs. Note here that variables
that intuitively should have an impact but have no visually or
statistically (in a linear model) clear impact on the probability of
experiencing a conservation loss such as the dispersal distances of
adults or larvae have importance scores comparable in magnitude to
habitat differences when considered in the context of nonlinear effects
and interactions with other variables
(Fig.\@ref(fig:fig-loss-drivers)-B).

```{r fig-loss-drivers, fig.cap = "Things"}

loss_reg <-   glm(
  loss ~   f_v_m  + critter + seasonal_movement + adult_movement_sigma + rec_form + recruit_movement_sigma + prop_mpa + sigma_habitat + spatial_allocation + placement_strategy + hyper,
  family = binomial(),
  data = critter_results %>% filter(between(prop_mpa,.2,.4))
)

loss_reg_coefs <- broom::tidy(loss_reg) %>%
  mutate(lower = estimate - 1.96 * std.error,
         upper = estimate + 1.96 * std.error) %>%
  mutate(
    term = str_replace(term, "placement_strategy", "Placement Strategy: "),
    term = str_replace(term, "rec_form", "DD Timing: "),
    term = str_replace(term, "spatial_allocation", "Fleet Allocation: "),
    term = str_replace(term, "hyper", "Hyperallometry present"),
    term = str_replace(term, "prop_mpa", "MPA size"),
    term = str_replace(term, "sigma_habitat", "Habitat differences"),
     term = str_remove_all(term, "critter")
  )

loss_coef_plot <- loss_reg_coefs %>%
  ggplot() +
  geom_hline(aes(yintercept = 0)) +
  geom_pointrange(aes(
    forcats::fct_reorder(term, estimate),
    estimate,
    ymin = lower,
    ymax = upper
  )) +
  scale_y_continuous(name = "Log-odds Ratio Effect") +
  scale_x_discrete(name = '') +
  coord_flip()

tmp_critter <-
  critter_results %>% filter(between(prop_mpa, .2, .4)) %>%
  select(-habitat) %>% 
  mutate(fctr_loss = factor(loss))

loss_rf <-   ranger::ranger(
  loss ~   f_v_m  + critter + seasonal_movement + adult_movement_sigma + rec_form + recruit_movement_sigma + prop_mpa + sigma_habitat + spatial_allocation + placement_strategy + hyper,
  data = tmp_critter,
  importance = "impurity"
)

loss_importance <- data.frame(importance = loss_rf$variable.importance %>% sort()) %>% 
  rownames_to_column(var = "term")

loss_importance_plot <- loss_importance %>% 
  ggplot(aes(forcats::fct_reorder(term, importance), importance)) + 
  geom_col() +
  scale_y_continuous(name = "Variable Importance") +
  scale_x_discrete(name = '') +
  coord_flip()

fig_loss_drivers <- loss_coef_plot + loss_importance_plot

fig_loss_drivers
```

```{r fig-delta-drivers, eval = FALSE}

delta_reg <-   rstanarm::stan_glm(
  delta_biodiv ~   f_v_m  + critter + seasonal_movement + adult_movement_sigma + rec_form + recruit_movement_sigma + prop_mpa + sigma_habitat + spatial_allocation + hyper,
  data = critter_results %>% filter(between(prop_mpa,.2,.4)),
  cores = 4,
  chains = 4
)

tmp_critter <-
  critter_results %>% filter(between(prop_mpa, 0.2, 0.4)) %>%
  select(-habitat) %>% 
  mutate(fctr_loss = factor(loss))
# 
# tmp_rcp <- recipes::recipe(loss ~ ., data = tmp_critter) %>% 
#   themis::step_upsample(fctr_loss, over_ratio = 0.75) %>% 
#   recipes::prep(training = tmp_critter, retain = TRUE) %>% 
#   recipes::juice()

delta_rf <-   ranger::ranger(
  delta_biodiv ~   f_v_m  + critter + seasonal_movement + adult_movement_sigma + rec_form + recruit_movement_sigma + prop_mpa + sigma_habitat + spatial_allocation + placement_strategy + hyper,
  data = tmp_critter,
  importance = "impurity"
)

delta_rf$variable.importance %>% sort()

plot(delta_reg)

```

## WCPO Case Study

<!-- The generalized experiments help us understand the potential impacts of -->

<!-- the evaluated drivers, but also highlight the challenge in making -->

<!-- definitive statements about the effect of a given driver on the conservation impacts of MPAs, given the complex ways in which all of these factors -->

<!-- interact, even within a relatively simple model such as this one. In -->

<!-- order to help illustrate how addressing the kinds of spatial dynamics -->

<!-- and technical interactions included in our model can affect and benefit -->

<!-- marine spatial planning, we then ran a case study intended to loosely -->

<!-- reflect the dynamics of WCPO tuna, billfish, and shark fisheries and -->

<!-- populations. -->

Our case study allows us to focus in on important questions around the
relative performance of "rule of thumb" MPA design strategies relative
to optimized networks. In particular, as any leakage induced losses from
MPAs must be 0 if MPA size is 100%, an important question then is within
this case study what tradeoffs are present between increasing the
probability of a successful conservation outcome by increasing MPA size
and the increasing probability of negative economic impacts with MPA
size (xx phrase better xx).

Given the set of parameters simulated here, we find little intrinsic
risk for large-scale unintended losses in total spawning biomass caused
by MPA effort displacement for the species included in our WCPO case
study. Each of the design scenarios produced at least some small loss in
biomass for one or more species at some MPA size. However, these losses
were generally small and isolated to a subset of species within a small
MPA size range, with the exception of the "Avoid_fishing" MPA placement
strategy that resulted in net losses for most species at large MPA sizes
(though simply minimal impact at smaller sizes). All design strategies
except "Avoid_fishing" and "Rate" resulted in relatively rapid declines
in profits for the three main tuna fisheries (skipjack, yellowfin, and
bigeye). Both "Avoid_fishing" and "Rate" maintained profits close to
business as usual up to a large MPA size, but did so at the expense of
conservation benefits (Fig.\@ref(fig:fig-cs-critters)). Results for
yield rather than profits are qualitatively similar, see SI.

```{r fig-cs-critters, fig.cap = "Simulated trajectories of changes in profits and spawning stock biomass (SSB) for selected species caught in the WCPO. Each column represents a different MPA design strategy. 'Optimized' refers to an MPA network optimized to maximize total SSB across all species."}

opt_comparison <- opt_obj %>%
  filter(alpha == max(alpha)) %>%
  mutate(placement_error = 0,
         prop_mpa = p_protected / 100,
         placement_strategy = "Optimized") %>%
  select(critter, placement_error, delta_biodiv, delta_econ, delta_yield, prop_mpa,placement_strategy)

 experiment_obj %>% 
   select(colnames(opt_comparison)) %>% 
  bind_rows(opt_comparison) %>% 
   mutate(placement_strategy = fct_relevel(placement_strategy,"Optimized")) %>% 
   left_join(sci_to_com, by = "critter") %>% 
  filter(placement_error == min(placement_error)) %>% 
   rename("Change in SSB" = delta_biodiv, "Change in Profits" = delta_econ) %>% 
  pivot_longer(c("Change in SSB", "Change in Profits"), names_to = "metric", values_to = "value") %>% 
  ggplot(aes(prop_mpa, value, color = common_name)) + 
   scale_color_discrete(name = '') +
   geom_hline(aes(yintercept = 0), linetype = 2)+
  geom_line(show.legend = TRUE) + 
   scale_y_continuous(name = '') +
   scale_x_continuous(name = "MPA Size",labels = scales::percent, guide = guide_axis(n.dodge = 2)) +
  facet_grid(metric~placement_strategy, scales = "free_y", switch = "y", labeller = labeller(placement_strategy = capitalize)) + 
   theme(strip.text.x = element_text(size = 8), 
         legend.position = "top", 
         legend.text = element_text(size = 8))

```

Viewed as totals, we found a small parameter space under the "Area" and
"Avoid_fishing" design strategies where MPAs were capable of causing a
net decrease in total spawning stock biomass across all evaluated
species. However, the efficiency frontier indicates that these points
are not Pareto efficient, and MPA design configurations exist that
achieve the same levels of economic performance while producing a net
gain in spawning biomass. Optimal MPA design was able to slow the
tradeoff between total profits and bycatch spawning biomass, but for any
conservation weighting any increases in total biomass of bycatch species
came at a decrease in total profits. However we also found limited need
for optimized networks, with the "Depletion" design scenario essentially
following the Pareto efficiency frontier (Fig.\@ref(fig:fig-cs-totals)).

```{r fig-cs-totals, fig.cap = "Efficiency frontiers for changes in total profits and total bycatch spawning stock biomass (SSB). Color gradient line shows optimized pareto frontier, with color showing weighting of conservation (total spawning stock biomass) in the objective function. Red points show outcomes of 'rule of thumb' MPA design strategies. "}
 total_opt_obj %>%
  group_by(alpha) %>%
  filter(obj_fun == max(obj_fun)) %>%
  ungroup() %>%
  ggplot() +
  geom_line(aes(total_bycatch_ssb_change, delta_econ, color = alpha), size = 1.5) +
  geom_point(
    data = total_experiment_obj,
    aes(total_bycatch_ssb_change, delta_econ),
    size = 2,
    alpha = 0.1,
    fill = "red",
    shape = 21
  ) +
  geom_vline(aes(xintercept = 0), linetype = 2) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_abline(intercept = 0, slope = -1) +
  facet_wrap(~placement_strategy, scales = "free_y", labeller = labeller(placement_strategy = capitalize)) + 
  scale_color_viridis_c(name = "Conservation Weighting in Optimization",
                        guide = guide_colorbar(frame.colour = "black", ticks.colour = "black", barwidth = unit("15", "lines"))) +
  scale_x_continuous(name = "Change in Total Bycatch SSB") +
  scale_y_continuous(name = "Change in Total Profits ") +
  theme(axis.text.x = element_text(size = 8),
        legend.position = "top")

```

```{r, eval = FALSE}

 total_opt_obj %>%
  group_by(alpha) %>%
  filter(obj_fun == max(obj_fun)) %>%
  ungroup() %>%
  ggplot() +
  geom_line(aes(delta_biodiv, delta_econ, color = alpha), size = 2) +
  geom_point(
    data = total_experiment_obj,
    aes(delta_biodiv, delta_econ),
    size = 2,
    alpha = 0.1,
    fill = "red",
    shape = 21
  ) +
  geom_vline(aes(xintercept = 0), linetype = 2) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_abline(intercept = 0, slope = -1) +
  facet_wrap( ~ placement_strategy) +
  scale_color_viridis_c(name = "Conservation Weighting in Optimization") +
  scale_x_continuous(name = "Change in Total Biodiversity Relative to BAU ") +
  scale_y_continuous(name = "Change in Total Profits Relative to BAU ") +
  theme(axis.text.x = element_text(size = 8),
        legend.position = "top")

```

# Discussion

MPAs can be a powerful tool for providing conservation benefits across
large swaths of species and space. Under the appropriate circumstances
they can also provide benefits to food security and economic outcomes.
However, while spatial closures are simple in concept, predicting the
likely effects of MPAs across groups of species and fishing fleets
becomes increasingly complicated as the systems being studied deviate
from assumptions of homogenously distributed species with perfectly
selective fleets. In this paper we used a new bio-economic model to
demonstrate how a range of variables often omitted in MPA modeling can
act to influence the outcomes of MPAs on conservation and economic
objectives, particularly with regards to protecting bycatch species. We
demonstrate that unintended conservation losses to both target and
bycatch species are possible, but only under a range of specific
conditions, and under most simulated cases did not occur. MPAs were
capable of causing small losses in biomass for the WCPO case study, but
only for a subset of design strategies and species, most of which could
be overcome by optimized design.

Our results indicate that as suggested by theory (citation) and
empirical evidence (citation), MPAs can indeed have unintended
consequences for conservation if they result in effort displacement
rather than attrition. However, our model found these to be a minority
of the simulations tested. This does not suggest that leakage may not a
problem for specific MPA. But, it also indicates that across the range
of plausible parameter combinations that we explored it was much more
common for MPAs to provide net conservation benefits even if effort is
displaced. This suggests then that we should of course not dismiss the
conservation benefits of MPAs in the presence of effort displacement,
but also that users must be aware of the potential for unintended
consequences that can indeed occur. While many different factors
interact in complex ways to affect the potential for unintended
conservation consequences, the most obvious is the degree of difference
in the habitats utilized by the species within an ecosystem in which an
MPA is to be placed. As a starting point, it is critical that
communities engaged in marine spatial planning create a localized assessment of
the degree of habitat overlap across potentially affected species to
assess the potential risk of unintended consequences.

The MPA literature is highly dominated by modeling studies (citation).
What then is the added value of the model and results presented here?
Our model contains many features missing from many general MPA models,
including size structure, seasonal movement, larval and adult habitats
and dispersal, and critically technical interactions xx clearer term for
this xx across species. A reader might reasonably ask about a range of
other factors not included, such as time-varying parameters and species
interactions. One could of course continue to add features to the model
until it reaches the complexity of an end-to-end model such as Atlantis
(citation). While such models have tremendous value, our goal here is to
present a model that can help provide closer to tactical advice for
users without the data or capacity for such high resolution models. One
justification for simpler MPA models is often computational
tractability: tracking the full age structure of multiple species and
the economic outcomes of multiple fleets dynamically across a high
resolution spatial grid can quickly become computationally expensive.
Faster implementations completely written in languages such as C++
however become inaccessible to many users. Our model seeks to fill a
middle ground here, with the core operations run in C++ while providing
a relatively user-friendly front end in the open-source and commonly
used R programming language. This allows users to ask the sorts of
questions addressed in this paper that require a more realistic model,
without requiring inordinate amounts of data, programming expertise, or
computational time.

Several results from our model may seem counterintuitive based on
existing literature. We found little effect of the timing and spatial
structure of density dependence in our model. Other studies have pointed
out that decisions about whether density dependence is local or global
can have major implications for fishery outcomes of MPAs (citations).
The difference here is that our model is fully age structured. As such,
yields are separated from recruitment, the location of density dependent
processes in our model, by growth, natural mortality, and adult movment.
In contrast biomass dynamics models assume that density dependence
occurs on a net "growth" parameter combining both recruitment and
biomass growth, and that this growth is instantly available to the
fishery. As such we suspect that the timing of density dependence plays
a much bigger role in the expected outcomes of MPAs when biomass
dynamics models are used as opposed to when it is included as an effect
on recruitment in an age structured model.

@marshall2021 pointed out that many fisheries models assume that
fecundity is simply proportional to spawning biomass, when in fact there
is ample empirical evidence that fecundity follows a hyperallometric
pattern, i.e. that fecundity increases faster than weight. In that study
the authors found that the presence of hyperallometric growth
substantially amplified the predicted benefits of spatiao-temporal
closures such as MPAs, relative to an assumption of allometric
fecundity. In contrast, we find that the presence of hyperallometric
fecundity, even at levels greater than the mean values reported in
@marshall2021, had little effect on the expected conservation outcomes
of MPAs (Fig.\@ref(fig:fig-loss-drivers), see SI for additional
analyses). The key distinction here is that @marshall2021 evaluated the
effect of MPAs under different assumptions of management, e.g.
management based on spawning potential ratio values calculated assuming
allmetric fecundity when in fact the relationship was hyperallometric.
Our study holds effort constant, and so reflects more the isolated
effect of hyperallometric on MPA outcomes independent of management
actions.

@rassweiler2012 and @rassweiler2014 both demonstrated the potential
importance of optimized MPA planning over more "rule of thumb" MPA
designs. In contrast, we found little value of optimization, with the
"Depletion" MPA allocation strategy more or less following the Pareto
efficiency frontier (Fig.\@ref(fig:fig-cs-totals)). As pointed out in
@rassweiler2014, the value of optimization (at least if the objective
function depends on both fishery yields and fish biomass) depends on the
degree of overfishing present. In our WCPO case study, best available
evidence is that none of the major tuna stocks that provide the bulk of
yield, revenues, and profits to the WCPO fleets are overfished or
experiencing overfishing (citations, need to update with assessments
since pons is out of date now). Our results then are consistent with
@rassweiler2014, and evaluations of our model on stocks experiencing
overfishing will likely find much higher value of optimization.

Value of information and fleet dynamics are two critical areas for
future study in this realm. Our study assumes perfect information in the
design of MPAs. We feel this is a useful benchmark for isolating the
potential impacts of key ecological and economic drivers of MPA effects.
However it is clearly not a realistic assumption, particularly for
pelagic ecosystems where species inhabit a complex, dynamic, and hard to
observe mosaic of habitats. This is likely to be particularly true for
rarer bycatch species such as sharks that may be the impetus behind many
pelagic MPA designs. Further research is needed to examine how the types
of dynamics explored here are affected by varying degrees of information
quality.

We focused our efforts on modeling the spatial distribution of a fishing
fleet conditional on a fixed pool of fishing effort. As discussed
previously, the outcomes of MPAs depend critically on the reaction of
fishing fleets to new spatial protections. In reality neither perfect
attrition nor perfect redistribution are likely to be realistic
scenarios. Instead individual fishing operations will adjust their
behavior according to their own objective functions. More realistic
models of total effort could include simple open-access dynamics in the
manner of @cabral2019, to more complex dynamics reflecting combinations
of policies such as effort controls, spatial closures, and rights based
fisheries management.

As demonstrated here there though there is a possibly endless list of
refinements and realism that could be added to simulation based studies.
While efforts such as this can help illustrate which drivers may be the
most important in particular circumstances in many cases as demonstrated
here the effect of specific variables are likely to be so context
dependent as to defy easy classification of "important" or "not". We
feel these models are better viewed as a foundation for empirical
action. In terms of planning, models such as this can be used as tool
for answering "what if" rather "what will be". Helping community members
visualize the potential impacts of alternative states of nature, such as
debates around life history or habitat distribution, can help set
expectations and guide data collection. Models such as this can also
serve as hypotheses for expected effects that can ideally be confronted
with data as MPAs are actually implemented.

# Conclusions

The bluntness of MPAs as a tool is both their strength and weakness.
Targeted fisheries management is often focused on the assessment and
management of individual stocks. While this has worked well in many
cases [@hilborn2014a; @hilborn2020; @worm2009], it is an unrealistic and perhaps undesirable 
model to replicate in many parts of the world, where data limitations
and fishery complexity make this sort of highly targeted management
untenable. In contrast, MPAs can provide protection to entire portions
of an ecosystem with minimal need for targeted interventions. As a
result and as shown by @hastings2017, MPAs can provide an effective
solution to bycatch problems under the right circumstances. We extend
the spirit of that analysis here, but expanding the model used to assess
the impact of MPAs on bycatch under broader range of scenarios. As
predicted by @hastings2017, as conditions deviate from the stylized
nature of their model the effects of MPAs on bycatch become more
nuanced. While we similarly find a large parameters space where MPAs
provide clear benefits to the conservation status of both target and
bycatch species, we also identify a smaller subset of enabling
conditions, such as degrees of habitat overlap, that can lead to MPAs
producing net reductions in bycatch biomass. As increasing numbers of
MPA networks are implemented around the world, tools such as this can
help communities both design and evaluate the performance of MPA
networks with the goal of better outcomes for marine ecosystems and the
human communities that depend on them.

# References

::: {#refs}
:::

# SI

```{r fig-cs-yield-critters, fig.cap = "Simulated trajectories of changes in yields and spawning stock biomass (SSB) for selected species caught in the WCPO. Each column represents a different MPA design strategies. Optimized refers to an MPA network optimized to maximize total SSB across all species."}

opt_comparison <- opt_obj %>%
  filter(alpha == max(alpha)) %>%
  mutate(placement_error = 0,
         prop_mpa = p_protected / 100,
         placement_strategy = "Optimized") %>%
  select(critter, placement_error, delta_biodiv, delta_econ, delta_yield, prop_mpa,placement_strategy)

 experiment_obj %>% 
   select(colnames(opt_comparison)) %>% 
  bind_rows(opt_comparison) %>% 
   mutate(placement_strategy = fct_relevel(placement_strategy,"Optimized")) %>% 
   left_join(sci_to_com, by = "critter") %>% 
  filter(placement_error == min(placement_error)) %>% 
   rename("Change in SSB" = delta_biodiv, "Change in Yield" = delta_yield) %>% 
  pivot_longer(c("Change in SSB", "Change in Yield"), names_to = "metric", values_to = "value") %>% 
  ggplot(aes(prop_mpa, value, color = common_name)) + 
   scale_color_discrete(name = '') +
   geom_hline(aes(yintercept = 0), linetype = 2)+
  geom_line(show.legend = TRUE) + 
   scale_y_continuous(name = '') +
   scale_x_continuous(name = "MPA Size",labels = scales::percent, guide = guide_axis(n.dodge = 2)) +
  facet_grid(metric~placement_strategy, scales = "free_y", switch = "y", labeller = labeller(placement_strategy = capitalize)) + 
   theme(strip.text.x = element_text(size = 8), 
         legend.position = "top", 
         legend.text = element_text(size = 8))

```

# Scraps

There two general results beg some important questions. First, are
unintended consequences inevitable given the kinds of technical
interactions explored here, or can optimized MPA networks reduce or
eliminate these unintended consequences. Second, one clear solution to
the problem of unintended consequences is to simply make the MPAs large:
if MPAs cover 100% of the area there can be no direct unintended
conservation consequences. The potential for negative impacts on
economic fishery outcomes increases as MPAs reach large size though.
What then are the tradeoffs between reducing the potential for
unintended conservation consequences and fishery revenues?

To address these questions, we ran a more limited "case study" set of
simulations, based on the dynamics of the WCPO tuna fisheries. Rather
than iterating over different movement dynamics and habitat
configurations, we fix habitat layers based on reported CPUE for the
species in question, and xx other things we fix xx.

The general experiments conducted show two general insights

1.  Even conditional on fishing exploitation and general life history
    strategy, the effects of MPAs on a group of species with technical
    interactions is highly variable and dependent on habitat
    distribution, design decisions, and data quality. xx make this more
    explicit.

2.  Some degree of unintended conservation consequences occured across
    nearly all of our simulations below MPA coverage near 100%, though
    small losses were much more common than large. Different design
    strategies are better suited to avoiding unintended consequences.

Thinking through the "naive" MPA design strategies to compare against
the optimized outcomes. Thoughts at the moment below. These can be run
in a few seconds each, so we cab be be pretty liberal in coming up with
combinations if there are other ideas people want to try, but also if
some of these seem pointless can drop as well! 6:11 Design strategies 1.
Assign protection by SSB / SSB0 weighted abundance in space (protect the
highest concentrations of the most threatened species) 2. Assign
protection by bycatch rates (distribution of biomass bycatch / biomass
target) 3. Assign protection by targeting fishing effort 4. Assign
protection by avoiding fishing effort Knowledge states 1. Degrees of
accuracy over the thing being measured (i.e. error in the design
strategy) 2. Number of species considered (e.g. design aronnd only 2
bycatch species instead of all) 3. über naive: just close off random
patches up to a specified size We then plot these points in objective
function space and compare to the efficiency frontier / consider
frequency and magnitude of conservation and economic tradeoofs The
question all this is answering: how sensitive are conservation and
economic outcomes to design strategy / uncertainty when there are
technical interactions and heterogeneity in habitat? The first stage is
teeing up the idea that beyond conservation / yield tradeoffs, as you
increase habitat heterogeneity with technical interactions you increase
the potential for conservation tradeoffs. But, the degree is going to be
super context dependent, so to dive in a bit we then focus on this case
study of the particular species / habitat overlap in the WCPO. Further
work might reveal more generalities in this, but htis is basically
getting a sense for the potential magnitude of the challenge.

XX stress a key difference: may other papers on this are static, in that
they just some up displacement on different layers, rather than
addressing dynamic respones over tim xx

We show.... - sensitivity of both conservation and fishery outcomes to
pelagic realism and design choices - compared to targeted / dynamic
intervention... - value of information would be great...

Contrast to coastal MPA theoretical work - @hastings2017 - @ovando2021 -
@white2010 - not quite sure what to do about points raised in
@fulton2011a - to what extent has this been done...

The paragraph from @fulton2011a that broke me...

" Spatial management was insufficient as the dominant management lever
because it proves quite difficult to strike the balance between
protecting areas that are of a size sufficient to produce any
conservation benefit and providing sufficient fishing grounds to sustain
an econom- ically viable industry (which is not possible in many
fisheries if closures on the order of 70--80% or more are imposed,
Fulton et al. 2007). More- over, Savina etal. (2008) found that effort
displaced by closures potentially undermines any benefits coming from
the zoning. For instance, strong improvements in diversity (by nearly
20%) under heavy fishing pressure occurred if effort outside a marine
protected area (MPA) was controlled. However, without effort control,
there was a significant (\>20%) drop in overall diversity. Fulton et al.
(2007), Savina et al. (2008), Smith et al. (2010) and unpublished work
associated with Horne et al. (2010) all found that while spatial
management appears to be an effective means of satisfying conservation
objectives con- cerning habitat and restoring age-structures of less
mobile species, benefits from closures for individual species depend on
its trophic position, degree of mobility, geographical distribution and
extent of ontogenetic shifts in habitat use and depth (highly mobile or
prey species do not benefit in the way that more site attached or
predatory species do)."

Lots of studies have shown that lots of things matter, can we provide
guidance on which things are priorities?

Where does this fit in context of @smith2021? One important difference
is that they point out differences in effectiveness, where effectiveness
is still almost always positive, since they just measure leatherbacks.
So their fig 5, there are a few cases where leatherback catches actually
go up, but it's rare. So, extending to multiple species and keeping
track off all of them shows it's even harder

How does this contrast with the the large and growing dynamic ocean
management lit? - By definition, with perfect information dynamic $\geq$
static spatial controls - DOM literature is mostly static, tracks
displacement onto static abundance / risk layers - Doesn't allow for
dynamic impacts, for example fleet adjustments based on updated economic
surface, or short-term costs for long-term benefits (can overstate
costs?) - Dynamic model allows for delta between dynamic policies, e.g.
Fmsy vs DOM - analytical models a la hastings have value but do / don't
stand up to realism

Presentation of the model as a MICE model for MSP - reasonable
parameterizable - flexible - fast

Focus of debate has been on fisheries and yields. As MPAs move to
pelagic systems though, we show that conservation outcomes also depend
on correctly identifying ecological state / design decisions etc.

What are the considerations of protection you need to make good
decisions. It's not always obvious how to design things to get the
outcome you want. If we could make the case that something like closing
the high seas is insufficient conservation management without
conservation management inside EEZs / what is the

Need to do an optimization here, calculate the marginal value of added
cells.

What's this about? Pointing out a problem, which to make impactful would
require making the case that there is a problem with current design
strategies. e.g.. "MPAs are often designed like this (simple dispersive
models with constant habitat, or MARXSAN style things), and doing that
does/does not create problems A tactical evaluation. Picking a handful
of high profile MPAs, parameterizing all the knowns as well as possible,
iterating over unknowns, and see what the range of outcomes are.
Cleaner, more impactful, but also opens it up (depending on the results)
to a details fight about very specific modeling choices How plausible
are losses? Maybe we drill down on this a bit. From a purely bycatch
perspective, the usual questions of fisheries tradeoffs aren't on the
table (but can add to the discussion considerations of how much target
fishing you give up). Given that then, while it's possible that
"targeted" interventions would work better (ignoring costs/enforcement,
etc), there's really no reason not to do MPAs. Unless, you actually end
up concentrating fishing pressure outside on populations. So, we
consider how plausible is it for that to really happen. Issue here is
that seems likely to be pretty hand-wavey. Most likely that the vast
majority of simulations will show positive impacts, but have no way of
assigning probabilities to any one state of nature.

The Role of Blue-Water Marine Protected Areas in Bycatch Reduction

What's the question: How effective are Blue-Water MPAs in reducing
fishing mortality of important pelagic bycatch species?

Another angle. There's going to be a lot of unknowns in designing BWMPA.
How much does not fully understanding a few key life history (e.g.
movement) matter, and how much does the role that the species plays in
the broader fleets portfolio matter

Could simulate habitat from a multivariate normal, where you use the
sigma to control how similar rasters are across species

Marine protected areas are increasingly proposed to protect biodiversity
in pelagic systems. However, maby paradigms of mpa design come from
coastal systems. Here we ask, how likely are blue-water MPAs to provide
conservation benefits to bycatch species, and what are the key
uncertainties underlying these outcomes? We begin with a general
discussion of the ways in which traits of open-ocean systems might
influence the conservation performance of BWMPAs. Given the massive
parameter space, we then constrain ourselves to a styulized version of
the western central pacitic ocean to ask, how feasible is it to provide
broad protection of bycatch species through common MPA design paradigms
in the context of this system. Our results demonstrate the critical
importance of XX and XX in designing MPAs in pelagic systems, and
highlight the potential risks of not accounting for XX. Introductionhere

Bycatch is a significant driver of population declines for many pelagic
species Sharks, seabirds, mammals, etc MPAs increasingly turned to as a
solution for conservation problems, etc. MPA science is largely based on
dynamics of coastal systems. Open-ocean ecosystems have very different
dynamics that require dedicated research Burgess et al. (2018) estimated
the reductions in fishing mortality needed to halt declines for a number
of bycatch species Here we ask, what are the key drivers of bycatch
conservation outcomes from MPAs? There are all of these species; you
reduce fishing mortality. You can either do fisheries management, or you
can do MPAs, let's focus in on the MPA side. This could reduce fishing
mortality, or it could do, and it depends on the dynamics of the species
and the fleets. When can MPAs reduce fishing mortality enough for
bycatch species? And where can you go wrong?

It doesn't make any sense to ask "do MPAs help bycatch": the question is
obviously it depends. To quickly make this case, we'll assume that a
species is in need of recovery. We then do the simple mako example to
show that effects of spatial closure depend on the overlap of habitat
and the assumption of fleet dynamics, and provide some empirical
evidence that in particular effort redistribution is not unlikely.

Determining the effect of an MPA network on bycatch then will also
depend on trying to get close to a tactical model. But, this creates a
problem for blue water systems in particular. Can make the argument that
given the centrality of relatively static habitat features in coastal
MPAs "protect the key habitats" might not be a bad strategy from a
conservation strategy. But in highly mobile pelagic systems there (may
be) a lot more variables in play. Where should researchers direct their
attention?

To reduce the parameter space, we constrain ourselves to a system where
1) we know the species involved 2) we have some sense of the relative
depletion and value of the different species and 3) some basic life
history of the species. We then vary commonly unknown parameters,
particularly preferred habitats, spatial catchability, movement rates
etc, and pair these simulations with different MPA designs, and examine
the conservation outcomes for bycatch species.

What i don't like about this

There isn't an "answer" to this: baring something like "changing the
type of density dependence never had any effect greater than XX%", for
anything that on occasion has an impact the answer of "how much" will
almost always depend on what else is going on.

So, it seems like there are two real angles of attack here

Pointing out a problem, which to make impactful would require making the
case that there is a problem with current design strategies. I.e. "MPAs
are often designed like this (simple dispersive models with constant
habitat, or MARXSAN style things), and doing that does/does not create
problems A tactical evaluation. Picking a handful of high profile MPAs,
parameterizing all the knownns as well as possible, iterating over
unknowns, and see what the range of outcomes are. Cleaner, more
impactful, but also opens it up (depending on the results) to a details
fight about very specific modeling choices

Methods

Collect life history for key bycatch species and target species, cite
Bradley et al. pulling together species distribution and threats etc.
Introduce marlin Parameterize marlin runs with "knowns" from life
history of bycatch and target species Highlight key "unknowns" that we
will iterate over, e.g. fleet models, density dependent movement, timing
of density dependence etc Simulate effects of MPAs given reasonably
realistic situations but iterating over key unknowns Evaluate the range
of outcomes and key drivers of MPA effects I'm thinking we try and avoid
a really specific case study but hard not to accidentally do one without
making everything up....

Results Range of likely changes in fishing-induced mortality for the
species selected Key drivers of uncertainty in those outcomes, focusing
on things that switch states from reduction to increase in mortality How
common are conservation tradeoffs between bycatch species? Fishery
impacts

Discussion What does this say about BWMPAs and bycatch? Not going to get
into design decisions here, depending how common tradeoffs are I think
that will be the main focus Highlight any species that frequently
induced tradeoffs Tee up future research MPAs in context of other
management MPAs as risk reduction strategy Etc.

We do not focus into the individual effects of particular variables as
we found them to be highly dependent on the exact question asked and the
set of data included, and since we cannot ascribe any actual likelihood
to any individual state of nature simulated here.

However, three important features do emerge. The first is that life
history plays an important role in the potential for unintended
conservation consequences from MPAs. All else equal, MPAs were much less
likely to produce losses in spawning biomass for our tuna-like species
than the marlin or shark-like species. Given that within the broad life
history traits we do vary adult and larval movement, fishing mortality,
as well as the timing of density dependence, we believe this result is
mostly due to the high resilience (high steepness, low age at maturity,
fast growth) of the tuna-like species relative to the marlin and
shark-like species.

The second is that design matters...

```{r}

```

The third is that extreme outcomes (both positive and negative),
particularly at lower MPA sizes, are largely driven by the presence of
small range sizes interacting with seasonal migrations. Across species,
simulations without seasonal migrations and small ranges were much more
likely to produce extreme results than those with seasonal migrations
and/or larger ranges. Presence of seasonal migrations acts as a
dilutant: As a given static MPA only provides protection, or displaces
fishing effort, onto a particular area in only half of the year, both
the benefits and costs of MPAs are smaller. When seasonal migrations are
not present and the species range is small, MPAs are more capable of
providing either large benefits or large losses.

\\Big message: even constrained by the same initial depletion and core
biology (e.g. movement rates), a wide range of outcomes are possible for
each species for a given MPA size, dependent on the design strategy,
habitat overlap, migrations, quality of information, etc. Across the
board as MPA size increases the ratio of positive to negative outcomes
increases, until all species achieve their maximum conservation
potential in every iteration at 100% MPA. HOwever, every species
experienced a set of simulations, roughyl xx%, that resulted in net
losses in SSB/SSB0 relative to BAU. Losses were smaller than potential
gaines, but reached meaningful levels in some species, and interestingly
peaked in many cases right before max MPA (reflecting concentration of
fishing in small amounts of habitat in habited by the species).Important
to stress in here that there are no probabilities in this, no way of
knowing which state is more probable than others, just given the range
of things simulated more ways for a bigger positive effect. Also make
clear though, biggish losses (10% or more) possible, but rare.