### 2021-10-12 README.Rmd
Cpue cs (#5)
* interesting. open access now works

* safety stop, things working but tricky to think through optimization

* testing reverse optimization

* adding in hyperallometry

* revised version focused on bycatch

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Something about MPAs and Bycatch

Repo for manuscript exploring the ways in which habitat overlap, fleet spillover, and MPA design strategies can interact to affect conservation and fishery outcomes of MPAs, focused on WCPO-ish system.

starginw thi hastings botsford weakstock species etc. can you hedge around this or not?Build intuition around BWMPAs for weakstocks

Covariance matrix across species: perfectly correlated to totally uncorrelated and everything in between

Clean up by setting Hastings and Botsford as a perfect template for this. PNAS? Isolate the life history

It is clear that if species are mutually exclusive, there is no potential for overlap protection... but where does it fall when overlap is something in between

Hmmm you could use virtual species.. or I guess that's even too complicated

What if you just picked an "environmental" variable, and generated a random matrix

For each species, you then assign a beta to that variable, where when they are all the same the covariance across the species is high. When they are all near zero there is no correlation, when they are opposite signs they are negatively correlated?

Suppose you had two variables actually. One specifies kind of the "default" distributino for all species. The other is an environmental variable, and species can have coefficients that on average range from all 1, evenly split 1 and -1, and near zero

Ah ok what about this. YOu define a set of benchmark species, where those benchmark species are defined by their relationship to a habitat. FOr the remainder of the species, their habitat is on average with noise on a gradient from -1 to 1, where one means they have the same as the benchmark, 0 means they are uncorrlated with the benchmark species, and -1 they are negatively correlated.

```{r}

library(marlin)
library(tidyverse)

theme_set(marlin::theme_marlin())

resolution <- 20 # resolution is in squared patches, so 20 implies a 20X20 system, i.e. 400 patches

seasons <- 1

years <- 50

tune_type <- "depletion"

steps <- years * seasons

# for now make up some habitat

base_hab <- expand_grid(x = 1:resolution, y = 1:resolution) %>%
  mutate(habitat =  .05 * x * y) %>% 
  mutate(habitat = as.numeric(scale(habitat)))

hab_scenes <- expand_grid(scene = c("friends", "enemies","frenemies"), species = fruit[1:10]) %>% 
  group_by(scene) %>% 
  nest()

generate_hab <- function(species, homogifier = 1,scene,base_hab,sigma_hab = 1, beta_hab = 1){

  sigma_hab = .2

  sigma_block <- 0.1

  homogifier <- 0

  species <- hab_scenes$data[[1]]

  block_a <- species$species[1:5]

  block_b <- species$species[6:10]

  block_a_betas <- rnorm(length(block_a),1,sigma_block)

  block_b_betas <- rnorm(length(block_b), homogifier, sigma_block)

  species$betas <- c(block_a_betas, block_b_betas)

  species$habitat <- vector(mode = "list", length = nrow(species))

  for (i in 1:nrow(species)){

    base_hab$realized_hab <- rnorm(nrow(base_hab), base_hab$habitat * species$betas[i], sigma_hab)

    species$habitat[[i]] <- base_hab

  }

  (species$habitat[[1]] %>% 
    ggplot(aes(x,y,fill = realized_hab))+ 
    geom_tile() + 
    scale_fill_viridis_c()) + 
      (species$habitat[[10]] %>% 
    ggplot(aes(x,y,fill = realized_hab))+ 
    geom_tile() + 
    scale_fill_viridis_c())

}

```

# idea for fleet revamp

I think the most straightforward route is just to build a bit more realism in.

1. Burn the model in

2. Calculate costs such that profit per unit effort is 0, where costs are calculated based on f, not effort.

3. Redistribute effort, and going forward use open access dynamics. So, if you set that up right, there will now be a tradeoff between increased catch from effort concentration but also increased cost from f approaching 1.

Worth a shot

```{r}

q <- 0.1

m <- 0.2

effort <- seq(0,100, by = 1)

f <- effort * q

b <- 100

price <- 10

catch <- f / (f + m) * b * (1 - exp(-(f + m)))

biomass <- b * (exp(-(f + m)))

revenue <- catch * price

cost = effort^2

plot(revenue - cost)

plot(revenue)

f <- 10 * q

r1 <- 2 * (f / (f + m) * b * (1 - exp(-(f+ m))) * price - 10^2)

f <- 20 * q

r2 <- (f / (f + m) * b * (1 - exp(-(f+ m))) * price - 20^2)

r1

r2

```
 Yeah that checks out. So, the idea is what? Run the population to equilibrium, and then find a cost coefficient such that system-wide profits are zero

```{r}

q <- 0.1

m <- 0.2

effort <-10

f <- effort * q

b <- 100

price <- 10

cost <- 10

max_delta <- 0.25

cost = (f / (f + m) * b * (1 - exp(-(f+ m))) * price) / effort^2

p1 <- (f / (f + m) * b * (1 - exp(-(f+ m))) * price) - cost * effort^2

theta = log((effort * (1 + max_delta)) / effort) / (f / (f + m) * b * (1 - exp(-(f+ m))) * price)

e2 <- effort * exp(theta * (f / (f + m) * b * (1 - exp(-(f+ m))) * price))

e2 / effort - 1

e2 <- effort * exp(theta * (10000))

e2 / effort

# so what if you tuned it so that if you caught all of ssb0 it would go up by 25%
```

 What about a sigmoid form, so that it's symetric in increase vs decrease?

```{r}

p = seq(-10,10)

x = 1.25 * (1 / (1 + exp(-p)))

plot(p,x)

```