### 2022-10-12 Untitled.qmd
v0.1

---
title: "Untitled"
format: pdf
editor: visual
crossref:
  fig-prefix: Fig.
  tbl-prefix: Tbl
---

## Quarto

Quarto enables you to weave together content and executable code into a
finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that
includes both content and the output of embedded code. You can embed
code like this:

```{r}
1 + 1
```

You can add options to executable code like this, see @fig-test

```{r}
#| label: fig-test
#| fig-cap: "Hello"
plot(1:10)
```

The `echo: false` option disables the printing of code (only output is
displayed).