### 2022-10-12 marlin-paper.qmd

---
title: "Efficient Simulation of Economic and Ecological Dynamics in Aquatic Systems"
author:
    - Daniel Ovando
    - Darcy Bradley
    - Echelle Burns
    - Lennon Thomas
    - James Thorson
date: "`r Sys.Date()`"
format:
  pdf:
    toc: true
    fig-pos: 'H'
  html: default
  docx: default
execute:
  echo: false
  message: false
  warning: false
bibliography: references.bib
params:
  run_name: ["v1.2"]
abstract: "Very rough cut. Communities around the world are increasingly looking to spatial strategies for managing marine ecosystems. However, the design of spatial management strategies faces numerous challenges. Marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond to changes in marine ecosystems in highly complex ways. We present a framework for efficient simulation of the bio-economic dynamics of marine systems, designed to support both management and research. We demonstrate the model's capabilities by focusing on three case studies on the conservation and food security impacts of marine protected areas (MPAs)."
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers

---

```{r setup}
#| echo: false
#| message: false
#| include: false

source(here::here("scripts", "00_setup.R"))

theme_set(marlin::theme_marlin())

sci_to_com <- taxize::sci2comm(unique(marlin_inputs$scientific_name)) %>% 
  bind_rows(.id = "critter") %>% 
  pivot_longer(everything(),names_to = "critter",values_to = "common_name") %>% 
  mutate(common_name = stringr::str_to_title(common_name))

results_files <- list.files(file.path(results_path))

rdses <- results_files[str_detect(results_files,'.rds')]

walk(rdses,~assign(str_remove(.x,".rds"), read_rds(file.path(results_path,.x)), envir = .GlobalEnv)) #load and assign all .rds objects in the appropriate results folder

rdatas <- results_files[str_detect(results_files,'.Rdata')]

walk(rdatas, ~load(file.path(results_path,.x), envir = .GlobalEnv)) # load results into global environment

```

<!-- Single column, double line spaced -->

<!-- Within the word count (no more than 7000-8000 words for Standard Articles, 3000-4000 words for Applications and Practical Tools) -->

<!-- Continuous line and page numbering throughout -->

<!-- Clearly defined manuscript structure as standard: Author details, Abstract (must be numbered 1-4 according to Manuscript Specifications), Keywords, Introduction, Materials and Methods, Results, Discussion, Figures and Tables with captions -->

<!-- Figures and Tables can be embedded within the text where referenced to facilitate reviewing -->

<!--         Author Contributions (if >1 author – see Authorship Policy). -->

<!--         Statement of where you intend to archive your data -->

## Introduction

Communities around the world are increasingly looking to spatial
strategies for managing marine ecosystems. For example, the "30x30"
movement calls for implementation of marine protected areas, a form of
spatial management, across 30% of the world's oceans by 2030. While the
objectives of individual spatial management strategies vary widely,
spatial management strategies are often supported based on their ability
to provide protection to broad swaths of marine ecosystems with one
management action, with potentially fewer costs and complications than
for example assessing, setting, and enforcing catch quotas for
individual species.

However, the design of spatial management strategies faces numerous
challenges. Marine ecosystems are highly dynamic and opaque, and
extractive entities such as fishing fleets respond to changes in marine
ecosystems in highly complex ways. These dynamics post challenges even
for single-species assessment and management strategies [@field2006a],
which are only amplified when we consider management policies
specifically designed around spatial features in the oceans. For
example, policies such as MPAs designed for both conservation and food
security must consider factors such as the optimal size and placement of
a protected area network given a wide array of life history,
exploitation levels, fleet dynamics, and preferred habitats, all of
which may change over time, particularly given the impacts of climate
change.

These challenges are amplified by the fact that our knowledge regarding
the design and implementation of spatial management strategies such as
MPAs is largely based on theoretical studies; the size of protected
areas [@ovando2021a] and time required[@nickols2019] to see substantial
effects from many MPAs means that we have relatively little empirical
evidence as to their large-scale effects under a range of
social-ecological settings [@mcclanahan2021].

Many models have been developed to support the theory and design of
spatial management strategies (see @fulton2015 for a review of existing methods). However, they have tended
to fall into one of either highly complex and tactical models designed
for use in a specific location, or to be highly stylized representations
intended to provide theoretical insights with less relevance for highly
specific design questions. This complex landscape presents a challenge
to stakeholders charged with marine spatial planning exercises lacking
accessible models suited for more tactical applications, and a barrier
to the scientific community, wherein comparisons across spatial
management simulation studies are clouded by discrepancies in underlying
model structures beyond differences in the core phenomenon in question.

In part to help address this challenge, we present here a novel
bio-economic modeling framework called `marlin` that allows for
efficient simulation of spatio-temporal biological and economic
dynamics. `marlin` allows users to simulation the impacts spatial
policies such as MPAs across a range of species targeted by various
fleets across heterogeneous and dynamic seascapes. Here we present the
core methods and functionality of this framework, and demonstrate its
use in three xx case-study applications. In the first, we show how
`marlin` can be used to assess critical drivers for positive outcomes
and unintended consequences of MPAs as a bycatch-mitigation strategy. In
the second, we present a more tactical application, demonstrating how
`marlin` can be used to compare and design MPA networks to achieve
societal objectives.

## Methods

`marlin` follows a "models of intermediate complexity" (MICE) philosophy [@plagányi2014]
targeted specifically towards capturing heterogeneous habitats, movement dynamics, and technical interactions across species. In its general form, `marlin` simulates the behavior of populations of spatially explicit biologically independent age-structured animal populations affected by one or more fishing fleets in time steps specified by the user (generally annual or finer).  While `marlin`
does not simulate species interactions, fishing fleets in `marlin` are
capable of targeting and affecting multiple species simultaneously. This allows the model to simulate processes such as fisheries bycatch or effort displacement in a way that accounts for the ways that changes in fleet behavior might affect multiple species in the system. The ability to track the impacts of fishing fleets across multiple species simultaneously is particularly important as very few fishing fleets are truly single species.

Below we provide a summary of the `marlin` model, as well as details of
the generalized simulation exercises and case study analyses. We note that `marlin` is a complex package with many options to control detailed behaviors. Given that these specific features will evolve over time, we do not describe in detail every option of the model here. Rather, we focus on the equations and methods describing the core foundations of our framework for efficiently simulating bio-economic dynamics in aquatic systems.

### Fleet Dynamics

A fishing fleet is defined in `marlin` by a total pool of
effort that can have different catchability, prices, and selectivity for
individual species within a simulation (each of which we denote as a *metier*. For example, both a longline and
purse seine fleet may capture bigeye (*Thunnus obesus*) and skipjack
(*Katswamus pelamus*) tunas, but the longline fleet is much more likely to
catch bigeye than skipjack, and *vice versa*. Each fishing fleet
generates catches, revenues, and profits from individual species that it
targets, and makes decisions based on its total objective function
across all species.

Revenues *R* for fleet *f* in time step *t* are given by

$$
 R_{t,p,f} = \sum_s^S \sum_a^A \Pi_{s,f}\frac{\alpha_{a,s,f}q_{s,f}E_{t,p,f}}{u_{t,p,s,a}} \times Y_{t,p,s,a}
$$ {#eq-revenue}

where *p* is patch, *s* is species, *a* is age, $\alpha$ is selectivity
at age, *q* is catchability , *u* is total fishing mortality, $\Pi$ is
price per unit weight, and *Y* is total catches.

$$
Y_{t,p,s,a} = \frac{u_{t,p,s,a}}{z_{t,p,s,a}} \times b_{t,p,s,a} \times (1 - e^{-z_{t,p,s,a}})
$$ {#eq-yield}

where *b* is biomass and *u* is total fishing mortality, given by

$$
u_{t,p,s,a} = \sum_{f=1}^F\alpha_{a,s,f}q_{s,f}E_{t,p,f}
$$ {#eq-u}

Selectivity $\alpha$ is modeled through a logistic form per

$$
\alpha_{a,s,f}=\frac{1}{(1 + e^{-log(19)\times\frac{l_{a,s} - l^{sel}_{s,f}}{\delta^{sel}_{s,f}}})}
$$ {#eq-selectivity}

where $l$ is length at age, $l^{sel}$ is the length at 50% selectivity,
and $\delta$ is the difference between the length at 50% selectivity and
the length at 95% selectivity.

or through a dome shaped form of

XX add dome shaped equation XX

*z* is total mortality given by

$$
 z_{t,p,s,a} = u_{t,p,s,a} + m_{s,a}
$$ {#eq-z}

Where *m* is natural mortality.

This means that each species experiences a total mortality at age in a
time step in a patch, individual fractions of which are portioned off as
catches and subsequently revenues for each fleet. Given revenues *R*, we
then calculate profits $\phi$ per

$$
\phi_{t,p,f} = R_{t,p,f} - C_{t,p,f}
$$ {#eq-profits}

where costs *C* are calculated

$$C_{t,p,f} = \gamma_f  \left( E_{t,p,f}^{\beta_f} + \eta_{f,p} E_{t,p,f} \right)$$

where $$\gamma_f$$ the scales the cost per unit effort for fleet *f*, $\beta$ allows for the cost of effort to scale non-linearly, and $\eta$ is an additional cost per unit effort of fishing in each patch *p*, allowing for the model to account for travel costs for different patches. Travel costs are calculated based on the euclidean distance of each patch to the nearest port of a given fleet, users can specify any number from zero to the number of patches of port locations. When no ports are specified, travel costs are zero.

#### Calculating Total Effort

`marlin` allows for two general modes of effort. The simplest is "constant effort", in which the total effort of each fishing fleet remains constant over time.

The more complex option is "open access". Under open access, the total amount of effort in a given time step for fleet *f* is a function of the profitability of that fleet in the proceeding time steps in which fishing was open.

$$
E_{t+1,f} = E_{t,f} \times e^{\theta_f log(R_{t,f} / C_{t,f})} 
$$ {#eq-oa}

There $\theta$ controls the responsiveness of fleet *f* to the log of the ratio of total revenues *R* to total costs *C*, and approximates the percentage amount effort changes in response to a one unit changes in the log revenue to cost ratio.

Three general exceptions exist for the effort dynamics. When MPAs are implemented, effort that used to operate inside the MPAs can either leave the fishery, or be redistributed outside the the MPA (the default behavior). The model can also accept total catch quotas for each species in each time step. When quotas are activated, if the total catch across all fleets for a given species under the model's effort dynamics would exceed the allowable quota for that species, the total effort for each fleet is decrease proportional to its contribution to the total catch until the quota is satisfied. `marlin` also allows users to specify closed fishing seasons for one or more species in the system. During closed seasons, no effort is allocated to patches that would result in capture of the covered species, potentially resulting in zero effort if the species cannot be avoided. Under these conditions, the amount of effort in the next open season is a function of the effort and profits in the last open fishing season.

#### Tuning Fishing Fleets

The degree of fishing pressure exerted by a given set of fishing fleets on each species is a function of a range of parameters including the total amount of effort, the selectivity ogives, cost functions, the spatial distribution of the resource, the relative prices across species, and the catchability coefficient of each fleet for each species. `marlin` provides a tuning option to help users achieve desired biological outcomes from their fleets.

Users can tune their fleets in one of two ways. First, they can specify a target exploitation rate *u* for each species in their model run. Taking all the other parameters of the model as given, `marlin` then adjusts the catchability coefficients *q_{f,s}* for each fleet *f* and species *s* such that the equilibrium exploitation rate for each species matches the desired level. Second, they can specify a target depletion, defined as total spawning stock biomass under fishing divided by total unfished spawning stock biomass, and the model will adjust the catchability coefficients *q_{f,s}* for each fleet *f* and species *s* such that the equilibrium depletion for each species matches the desired level.

#### Spatial Allocation of Effort

Each fishing fleet decides how to allocate its effort in space
based on one of four possible spatial allocation strategies. The ideal
free distribution (IFD) is the standard method for distributing fishing
fleets in spaces (see @gillis2003 and references therein). However, the
IFD presents a number of complications for our model. First, under the constant effort model, the total amount of effort required to satisfy the IFD may not equal the total amount of
effort required by the model. Second, even if we were to implement the
IFD for a single fleet, the model is capable of tracking multiple fleets
at once. This means that the IFD for a single fleet would be a
game-theoretic outcome of the actions of each of the individual fleet
conditional on the actions of all other fleets. While possible to
implement this would slow down our model runs to the point of
practically preventing large-scale evaluation of the effects of MPAs on
bycatch.

As such we explored a series of "next best" fleet distribution
algorithms. While not the IFD in any individual time step, over time
they start to approximate the IFD, as each assumes that fleets base
their decisions for the current time step on the outcomes in the prior
timestep, meaning that the impacts of the actions of other fishing
fleets are somewhat internalized. The four possible fleet distribution
algorithms are

1.  Revenue per unit effort (RPUE): The fleet distributes itself in
    space based on revenue per unit effort
2.  Revenue: The fleet distributes itself in space based on revenue
3.  Profit per unit effort (PPUE): The fleet distributes itself in space
    based on profit per unit effort
4.  Profit: The fleet distributes itself in space based on profit

Revenue based spatial distribution is not likely to be very realistic .
However, due to the complexity of parameterizing cost functions fleet
dynamics are often evaluated based on yield or revenue alone, and so we
include those scenarios here to evaluate the potential implications of
this choice. The decision on whether to allocate the fleet based on
absolute or relative (per unit effort) is more complex. When effort
represents the actions of separate and individual fishing actors (e.g.
independent fishing vessels), a per-unit-effort strategy may be more
realistic, in which fishers distribute themselves based on the expected
catch of their individual efforts [@hilborn1987]. Conversely, a system
defined by a sole owner seeking to maximize total profits might be
better represented by a fleet model based on total profits

### Population Model

The underlying population model used is an age structured single-species
model in the manner of @ovando2021a.

For the population model, numbers *n* at time *t* for age *a* are given
by

$$
n_{t,p,s,a} = \begin{cases}
      = BH(ssb_{t-1,p,s,a}) & \text{if $a = 1$}\\
     = n_{t-1,p,s,a-1}e^{-(z_{t-1,p,s,a-1}}, & \text{if $1< a < max(age)$}\\
     =  n_{t-1,p,s,a}e^{-(z_{t-1,p,s,a})} + n_{t-1,a-1}e^{-(z_{t-1,p,s,a-1})}, & \text{if $a = max(a)$}
  \end{cases}
$$ {#eq-pop}

where *BH* is the Beverton-Holt recruitment function, *ssb* is
spawning-stock-biomass.

where $l_a$ is the mean length at at age, $l_sel$ is the length at which
on average 50% of individuals are selected by the fishery, and
$\delta_{sel}$ are the additional units of length at which on average
95% of fish are selected by the fishery.

*ssb* is calculated by converting age to mean length, calculating weight
at age, maturity at age, and then calculating spawning stock biomass as
the sum of spawning potential at age in a given time step.

$$
  l_{a,s} = l_{\infty,s}\left(1 - e^{-k_s(a - a0_{s})}\right)
$$ {#eq-len}

Biomass *b* at age is then given by

$$
  b_{a,s} = wa_s \times l_{a,s}^{wb_s}
$$ {#eq-weight}

and maturity *mat* is calculated as

$$
  mat_{a,s} = \frac{1}{(1 + e^{-log(19)\times\frac{l_{a,s} - lmat_s}{{\delta}mat_s}}}
$$ {#eq-mat)}

where $l_{mat}$ is the length at which on average 50% of individuals are
sexual maturity, and $\delta_{mat}$ is the units of length beyond
$l_{mat}$ at which on average 95% of fish are sexually mature.

Spawning stock biomass at time *t* is then calculated as

$$
  ssb_{t,p,s} = \sum_{a=1}^Aw_{a,s,t}^{\gamma_s} mat_{t,a} n_{t,p,s,a}
$$ {#eq-ssb}

#### Recruitment

Recruitment follows Beverton-Holt dynamics. We do however allow for
three variants in the timing of density dependence:

1.  Global density dependence: Density dependence is a function of the
    sum of spawning biomass across all patches, and recruits are then
    distributed according to habitat quality

$$
n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p}}\times{h}\times\sum_{p=1}^P{ssb_{t-1,p}}}{0.2\times{\sum_{p=1}^Pssb0_p}\times(1 - h)+(h - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p}}}\right)\times {hab_p}
$$ {#eq-dd2}

where *hab* is a vector of recruitment habitat quality by patch degined
by

$$hab_p = r0_p / \sum_p^P{r0_p}$$

2. Local density dependence: Density
dependence occurs independently in each patch and recruits are retained
in their home patch.

$$
n_{t,p,s,a = 1,p} = \left(\frac{0.8{\times}r0_{s,p}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p}}}\right)
$$ {#eq-dd1}

3.  Local density dependence then disperse: Density dependence occurs
    independently in each patch and recruits are then dispersed.

$$
  n_{t,p,s,a = 1,p} = \left(\frac{0.8{\times}r0_{s,p}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p}}}\right)\times \boldsymbol{d^l} \times \epsilon_t
$$ {#eq-dd3}

where **d^l^** is the recruitment movement matrix

4.  Post-dispersal density dependence: Larvae are distributed throughout
    the system, and then density dependence occurs based on the density
    of adult biomass at the destination patch.

$$
  larv_{t,p} = ssb_{t-1} \times\boldsymbol{d^l}
$$ {#eq-larvmove}

$$
n_{t,a = 1,p} = \left(\frac{0.8{\times}r0_{p}\times{h}\times{larv_{t,p}}}{0.2\times{ssb0_p}\times(1 - h)+(h - 0.2)\times{larv_{t,p}}}\right) \times \epsilon_t
$$ {#eq-dd4}

#### Dispersal

`marlin`'s movement dynamics are based on a continuous-time Markov
chain (CTMC), as described in @thorson2021a. Within this framework, the model
allows for movement to be broken down into three components of
*diffusion*, *advection* (follow currents), and *taxis* (active movement
towards say a preferred habitat). For now, our model focuses just on the diffusion and taxis components of this model, though future extensions could easily incorporate advection vectors from say ROMs models (citations). In this way then, `marlin` allows users to run anything from a simple Gaussian dispersal kernel up to a system governed by species that
passively diffuse around a dynamic thermal range.

@thorson2021a described the exact steps of the CTMC approach in detail. We briefly the general CTMC method here, but see @thorson2021a for a detailed description of the method. Under this framework, movement from each patch to each other patch in the system in a given timestep *t* for life stage *g* of species *s* is defined by a movement matrix $\boldsymbol{M}_{t,s,g}$. $\boldsymbol{M}_{t,s,g}$ is calculated as the matrix exponent of the sum of the diffusion and taxis matrices scaled by the width of the time step that the model is running under.

$$
\boldsymbol{M}_{t,s,g} = e^{\Delta t(\boldsymbol{D}_{t,s,g} + \boldsymbol{X}_{t,s,g})} 
$$ {#eq-movement}

Where $\boldsymbol{D}$ is the diffusion matrix and $\boldsymbol{X}$ is the taxis matrix.

The diffusion matrix is a function of a user specified diffusion rate, see @thorson2021a. We slightly modify the taxis matrix in the following manner

xx need to fix notation somewhere here, see this line of code for much clearer depiction of what's happening, since sometimes code is cleaner. https://github.com/DanOvando/marlin/blob/3563e7319ad0267f89789267a4c6ee58a9ad0d60/R/fish_class.R#L525 xx

$$
n_{t,p,s,a} = \begin{cases}
      = (1 + \lambda d_{s,f})max(0, h(p2_{t,s,g}) - h(p1_{t,s,g})) & \text{if p2 and p1 are adjacent}\\
     = -\sum_{g' \neq g1}\{ (1 + \lambda d_{s,f})max(0, h(p2_{t,s,g}) - h(p1_{t,s,g}))\} & \text{if p1 = p2}\\
     = 0 & \text{otherwise.}
  \end{cases}
$$ {#eq-taxis}

The $lambda d_{s,f}$ term is a scalar in which the magnitude of the habitat gradient between patches can be scaled relative to the diffusion parameter $d$.

For both the diffusion and taxis matrix, we allow for the inclusion of physical barriers to movement (i.e. land). Pairs of patches that are adjacent but in which one or both patches are a barrier to movement are set as non-adjacent. The CTMC model then produces movement dynamics that move around barriers rather than over them. This is different than setting habitat in a patch to zero, which implies a patch that a given animal will not move towards for its own sake, but can move over towards another better patch.

While this CTMC approach to movement simulation is to date not commonly seen in the
marine modeling literature, it has numerous advantages that warrant its
seeming complexity. First, when only diffusion is present, the model
will generalize to the familiar dynamics of a Gaussian dispersal kernel. However, the time component *t* in @eq-movement allows the movement model to be run on arbitrarily fine time steps. Second, the *taxis* model allow for more active habitat choices by
species, allowing us to simulate both preferences of species in space
and time efficiently.Lastly, this form has the advantage that the parameters are directly
estimable from real data. So, if provided with say CPUE data and a
tagging study, users can estimates the movement parameters in the same
manner as @thorson2021a, and then pass their estimated parameters to
`marlin` for simulation.

#### Reference Points

Sigh. I need to work on this section, but I am *le tired*.

A note here Why no reference points? The problem: Fmsy and MSY are
selectivity dependent, and once we allow for different fleets with
different selectivities that critical can move in space, that movement
affects the net selectivity and fishing mortality on the population.

Bmsy is also affected, since if selectivity is very low, you might need
more biomass of immature individuals relative to less biomass of more
mature individuals; a biomass of X can be more or less reproductively
productive depending on the size / age makeup of that biomass.

SSBmsy is more conserved, as over the long term the key things
is the flow of recruits, so we do provide an option for that, but I haven't messed with it for a while.

There are many different strategies for estimating reference points
conditional on selectivity and regimes, see citations. We leave it to
users to find relevant reference point as required by their methods on
their own, using the sort of MSE examples available in the package.

### MPA Optimization

Our MPA optimization algorithm seems to create an MPA network that
maximizes an objective function, given by

$$
obj_p = \Phi \times scale(\sum_f^F \boldsymbol{\phi_{f,p}}) + (1 - \Phi)  scale(\boldsymbol{\sum_s^S\frac{ssb_s}{ssb0_{s,p}}})
$$ {#eq-obj}

Where *scale* rescales the values to be between 0 and 1 such that the
weight parameter is comparable between vectors. The algorithm works by
first assigning every patch in the system to an MPA. For the first
round, the algorithm then calculates the equilibrium objective function
value resulting from sequentially removing each individual patch from
the MPA network. The cell whose removal from the MPA produces the
highest objective function value is then removed from the MPA network,
and the marginal value of the objective function each of the remaining
cells from the last steps are then stored. For the next steps, we
sampled 25% of the remaining MPA cells with probability proportional to
their marginal contribution to the objective function in the prior
iteration, selecting the cell whose removal produces the highest total
objective function as the cell to remove from the network, and updating
the marginal value of the sampled-but-not-selected cells. This in
essence is a sampling-importance-resampling algorithm. The algorithm
continues until no more cells are left in the MPA. This then provides
for a given objective weighting value $\Phi$ the near-optimal MPA
network at every MPA size from 0 to 100%. We say "near" as they is some
path dependency here, in that for example the network that maximizes the
objective function with 40% of the MPAs does not necessarily contain the
optimal network at 30%. However, the reversed nature of the algorithm,
starting at 100% MPA coverage and working better, produces a more stable
network with regards to this problem than starting at 0% and growing to
100%.

### Use Cases

We include three examples demonstrating how `marlin` can be used to
support marine spatial planning. In the first, we show how `marlin` can
be used to compare the total and distributional impacts of marine
protected areas design in a heavily fished coral reef ecosystem. In the
second, we demonstrate how `marlin` can be used to assess components of
climate resilience of alternative MPA design strategies. In the third,
we show how `marlin` can be used to run large-scale simulation testing
of management strategies.

#### Rule-of-thumb MPA Design Strategies

Across all of our case studies we make use of four potential "rule of
thumb" MPA design strategies:

1.  Rate: MPAs are placed in based on the pre-MPA depletion-weighted
    catch relative to the total catch in a patch. So, patches with high
    rates of catch of depleted species relative to total catch are
    prioritized.

2.  Avoid_fishing: MPAs locations are prioritized inversely to fishery
    catches. So, patches with low catches are prioritized over patches
    with high catches.

3.  Target_fishing: MPAs locations are prioritized proportionally to
    fishery catches. So, patches with high total catches are prioritized
    over patches with low catches.

4.  Area: For the "area" strategy, we simply assign patches in
    sequential order until the desired total MPA size is reached.

MPAs never to our knowledge randomly placed; their locations are decided
based on stakeholder negotiation based on available information and
objectives. In theory the design MPA networks can be optimized through
the use of a modeling framework, and indeed conditional on the validity
of the model itself this process is likely to produce better outcomes
than manually-designed strategies [@rassweiler2012; @rassweiler2014].
However, in practice the ability of communities to conduct numerical
optimization of MPA networks can be constrained by the number of
objectives and dimensionality of the operating model, or by simple
capacity of local stakeholders. As such, we show here how `marlin` can
be used to examine the performance of potentially sub-optimal but more
broadly usable design strategies xx.

#### Coral Reefs

In our coral reef example, we model the dynamics of four species: a
grouper (XX), a shallow-reef snapper (xx), a deep-water snapper (XX),
and a reef shark (XX). Groupers undergo a mass migration to a spawning
aggregation once per year, followed by the sharks. Shallow-reef snappers
stay in reefs closer to shower year-round, while deep-water snappers
stay in the deeper reefs (@fig-coral).

These species are targeted by two different fleets. Fleet one primarily
targets the grouper and nearshore snapper populations, while obtaining
some funds from sharks. Fleet one has a dome shaped selectivity pattern
for both grouper and shallow-reef snapper due to demand for "plate
sized" fish (citation), and logistic selectivity for sharks, as any
shark caught is retained. This fleet is a subsistence fleet,
where the local community takes every possible opportunity to fish that
they can, and as such we model it as a "constant effort" fishery.

The second fleet primarily targets the deep-water snapper population.
While plate-sized snapper are still prized from this population, we
model fleet two's selectivity as logistic due to high levels of discard
mortality for deep-water snapper resulting from barotrauma (xx). This
fleet catches some grouper, though less than fleet one, and receives no price for
sharks due to the requirements of a certification program through which
they sell their deep-water snapper, though they do accidentally catch and kill them on occasion. This fleet operates under open-access
dynamics, as fishing is not the primary means of subsistence and so
short-term effort expands and contracts in respond to profitability of
the primarily grouper-driven fishery xx.

```{r}
#| label: fig-coral
#| fig-cap: "Distribution of animals each season under unfished conditions. Numbers show the approximate location of each fleet's port."

coral_fauna <- coral_fauna_and_fleets$fauna

coral_fleets <- coral_fauna_and_fleets$fleets

grid <- expand_grid(x = 1:resolution, y= 1:resolution) %>%
  mutate(patch = 1:nrow(.))

mpa_locations <- expand_grid(x = 1:resolution, y = 1:resolution) %>%
  mutate(mpa = TRUE)

coral_sim <- simmar(
  fauna = coral_fauna,
  fleets = coral_fleets,
  manager = list(mpas = list(
    locations = mpa_locations,
    mpa_year = 1
  )),
  years = 2
)

patch_biomass <-
  map_df(coral_sim, ~ map_df(.x, ~tibble(biomass = rowSums(.x$ssb_p_a), patch = 1:nrow(.x$ssb_p_a)), .id = "critter"), .id = "step") %>%
  mutate(step = as.numeric(step)) %>%
  left_join(grid, by = "patch")

patch_biomass %>%
  group_by(critter, step) %>%
  mutate(sbiomass = biomass / max(biomass)) %>%
  ungroup() %>%
  filter(between(step,1,2)) %>% 
  mutate(Season = step) %>% 
  ggplot()+
  geom_tile(aes(x,y, fill = sbiomass),show.legend = FALSE) +
  geom_text_repel(data = coral_ports, aes(x, y, label = fleet), color = "red") +
  facet_grid(Season~critter) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_viridis_c() + 
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank())

```

We then simulate the impacts of MPAs...

#### Pelagic Systems

We modeled our pelagic use case loosely on the characteristics of the
Western and Central Pacific Ocean tuna fisheries. We simulate
trajectories of XX species commonly caughts in the region, including
both the highly abundance skipjack tuna (*Katsuwonus pelamis*) to the
heavily depleted oceanic whitetip shark (*Carcharhinus longimanus*). We
pulled publicly available data on catch-per-unit-effort of each of these
species from the WCPO as a proxy for baseline habitat distributions
(Fig.XX). However, for the pelagic case study we simulate a scenario
where climate change is inducing a northerly migration of XX degress per
year in XX species.

These species are targeted by a longline fleet that targets large adult
tunas such as bigeye and yellowfin primarily for high-grade consumption,
and a purse-seine fishery that primarily targets skipjack tunas for bulk
canning. XX. Both fleets are subject to a quota XX add this back in XX.

```{r}
#| label: fig-blue
#| fig-cap: "Distribution of animals under unfished conditions for blue-water simulation."

blue_fauna <- blue_water_fauna_and_fleets$fauna

blue_fleets <- blue_water_fauna_and_fleets$fleets

grid <- expand_grid(x = 1:resolution, y= 1:resolution) %>%
  mutate(patch = 1:nrow(.))

mpa_locations <- expand_grid(x = 1:resolution, y = 1:resolution) %>%
  mutate(mpa = TRUE)

blue_sim <- simmar(
  fauna = blue_fauna,
  fleets = blue_fleets,
  manager = list(mpas = list(
    locations = mpa_locations,
    mpa_year = 1
  )),
  years = 2
)

patch_biomass <-
  map_df(blue_sim, ~ map_df(.x, ~tibble(biomass = rowSums(.x$ssb_p_a), patch = 1:nrow(.x$ssb_p_a)), .id = "critter"), .id = "step") %>%
  mutate(step = as.numeric(step)) %>%
  left_join(grid, by = "patch")

patch_biomass %>%
  group_by(critter, step) %>%
  mutate(sbiomass = biomass / max(biomass)) %>%
  ungroup() %>%
  filter(step == max(step)) %>% 
  ggplot()+
  geom_tile(aes(x,y, fill = sbiomass),show.legend = FALSE) +
  facet_wrap(~critter) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_viridis_c() + 
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank())

```

#### Key Drivers

The MPA literature points to the ability of a wide range of drivers to
affect the performance of MPAs with regards to conservation and economic
outcomes such as adult and larval dispersal [@gerber2005], the timing of
density dependence [@babcock2011], correlations in space and time across
species [@pons2022; @abbott2012] , fleet dynamics and economic
incentives [@smith2003a; @cabral2019], relationships between size and
fecundity (e.g. hyperallometry) [@marshall2021], heterogeneity in stock
status [@hastings2017; @abbott2012], and many other traits [@gaines2010;
@hilborn2004a].

These studies have all helped guide the science of marine protected
areas, but often focus on modeling the impact of subset of variables of
interest at a time. We demonstrate here how `marlin` can be used to
conduct Monte Carlo simulations across a wide range of drivers to assess
the relative contributions each to the outcomes of protected areas under
a particular set of circumstances.

For this example we model the dynamics of a skipjack tuna *Katsuwonus
pelamis*, a striped marlin *Kajikia audax*, and a oceanic whitetip shark
*Carcharhinus longimanus*. Life history parameters for each species were
pulled from FishLife (@thorson2020), with the exception of steepness,
set to XX, and fecundity. For the tuna and marlin species, fecundity is
assumed to scale with weight modified by the hyperallometric exponent.
For the shark species, we assume a constant value of six pups per year
at all ages greater than the age at maturity. We assume that tunas have
an unfished biomass of XX, marlins XX, and sharks XX, to reflect a
large, medium, and small population.

For the generalized experiments we only include one fishing fleet. We
assign a price per KG of \$XX for tuna, intended to reflect a low value
high volume fishery, \$XX for marlin for a higher value low volume
fishery, and \$0 for sharks to reflect a true bycatch species. We assign
a selectivity of 60% of length at maturity for skipjack to allow for
some overfishing when F/M is high enough, 70% for marlin, and 90% for
shark xx.

We ran `r nrow(state_experiments)` state experiments, varying the state
of nature experienced by a simulation. The candidate states include

| State                  | Distribution          | Description                                     |
|------------------|------------------|-----------------------------------|
| $\sigma_{centroid}$    |                       | Standard deviation of habitat centroid          |
| $\sigma_{habitat}$     |                       | Standard deviation around centroid              |
| Spatial catchability   |                       | Catchability varies in space                    |
| Spatial allocation     |                       | Fleet allocation strategy                       |
| F/M                    |                       | Fishing mortality relative to natural mortality |
| Adult diffusion rate   |                       |                                                 |
| Recruit diffusion rate |                       |                                                 |
| Hyperallometry         | $U(1,2)$              | Presence of hyperallometric fec                 |
| DD timing              |                       | Timing of density dependence                    |
| Seasonal Movement      | $\subset[TRUE,FALSE]$ | Seasonal Movement Present                       |

We simulate degrees of species overlap using two parameters, $\sigma_c$
and $\sigma_h$. All species start with the assumption that the centroid
of their preferred habitat occurs in the center-most patch of the two
dimensional system simulated by marlin. For our simulations, we then
generated a random deviation from that default centroid for each species
using $\sigma_c$ in the manner

$$centroid_s \sim N(center,\sigma_c)$$ where *s* denotes species, and
*center* is the coordinate of the central cell. In this way, the larger
$\sigma_c$ grows, the more variation there is between the centroid of
the preferred habitat for each species.

We then control the spread of a creatures preferred habitat around that
centroid per a Gaussian dispersal kernel in the manner

$$habitat_{p,s} = \frac{1}{\sqrt{2\pi\sigma_h^2}}e^{-\frac{d_p^2}{2\sigma_h^2}}$$

Where $d$ is the euclidean distance from the $centroid_s$ to patch $p$
and $\sigma_h$ controls the dispersal of the habitat around that range.

For each of these `r nrow(state_experiments)` state experiments, we then
apply `r nrow(placement_experiments)` placement experiments defined by
an MPA size, an MPA placement strategy, and the number of animals to be
considered in the MPA planning, ranging from one to three, where one
indicates that only one animal is considered in the MPA design process,
while three means that all three species are considered.

## Results

### Coral Reefs

```{r}
#| label: fig-coral-results
#| fig-cap: "Something something."
#| fig-height: 8
#| fig-width: 6

coral_mpa_experiements$mpas <- map(coral_mpa_experiements$results, "mpa")

coral_mpa_experiements$obj <- map(coral_mpa_experiements$results, "obj")

coral_results <- coral_mpa_experiements %>%
  unnest(cols = obj)

coral_fleet_frontier <- coral_results %>%
  pivot_longer(starts_with("fleet_"), names_to = "fleet", values_to = "fleet_yield") %>%
  group_by(prop_mpa, fleet, placement_strategy) %>%
  summarise(yield = sum(fleet_yield),biodiv = sum(unique(biodiv))) %>%
  ggplot(aes(biodiv, yield, color = placement_strategy)) +
  geom_line() +
  scale_x_continuous(name = "Change in Total Depletion",limits = c(0, NA)) +
    scale_y_continuous(name = "Total Yield",limits = c(0, NA)) +
  facet_wrap(~ fleet, scales = "free_y")

coral_yield <- coral_results %>% 
  group_by(prop_mpa, placement_strategy) %>% 
  summarise(biodiv = sum(biodiv), yield = sum(yield)) %>% 
  group_by(placement_strategy) %>% 
  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1,
         delta_biodiv = biodiv / biodiv[prop_mpa == 0] - 1) %>% 
  ggplot(aes(delta_biodiv, delta_yield, color = placement_strategy)) + 
  geom_vline(aes(xintercept = 0), linetype = 2) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_line(size = 2) + 
  scale_x_continuous(name = "Change in Conservation", labels = scales::label_percent(accuracy = 1)) +
  scale_y_continuous(name = "Change in Yield", labels = scales::label_percent(accuracy = 1)) + 
  theme(legend.position = "top")

coral_fleet_yield <- coral_results %>%
  pivot_longer(starts_with("fleet_"),
               names_to = "fleet",
               values_to = "fleet_yield") %>%
  group_by(prop_mpa, fleet, placement_strategy) %>%
  summarise(yield = sum(fleet_yield), biodiv = sum(unique(biodiv))) %>%
  group_by(fleet, placement_strategy) %>%
  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1) %>%
  ggplot(aes(prop_mpa, delta_yield, color = fleet)) +
    geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_line() +
  facet_wrap( ~ placement_strategy, scales = "free_y") +
  scale_x_continuous(name = "MPA Size",
                     labels = scales::label_percent(accuracy = 1)) +
  scale_y_continuous(name = "Change in Yield",
                     labels = scales::label_percent(accuracy = 1))

coral_yield / coral_fleet_yield

```

### Pelagic Systems

```{r}
#| label: fig-pelagic
#| fig-cap: "Placeholder for pelagic results is we stick with them"

blue_water_climate_experiments$mpas <- map(blue_water_climate_experiments$results, "mpa")

blue_water_climate_experiments$obj <- map(blue_water_climate_experiments$results, "obj")

blue_results <- blue_water_climate_experiments %>%
  unnest(cols = obj)

blue_fleet_frontier <- blue_results %>%
  pivot_longer(names(blue_fleets), names_to = "fleet", values_to = "fleet_yield") %>%
  group_by(prop_mpa, fleet, placement_strategy) %>%
  summarise(yield = sum(fleet_yield),biodiv = sum(unique(biodiv))) %>%
  ggplot(aes(biodiv, yield, color = placement_strategy)) +
  geom_line() +
  scale_x_continuous(name = "Change in Total Depletion",limits = c(0, NA)) +
    scale_y_continuous(name = "Total Yield",limits = c(0, NA)) +
  facet_wrap(~ fleet, scales = "free_y")

blue_frontier <- blue_results %>% 
  group_by(prop_mpa, placement_strategy) %>% 
  summarise(biodiv = sum(biodiv), yield = sum(yield)) %>% 
  group_by(placement_strategy) %>% 
  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1,
         delta_biodiv = biodiv / biodiv[prop_mpa == 0] - 1) %>% 
  ggplot(aes(delta_biodiv, delta_yield, color = placement_strategy)) + 
  geom_vline(aes(xintercept = 0), linetype = 2) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_line(size = 2) + 
  scale_x_continuous(name = "Change in Conservation", labels = scales::label_percent(accuracy = 1)) +
  scale_y_continuous(name = "Change in Yield", labels = scales::label_percent(accuracy = 1)) + 
  theme(legend.position = "top")

blue_fleet_yield <- blue_results %>%
  pivot_longer(names(blue_fleets),
               names_to = "fleet",
               values_to = "fleet_yield") %>%
  group_by(prop_mpa, fleet, placement_strategy) %>%
  summarise(yield = sum(fleet_yield), biodiv = sum(unique(biodiv))) %>%
  group_by(fleet, placement_strategy) %>%
  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1) %>%
  ggplot(aes(prop_mpa, delta_yield, color = fleet)) +
    geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_line() +
  facet_wrap( ~ placement_strategy, scales = "free_y") +
  scale_x_continuous(name = "MPA Size",
                     labels = scales::label_percent(accuracy = 1)) +
  scale_y_continuous(name = "Change in Yield",
                     labels = scales::label_percent(accuracy = 1))

blue_critter_bio <- blue_results %>%
  group_by(prop_mpa, critter, placement_strategy) %>%
  summarise(biodiv = sum(unique(biodiv))) %>%
  group_by(critter, placement_strategy) %>%
  mutate(delta_bio = biodiv / biodiv[prop_mpa == 0] - 1) %>%
  filter(prop_mpa < 0.4) %>% 
  ggplot(aes(prop_mpa, pmin(1,delta_bio), color = critter)) +
    geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_line() +
  facet_wrap( ~ placement_strategy) +
  scale_x_continuous(name = "MPA Size",
                     labels = scales::label_percent(accuracy = 1)) +
  scale_y_continuous(name = "Change in Spawning Biomass",
                     labels = scales::label_percent(accuracy = 1))

# blue_frontier / blue_critter_bio + plot_layout(heights = c(1,2))

blue_critter_bio
```

### Key Drivers

```{r}
#| label: fig-drivers
#| fig-cap: "Distribution of conservation and food security impacts of MPAs covering 30% of available habitat across a range of simulations."
#| fig-height: 7
#| fig-width: 7

var_effects <- driver_results %>%
  ungroup() %>%
  mutate(biodiv_effect = biodiv / bau_biodiv - 1,
         econ_effect = econ / bau_econ - 1) %>%
  select(
    biodiv_effect,
    f_v_m,
    econ_effect,
    critter,
    adult_diffusion,
    recruit_diffusion,
    density_dependence,
    placement_strategy,
    sigma_habitat,
    hyper,
    seasonal_movement,
    fleet_model,
    spawning_aggregation,
    seasonal_movement,
    prop_mpa
  )

continuous <- map_lgl(var_effects, is.numeric)

bio_effects <- var_effects[,continuous] %>% 
  filter(between(prop_mpa,0.29, .31)) %>%
  select(-econ_effect, - prop_mpa) %>% 
  pivot_longer(-c(contains("_effect")),names_to = "variable", values_to = "value")

econ_effects <- var_effects[,continuous] %>% 
  filter(between(prop_mpa,0.29, .31)) %>%
  select(-biodiv_effect, - prop_mpa) %>% 
  pivot_longer(-c(contains("_effect")),names_to = "variable", values_to = "value")

biodiv_effects_plot <- bio_effects %>%
  ggplot(aes(value, pmin(2.5,biodiv_effect))) +
  geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
  geom_bin2d(show.legend = FALSE) +
  scale_fill_viridis_c() +
  scale_y_continuous(name = "Conservation Impact", labels = scales::label_percent(accuracy = 1)) + 
  scale_x_continuous(name = "") +
  facet_wrap(~variable, scales = "free_x") +
  theme(axis.text.x = element_blank(),
        plot.margin = unit(c(0,0,0,0), units = "points"))

econ_effects_plot <- econ_effects %>%
  ggplot(aes(value, pmax(-1,econ_effect))) +
  geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
  geom_bin2d(show.legend = FALSE) +
  scale_fill_viridis_c() +
  scale_y_continuous(name = "Food Impact", labels = scales::label_percent(accuracy = 1)) + 
  scale_x_continuous(name = "Metric Value") +
  facet_wrap(~variable, scales = "free_x")

biodiv_effects_plot / econ_effects_plot & theme(strip.text = element_text(size = 8))

```

## Discussion

Waiting on a fleshed out discussion until we are all happy with the content, but in general

* Growing need for more spatially explicit and ecosystem-based management and assessment

* `marlin` fills an important gap in the modeling literature
    * MPAs
    * Policy interaction
    * spatial-temporal MSE
    * Range shifts
    * make clear it's not just an MPA model

* Examples demonstrate this
    * total benefits and Distributional impacts in coral case
    * Something in the pelagic case
    * Complex interactions in the final case

* `marlin` and the methods within as a platform for future practical and academic bio-economic modeling needs

## References

::: {#refs}
:::

## scraps