### 2022-11-17 README.Rmd
updating makefile instructions

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Policy Implications of Coupled Economic and Ecological Dynamics in Aquatic Systems

Paper Introducing the `marlin` package.

## Reproducing results on UNIX machines

This repo is equipped with a `make` file. To reproduce results this way

1. Make sure you have make installed

2. Open a terminal and make sure the working directory of the terminal is set to the location of this repo

3. Make sure you have all dependencies installed by opening R and running `renv::restor()`

4. In the terminal, run `RUN_NAME:MyRunName FORMAT:MyFormat make` where MyRunName is the name of the folder you want to create to store results (e.g. v0.1) and FORMAT is one of pdf, html, docx

5. Wait for a bit! This command will run all the required scripts and knit the paper.