### 2023-01-09 README.Rmd
prepping repo for submission

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Policy Implications of Coupled Economic and Ecological Dynamics in Aquatic Systems

Paper Introducing the `marlin` package.

## renv

This repository is equipped with [`renv`](https://rstudio.github.io/renv/) for package management.

To install all dependencies for this project

1. Open R

2. Make sure you have `renv` installed (if not, `install.packages("renv")`)

3. Inside R, making sure the working directory is set to the root directory of this repository, run `renv::restore()`

4. Follow all prompts. This will install local versions of all the dependencies needed for this project

5. If you have trouble installing `marlin`, see instructions [here](https://danovando.github.io/marlin/#installation-troubleshooting)

## Navigating Repo

THe main scripts needed to reproduce results are

1. 00_setup.R
2. 01_coral_case_study.R
3. 02_blue_water_cast_study.R
4. 03_marlin_paper.qmd

The best way to reproduce the results is through the `make` commands described below. To reproduce results manually, first open 00_setup.R and set parameters as desired. Then, run 01_.. 02_... and 03_.

## Reproducing results on UNIX machines

This repo is equipped with a `make` file. To reproduce results this way

1. Make sure you have make installed

2. Open a terminal and make sure the working directory of the terminal is set to the location of this repo

3. Make sure you have all dependencies installed by opening R and running `renv::restore()`

4. In the terminal, run `RUN_NAME:MyRunName FORMAT:MyFormat make` where MyRunName is the name of the folder you want to create to store results (e.g. v0.1) and FORMAT is one of pdf, html, or docx

5. Wait for a bit! This command will run all the required scripts and knit the paper. Should take no more than a few hours.