### 2023-03-28 03_marlin_paper.qmd
passing back to abalone for a final top to bottom run

---
title: "Revealing Benefits, Costs, and Tradeoffs of Spatial Management in Marine Systems Arrising from Coupled Economic and Ecological Dynamics"
author:
    - Daniel Ovando
    - Darcy Bradley
    - Echelle Burns
    - Lennon Thomas
    - James Thorson
date: "Updated `r Sys.Date()`"
format:
  pdf:
    toc: true
    fig-pos: 'H'
  html: default
  docx: default
execute:
  echo: false
  message: false
  warning: false
  cache: false
bibliography: references.bib
params:
  run_name: ["v1.0"]
abstract: "Abstract: Designing effective spatial management strategies is challenging because marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond endogenously to ecosystem changes in ways that depend upon ecological and policy context. We present a tool, `marlin`, that can be used to efficiently simulate the bio-economic dynamics of marine systems in support of both management and research. We demonstrate `marlin’s` capabilities by focusing on two case studies on the conservation and food production impacts of marine protected areas (MPAs): a coastal coral reef and a pelagic tuna fishery. In the coastal coral reef example, we explore how heterogeneity in species distributions and fleet preferences can affect distributional outcomes of MPAs. In the pelagic case study, we show how our model can be used to assess the climate resilience of different MPA design strategies, as well as the climate sensitivity of different fishing fleets. This paper demonstrates how intermediate complexity simulation of coupled bio-economic dynamics can help communities predict and potentially manage trade-offs between conservation, fisheries yields, and distributional outcomes of management policies affected by spatial bio-economic dynamics."
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
---

```{r setup}
#| echo: false
#| message: false
#| include: false

source(here::here("00_setup.R"))

theme_set(theme_classic(base_size = 10, base_family = "Helvetica") + theme(strip.background = element_blank()))

if (!file.exists(here("data","sci_to_com.csv"))){
  sci_to_com <- taxize::sci2comm(unique(marlin_inputs$scientific_name),simplify= TRUE,db='ncbi') %>% 
  bind_rows(.id = "critter") %>% 
  pivot_longer(everything(),names_to = "critter",values_to = "common_name") %>% 
  mutate(common_name = stringr::str_to_title(common_name))

  write_csv(sci_to_com, file = here("data","sci_to_com.csv"))

} else{
 sci_to_com <-  read_csv(file = here("data","sci_to_com.csv"))

}

print(results_path)
results_files <- list.files(file.path(results_path))

rdses <- results_files[str_detect(results_files,'.rds')]

walk(rdses,~assign(str_remove(.x,".rds"), read_rds(file.path(results_path,.x)), envir = .GlobalEnv)) #load and assign all .rds objects in the appropriate results folder

# rdatas <- results_files[str_detect(results_files,'.Rdata')]
# 
# walk(rdatas, ~load(file.path(results_path,.x), envir = .GlobalEnv)) # load results into global environment

```

<!-- Single column, double line spaced -->

<!-- Within the word count (no more than 7000-8000 words for Standard Articles, 3000-4000 words for Applications and Practical Tools) -->

<!-- Continuous line and page numbering throughout -->

<!-- Clearly defined manuscript structure as standard: Author details, Abstract (must be numbered 1-4 according to Manuscript Specifications), Keywords, Introduction, Materials and Methods, Results, Discussion, Figures and Tables with captions -->

<!-- Figures and Tables can be embedded within the text where referenced to facilitate reviewing -->

<!--         Author Contributions (if >1 author – see Authorship Policy). -->

<!--         Statement of where you intend to archive your data -->

## Introduction

Communities around the world are increasingly looking to spatial strategies for managing marine ecosystems .
For example, the "30x30" movement calls for implementation of marine protected areas (MPAs), a form of spatial management, across 30% of the world's oceans by the year 2030 [@grorud-colvert2021 and references therein].
However , designing spatial management strategies to achieve desired objectives -- which may include recovery and resilience of overfished species, increased food production and economic well being, and equitable distribution of benefits - is not straightforward.
Different species have different resilience to fishing, distributions in space and time, and value to fishing fleets; fishing fleets themselves can have varying ranges of incentives and reactions to policy structures.
These dynamics pose challenges even for single-species assessment and management strategies , which are only amplified when we consider management policies specifically designed around multiple species, fleets, and spatial features in the oceans [@field2006a].

To illustrate, policies such as MPAs designed for both conservation and food production must consider factors such as the optimal size and placement of a protected area network given a wide array of life history, species distributions, exploitation levels, fleet dynamics, and policy-dependent behavior [@reimer2017], all of which may change over time, particularly given the impacts of climate change.
Efforts to effectively design spatial management strategies such as MPAs are further constrained by a lack of empirical evidence describing the size [@ovando2021a] and time required [@nickols2019] for MPAs to produce substantial effects across a range of social-ecological settings [@mcclanahan2021].

Many models have been developed to support the theory and design of spatial management strategies (see @fulton2015 for a review of existing methods).
However, these models have tended to be either highly complex and tactical models designed for use in a specific location, or to be extremely stylized representations intended to provide theoretical insights with less relevance for specific applications.
This lack of accessible models suited for more tactical application presents a challenge to stakeholders charged with marine spatial planning exercises.
It also presents a barrier to the scientific community, wherein comparisons across spatial management simulation studies are clouded by discrepancies in underlying model structures beyond differences in the core phenomenon in question.

To help address this challenge, we present a novel bio-economic modeling tool called [`marlin`](https://danovando.github.io/marlin/) that allows for efficient simulation of spatio-temporal biological and economic dynamics.
`marlin` allows users to simulate the impacts of marine management policies across a range of species targeted by various fleets across heterogeneous and dynamic seascapes.
This model can help users seek Pareto-optimal solutions that produce win-wins or minimize trade-offs across multiple management objectives, such as food production and conservation [@lester2013].

Here we present the core methods and functionality of `marlin`, and demonstrate its use in two case-study applications; a coral reef system and a pelagic system.
Our results show how interacting ecological, economic, and design attributes can affect the degree of benefits, costs, and trade offs between conservation and food production outcomes of MPAs.
More broadly, this paper both demonstrates the functionality of our model and the critical need for considering sufficiently realistic coupled ecological and economic dynamics in policy evaluations.

## Methods

`marlin` follows a "models of intermediate complexity" philosophy [@plagányi2014] targeted specifically towards capturing heterogeneous habitats, movement dynamics, and the simultaneous impacts of fishing fleets across multiple species (frequently called "technical interactions").
In its general form, `marlin` simulates the behavior of populations of spatially explicit biologically independent age-structured animal populations affected by one or more fishing fleets in time steps and a spatial resolution specified by the user.

The front end of the software package accompanying this paper, [`marlin`](https://danovando.github.io/marlin/), is written in R [@rcoreteam2021] to facilitate use, while the underlying population model is written in C++, integrated through the `Rcpp` package [@eddelbuettel2018], for speed.
Users on standard computers should be able to simulate one age-structured population distributed across a 10 by 10 patch system for 20 years in fractions of a second.

While `marlin` does not simulate species interactions, fishing fleets in `marlin` are capable of targeting and affecting multiple species simultaneously.
This allows the model to simulate processes such as fisheries bycatch or effort displacement in a way that accounts for how fleet behavior might affect multiple species in the system.
The ability to track the impacts of fishing fleets across multiple species simultaneously is particularly important as very few fishing fleets are truly single species.

We define a few commonly used terms here.
*Yields* refers to the volume of catch from fishing activity.
*Spawning stock biomass* refers to the total biomass of reproductively mature fish in the water, a function of numbers, weight, fecundity, and sexual maturity at age.
We measure the size of the population by th ratio of the *spawning stock biomass* (SSB) in a given time step relative to *unfished spawning stock biomass* (SSB0), SSB/SSB0.
So, a SSB/SSB0 value of 1 means that the population is unfished, a SSB/SSB0 value of 0 means the population is extinct.

As a demonstration of the use of this modeling tool, we explore two general situations

1.  Trade-offs and distributional outcomes for food production and multi-species biomass in a coastal coral-reef fishery
2.  Implications of climate-driven range shifts for MPA design in a pelagic fishery

We chose these two examples to illustrate the use of `marlin` in contrasting systems in which spatial management strategies such as MPAs are increasingly considered.

Below we provide a summary of the `marlin` model, as well as details of the case studies.
We note that `marlin` is a complex software package with many options to control detailed behaviors.
Given that these specific features will evolve over time, we do not describe in detail every option of the model here.
Rather, we focus on the equations and methods describing the core foundations of our tool for efficiently simulating bio-economic dynamics in aquatic systems.

### Model Summary

`marlin` simulates the dynamics of one or more species, currently best representing fish-like species, using age-structured population dynamics.
Each time step, species move throughout the simulated seascape using both diffusion and "taxis" (active movement towards preferred habitat), experience natural and potentially fishing mortality, and if in a spawning season spawn using one of the possible forms of density dependence implemented in the model, with the potential to include auto-correlated stochastic deviations in the amount of offspring produced, generally called "recruitment deviates".

These species can be caught by fishing fleets.
A fishing fleet is defined in `marlin` by a set of fishers that have the same fishing skill, prices, and selectivity for individual species (each of which we denote as a *metier*).
For example, both a longline and purse seine fleet may capture bigeye (*Thunnus obesus*) and skipjack tunas, but the longline fleet can be made much more likely to catch larger bigeye than skipjack, and *vice versa*.
Each fishing fleet distributes its fishing effort in space according to a specified spatial allocation rule (see *Spatial Allocation of Effort* ), for example in proportion to profit-per-unit-effort, conditional on management policies in places such as quotas and/or the presence of any spatial restrictions such as MPAs.

Unconstrained by management, the total amount of effort exerted by each fleet can follow one of two dynamics: open access or constant effort (see *Calculating Total Effort*).
Under constant effort, the total amount of effort of each fleet is fixed over time, with the possible exception of attrition due to MPA placement.
Under open access, the total amount of effort in a given time step is a function of the profitability of the fleet in the previous time step, evolving until a bionomic equilibrium of zero total profits is reached.
Profitability is a function of the volume and price of each species caught, as well as the cost of the total amount of fishing effort per fleet and the travel costs as a function of distance from a port (see *Fleet Dynamics*).

The dynamics of each fleet can be modified by management in a variety of ways.
Managers can impose size limits for individual species within each fleet.
They can also impose total catch quotas for each species in each time step.
When quotas are activated, if the total catch across all fleets for a given species under the model's effort dynamics would exceed the allowable quota for that species, the total effort for each fleet is decreased proportional to its contribution to the total catch until the quota is satisfied.
As an alternative to catch quotas, managers can set an effort cap per fleet, which prevents effort from exceeding a given amount under open access dynamics, though effort may be reduced below this cap if required by the profit equation (i.e. the fleet can choose to fish less than the quota or effort cap).

`marlin` also allows users to specify closed fishing seasons for one or more species in the system.
When all species in a simulation share the same closed seasons, no fishing occurs in those seasons.
When not all species share the same closed seasons, fishing still occurs in the closed seasons, but the fleet avoids catching the species for which fishing is closed in a given season, assuming no discard mortality.
When closed seasons are in effect, the amount of effort in the next open season is a function of the effort and profits in the last open fishing season.
Lastly, managers can specify locations of no-take MPAs, which can change in size and location if needed.
When MPAs are implemented, fishing effort that used to operate inside the MPAs can either leave the fishery, or be redistributed outside the MPA (the default behavior).

Users can use the model to explore general properties of spatially explicit social-ecological systems, or can parameterize it to explore key uncertainties of more tactical applications.
As an example of the later case, we provide two case studies here designed to reflect the kinds of questions stakeholders might face if designing an MPA network for a coastal coral reef or a pelagic system.

### Fleet Dynamics

Each fishing fleet (*f*) generates catches or yield (*Y*), revenues (*R*), and costs (*C*) from fishing individual species (*s*) that it targets in a given time step (*t*) and patch (*p*).
The fleet then makes decisions around fishing locations and intensity, subject to regulatory constraints, based on the total profitability across all species in space and time.

Revenues for each fleet in a time step are a function of the total amount of each species caught and its price ($\Pi$).
The amount caught is a function of the selectivity at age of each species for each fleet ($\alpha$), the fishing efficiency of the fleet for that that species (also called "catchability", *q*), the amount of fishing effort of the fleet in question in that patch in a given time step (*E*), and the total instantaneous fishing mortality (*u*) at age (*a*) (including all other fleets) for that particular species in that patch and time step.

$$
 R_{t,p,f} = \sum_s^{N_s} \sum_a^{N_a} \Pi_{s,f}\frac{\alpha_{a,s,f}q_{s,f}E_{t,p,f}}{u_{t,p,s,a}} \times Y_{t,p,s,a}
$$ {#eq-revenue}

Total catches or yield *Y* are calculated through the Baranov equation [@baranov1918], which appropriately accounts for the total instantaneous mortality (both fishing and natural. *z*) and divides the total amount of the population biomass (*b*) killed between natural (*m*) and fishing (*u*) sources, with the amount of biomass killed through fishing called "yield".

$$
Y_{t,p,s,a} = \frac{u_{t,p,s,a}}{z_{t,p,s,a}} \times b_{t,p,s,a} \times (1 - e^{-z_{t,p,s,a}})
$$ {#eq-yield}

$$
u_{t,p,s,a} = \sum_{f=1}^{N_f}\alpha_{a,s,f}q_{s,f}E_{t,p,f}
$$ {#eq-u}

Selectivity at age $\alpha$ is modeled through either a logistic form or a dome-shaped form.
The logistic form is based on the lengths *l* at which 50% of individuals are selected by the fishing gear, $l^{sel}$, and $\delta$ which is the difference between the length at 50% selectivity and the length at 95% selectivity.

$$
\alpha_{a,s,f}=\frac{1}{(1 + e^{-log(19)\times\frac{l_{a,s} - l^{sel}_{s,f}}{\delta^{sel}_{s,f}}})}
$$ {#eq-selectivity}

We approximate the dome-shaped form as a normal distribution with mean $l^{sel}$ and standard deviation $\sigma$.
The normal density function is re-scaled such that the selectivity is 1 at $l^{sel}$.

$$
\alpha_{a,s,f}= \frac{1}{\sigma_{s,f} \sqrt{2\pi}}e^{-0.5 \left( \frac{l_{a,s} - l^{sel}}{\sigma_{s,f}}  \right)^2}
$$ {#eq-dome}

Lastly *z* is total mortality, the sum of fishing mortality and natural mortality (*m*).

$$
 z_{t,p,s,a} = u_{t,p,s,a} + m_{s,a}
$$ {#eq-z}

This means that each species experiences a total mortality at age in a time step in a patch, individual fractions of which are portioned off as catches and subsequently revenues for each fleet.
Given revenues *R*, we then calculate profits $\phi$ based on revenues and costs.
Costs *C* are calculated as a function of base costs per unit effort ($\gamma$) as well as potential additional costs per unit effort of fishing in particular patches ($\eta$) for fleet *f*.
$\beta$ allows for the cost per unit effort effort to scale non-linearly.
Travel costs ($\eta$) are calculated based on the Euclidean distance of each patch to the nearest port of a given fleet, and users can specify any number from zero to the number of patches of port locations.
When no ports are specified, travel costs are zero.

$$
\phi_{t,p,f} = R_{t,p,f} - C_{t,p,f}
$$ {#eq-profits}

$$
C_{t,p,f} = \gamma_f  \left( E_{t,p,f}^{\beta_f} + \eta_{f,p} E_{t,p,f} \right)
$$ {#eq-travel-costs}

#### Calculating Total Effort

`marlin` allows for two general modes of effort.
The simplest is "constant effort", in which the total effort of each fishing fleet remains constant over time.

The more complex option is "open access".
Under open access, the total amount of effort in a given time step for fleet *f* is a function of the profitability of that fleet in the proceeding time steps in which fishing was open, where $\theta$ controls the responsiveness of fleet *f* to the log of the ratio of total revenues *R* to total costs *C*, and approximates the proportional change in effort in response to a one unit changes in the log revenue to cost ratio.

$$
E_{t+1,f} = E_{t,f} \times e^{\theta_f log(R_{t,f} / C_{t,f})} 
$$ {#eq-oa}

This is essentially a Gompertz model for fishing effort, that has been used for other theoretical studies of fishing effort dynamics [@thorson2013].

#### Tuning Fishing Fleet Parameters

The degree of fishing pressure exerted by a given set of fishing fleets on each species is a function of a range of parameters including the total amount of effort (*E*), the selectivity ogives ($\alpha$,the curve specifying the proportion of individuals at a given age or size that will be caught by a unit of fishing mortality), fishing cost (*C*), the spatial distribution of the species affected by the fleet, the relative prices across species ($\Pi$), and the catchability coefficient of each fleet for each species ($q$).
`marlin` provides a tuning option to help users achieve desired biological outcomes from their fleets.

Users can tune their fleets in one of two ways.
First, they can specify a target exploitation rate *u* for each species in their simulation.
Taking all the other parameters of the model as given, `marlin` then adjusts the catchability coefficients $q_{f,s}$ for each fleet *f* and species *s* such that the equilibrium exploitation rate for each species matches the desired level.
Second, they can specify a target total spawning stock biomass under fishing divided by total unfished spawning stock biomass, and the model will adjust the catchability coefficients $q_{f,s}$ for each fleet *f* and species *s* such that the equilibrium ratio of fished to unfished spawning biomass for each species matches the desired level.

Users can also tune the fleet dynamics by specifying the ratio of costs to revenues.
Price data for use in the model can be obtained relatively simply through literature reviews, market surveys, or local experts, so long as they are converted to the same units of biomass used in the model.
However, cost parameters are more complicated, as translating say cost per day of fishing into the same representation of fishing effort used in the model is not straight forward to accomplish.
As an alternative, users can specify an equilibrium cost-to-revenue ratio for the fleet, essentially the profit margins of the fleet in question, and `marlin` will tune the cost parameter to achieve this desired cost to revenue ratio given the other parameters in the model (e.g. price and biomass of the target species).
So for example, while users may not have access to exact cost parameters, they may have a general sense that when operating under ideal conditions a given fleet has a profit margin of 10%, i.e. the cost to revenue ratio is 90%.
They can input the cost to revenue ratio as 0.9 then and allow `marlin` to tune the cost parameters accordingly.

#### Spatial Allocation of Effort

Each fishing fleet decides how to allocate its effort in space based on one of four possible spatial allocation strategies.
The ideal free distribution (IFD) is the standard method for distributing fishing fleets in spaces (see @gillis2003 and references therein).
However, analytical solutions to the IFD present a number of complications for our model.
Our model is capable of tracking multiple fleets at once.
This means that the IFD for a single fleet would commonly be modeled as a Nash-equilibrium based on the actions of each of the individual fleet conditional on the actions of all other fleets.
While possible to implement, this would slow down our model runs to the point of practically preventing large-scale evaluation of the spatial management policies.

As such we explored a series of "next best" fleet distribution algorithms.
While not the IFD in any individual time step, over time they start to approximate the IFD, as each assumes that fleets base their decisions for the current time step on the outcomes in the prior timestep, meaning that the impacts of the actions of other fishing fleets are somewhat internalized.
The four possible fleet distribution algorithms are

1.  Revenue per unit effort (RPUE): The fleet distributes itself in space based on the realized revenue per unit effort in each patch in the preceding time step
2.  Revenue: The fleet distributes itself in space based on the realized total revenue in each patch in the preceding time step
3.  Profit per unit effort (PPUE): The fleet distributes itself in space based on the realized profit per unit effort in each patch in the preceding time step
4.  Profit: The fleet distributes itself in space based on profit in each patch in the preceding time step

Revenue based spatial distribution is not likely to be very realistic; in general we would expect fishing fleets to respond to profits on some core level.
However, due to the complexity of parameterizing cost functions, fleet dynamics are often evaluated based on yield or revenue alone, and so we include those scenarios here to allow users to evaluate the potential implications of this choice.
The decision on whether to allocate the fleet based on absolute or relative (per unit effort) metrics is more complex.
When effort represents the actions of separate and individual fishing actors (e.g. independent fishing vessels), a per-unit-effort strategy, in which fishers distribute themselves based on the expected catch of their individual efforts, may be more realistic [@hilborn1987].
Conversely, a system defined by a sole owner seeking to maximize total profits might be better represented by a fleet model based on total profits

### Population Model

The underlying population model used is an age structured single-species model in the manner of @ovando2021a.
The population model requires many parameters.
However, if the user supplies either a scientific or common name for the species in question (scientific preferred), the model will supply default values for that species based on the values reported in FishLife [internet connection required, @thorson2020].
FishLife provides estimates of core life history parameters for fish species based on a model integrating published values and evolutionary connections.
FishLife provides reasonable default values for given species, but these default values should be not be taken as definitive and users should input best available parameters specific to the stock in question if they wish to best represent the dynamics of their specific system, and should check the plausibility of the default values prior to use.
All of the life history values can be replaced with user-supplied values as needed and should be for more tactical applications; the default values simply ensure that reasonable and internally consistent values for the species in question are supplied as a starting point.

#### Movement

`marlin`'s movement dynamics are based on a continuous-time Markov chain (CTMC), as described in @thorson2021a.
Within this framework, the model allows for movement to be broken down into three components of *advection* (drifting with currents), *taxis* (active movement towards preferred habitat), and *diffusion* (essentially remaining variation in movement not explained by advection or taxis).
For now, `marlin` focuses just on the diffusion and taxis components of this model, assuming that advection is zero, though future extensions could easily incorporate advection vectors from oceanographic models.
In this way, `marlin` allows users to run anything from a simple Gaussian dispersal kernel up to a system governed by species that passively diffuse out from a core habitat defined by a dynamic thermal range.

We provide a brief overview of the the general CTMC method here, see @thorson2021a for a detailed description.
Under this framework, movement from each patch to each other patch in the system in a given timestep *t* for life stage *a* of species *s* is defined by a movement matrix $\boldsymbol{M}_{t,s,a}$.
$\boldsymbol{M}_{t,s,a}$ is calculated as the matrix exponent of a function of the diffusion $\boldsymbol{D}$ and taxis $\boldsymbol{\tau}$ matrices scaled by the width of the time step $\Delta_{t}$ and the length of the edge of each patch $\Delta_d$ that the model is running under.
So for example, if we are running a model on a monthly timestep, and we define the units of time in years, then $\Delta_{t} = 1/12$.
Similarly, if we are running a model where one patch equals 100km^2^, then $\Delta_d = 10km$.
This parameterization allows users to set the effective area of the spatial domain through two avenues; the number of patches, which effectively scales the resolution of the model, and the area of each patch, which scales the spatial extent of the simulation.

$$
\boldsymbol{M}_{t,s,a} = e^{\Delta_{t}(\boldsymbol{D}_{t,s,a} + \boldsymbol{\tau}_{t,s,a})} 
$$ {#eq-movement}

The individual components (*m*) of the movement matrix are filled based on an adjacency matrix, which defines whether two patches are both adjacent and water (as opposed to land or another physical barrier), a diffusion rate $D$ defined in units of area of a patch per unit of time, and a habitat preference function *h* in units of length of a side of a patch per unit time.
For example, if we are defining the time units as years and the distance units as kilometers, for a tuna $D$ might be 1,000 $\frac{KM^2}{Year}$.
We then use parameters $\Delta_{t}$ and $\Delta_{d}$ parameters to translate the diffusion rate $D$ to match the time step and patch size used in a simulation.
This "scale free" parameterization means that appropriate value of $D$ can be identified for a species and then set, regardless of the time step or patch size used in the simulation model itself.
The taxis component of the movement process is a function of the difference in habitat quality *h*.
The habitat preference function itself can take any form the user wishes.
Exponentiating the difference in the habitat preference function between patches turns the taxis matrix into a multiplier of the diffusion rate *D*.
As such, when creating habitat layers for simulation, users can tune the scale of the habitat gradient function to result in realistic multipliers of the diffusion rate.
This parameterization ensure that the off-diagonal elements of the movement matrix $\boldsymbol{M}_{t,s,a}$ are all non-negative, a requirement of the CTMC method.

$$ 
m_{p1,p2,t,s,a} = \begin{cases}
      = \frac{\Delta_{t}}{\Delta_{d}^2}De^{\frac{\Delta_t(h(p2,t,s,a) - h(p1,t,s,a))}{\Delta_d}} & \text{if p2 and p1 are adjacent}\\
     = -\sum_{p' \neq p1} m_{p1,p2,t,s,a} & \text{if p1 = p2}\\
     = 0 & \text{otherwise.}
\end{cases}
$$ {#eq-diffusion}

For both the diffusion and taxis matrices, we allow for the inclusion of physical barriers to movement (i.e. land).
Pairs of patches that are adjacent but in which one or both patches are a barrier to movement are set as non-adjacent.
The CTMC model then produces movement dynamics that move around barriers rather than over them.
This is different than setting habitat in a patch to zero; an animal will not preferentially move towards a patch with zero habitat all else being equal, but can move over a patch with zero habitat towards another better patch.
An example might be a species that lives and feeds on coastal habitats transiting through a patch of open ocean to reach a new feeding ground.
In contrast, a physical barrier represents something like a peninsula that an animal must move around to transit from one side to another.

While this CTMC approach to movement simulation is to date not commonly seen in the marine modeling literature, it has numerous advantages that warrant its seeming complexity.
First, when only diffusion is present, the model will generalize to the familiar dynamics of a Gaussian dispersal kernel.
However, the time component *t* in @eq-movement allows the movement model to be run on arbitrarily fine time steps.
Second, the taxis model allow for more active habitat choices by species, allowing us to simulate preferences of species in space and time efficiently.
Lastly, this form has the advantage that the parameters are directly estimable from real data.
So, if provided with for example spatial abundance data and a tagging study, users can estimate the diffusion and taxis movement parameters in the same manner as @thorson2021a, and then pass their estimated parameters to `marlin` for simulation (so long as the estimating method uses the same functional form as the movement model in `marlin`).

#### Population Growth

For the population model, numbers *n* at time *t* for age *a* are a function of growth, death, and recruitment

$$
n_{t,p,s,a} = \begin{cases}
      = BH(ssb_{t-1,p,s,a}) & \text{if $a = 1$}\\
     = n_{t-1,p,s,a-1}e^{-(z_{t-1,p,s,a-1}}, & \text{if $1< a < max(age)$}\\
     =  n_{t-1,p,s,a}e^{-(z_{t-1,p,s,a})} + n_{t-1,a-1}e^{-(z_{t-1,p,s,a-1})}, & \text{if $a = max(a)$}
  \end{cases}
$$ {#eq-pop}

where *BH* is the Beverton-Holt recruitment function [@beverton1957] and *ssb* is spawning-stock-biomass.
Per convention, the model allows for a "plus group", wherein rather than tracking numbers of every possible age, individuals greater than or equal to a given maximum age are grouped together.

Spawning stock biomass *ssb* is calculated by converting age to mean length, calculating weight at age, maturity at age, and then calculating spawning stock biomass as the sum of spawning potential at age in a given time step, taking into account the potential for hyperallometry in the manner of @marshall2021.
Age is converted to length through the von Bertalanffy growth equation given parameters asymptotic length ($l_\infty$), growth (*k*) and theoretical age at length zero ($a0$) assuming log-normally distributed variation *u* in the length at age with CV $\sigma_s$.

$$
  l_{a,s} = l_{\infty,s}\left(1 - e^{-k_s(a - a0_{s})}\right)e^{u_s}
$$ {#eq-len}

$$
u_s \sim N(0,\sigma_s)
$$ {#eq-sigma}

Biomass *b* at age is then given by the weight at length equation governed by a scaling coefficient $\Omega_s$ and an exponent $\Phi_s$ that controls the rate at which volume scales with length

$$
  b_{a,s} = _s \times l_{a,s}^{wb_s}
$$ {#eq-weight}

The proportion of sexually mature individuals (*mat*) at a given age is then calculated as a logistic function where $l_{mat}$ is the length at which on average 50% of individuals are sexually mature, and $\delta_{mat}$ is the unit of length beyond $l_{mat}$ at which on average 95% of fish are sexually mature.

$$
  mat_{a,s} = \frac{1}{\left(1 + e^{-log(19)\times\frac{l_{a,s} - lmat_s}{{\delta}mat_s}}\right)}
$$ {#eq-mat)}

Spawning stock biomass at time *t* is then calculated as a function of the numbers at age, the maturity at age, and the weight at age raised by a parameter $\gamma$.
When $\gamma$ is greater than 1, the species is said to experience hyperallometric fecundity, i.e. fecundity increases faster than weight.

$$
  ssb_{t,p,s} = \sum_{a=1}^{N_a}w_{a,s,t}^{\gamma_s} mat_{t,a} n_{t,p,s,a}
$$ {#eq-ssb}

#### Recruitment

Recruitment (i.e. the number of age 1 individuals entering the population) follows Beverton-Holt dynamics parameterized around steepness (*h*) with log-normally distributed recruitment deviates $\epsilon$.
When steepness is one recruitment is independent of spawning biomass.
As steepness approaches 0.2 recruitment becomes a linear function of spawning biomass.
We allow for five variants in the timing of density dependent recruitment, building off of @babcock2011 :

1.  Global density dependence: Density dependent recruitment is a function of the sum of spawning biomass across all patches, and recruits are then distributed according to habitat quality

$$
n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p,s}}\times{h_s}\times\sum_{p=1}^P{ssb_{t-1,p,s}}}{0.2\times{\sum_{p=1}^Pssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p,s}}}\right)\times {r0_{p,s} / \sum_p^P{r0_{p,s}}\times \epsilon_{t,s}
}
$$ {#eq-dd2}

where $r0$ is is a vector of recruits under unfished conditions in a given patch.

2.  Local density dependence: Density dependent recruitment occurs independently in each patch and recruits are retained in their home patch.

$$
n_{t,p,s,a = 1} = \left(\frac{0.8{\times}r0_{p,s}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p,s}}}\right) \times \epsilon_{t,s}
$$ {#eq-dd1}

3.  Local density dependence then disperse: Density dependent recruitment occurs independently in each patch and recruits are then dispersed.

$$
  n_{t,p,s,a = 1} = \left(\frac{0.8{\times}r0_{p,s}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p,s}}}\right)\times \boldsymbol{d^l_s} \times \epsilon_{t,s}
$$ {#eq-dd3}

where **d^l^** is the recruitment movement matrix

4.  Post-dispersal density dependence: Larvae are distributed throughout the system, and then density dependent recruitment occurs based on the density of spawning biomass at the destination patch.

$$
  larv_{t,p,s} = ssb_{t-1,p,s} \times\boldsymbol{d^l_s}
$$ {#eq-larvmove}

$$
n_{t,a = 1,p,s} = \left(\frac{0.8{\times}r0_{p,s}\times{h_s}\times{larv_{t,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{larv_{t,p,s}}}\right) \times \epsilon_{t,s}
$$ {#eq-dd4}

5.  Global density dependence allocated by spawning biomass: Density dependence is a function of the sum of spawning biomass across all patches, and recruits are then distributed according to the distribution of spawning biomass

$$
n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p,s}}\times{h_s}\times\sum_{p=1}^P{ssb_{t-1,p,s}}}{0.2\times{\sum_{p=1}^Pssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p,s}}}\right)\times \frac{ssb_{t-1,p,s}}{\sum_{p=1}^Pssb_{t-1,p,s}} \times \epsilon_{t,s}
$$ {#eq-dd5}

Log-normal recruitment deviates are calculated with the potential for autocorrelation defined with strength $\rho$

$$ 
log(\epsilon_{t,s}) \sim \begin{cases}
       N(0,\sigma_{r,s}), & \text{if $t = 1$}\\
       \rho_s log(\epsilon_{t-1,s}) + \sqrt{1 - \rho_s^2} N(0,\sigma_{r,s}), & \text{if $t > 1$}
  \end{cases}
$$ {#eq-recdev}

And log recruitment deviates are converted to raw units using the bias correction factor

$$
\epsilon_{t,s} = e^{log(\epsilon_{t,s}) - \sigma_{r,s}^2/2}
$$ {#eq-bias}

#### Reference Points

Fisheries management is often concerned with measuring stock status relative to maximum sustainable yield (MSY) based reference points, though the exact level of stock status relative to MSY reference points desired by management or societal objectives may vary widely.
MSY based reference points present a problem for a multi-fleet and spatial-temporal model such as `marlin`.
MSY and the fishing mortality rate that would produce MSY, $F_{MSY}$, are a function of fishery selectivity.
Fishery selectivity in this model can vary by fleet, and species can be distributed unevenly in both space and time.
This means that the net effective fishing selectivity on a species can vary depending on the dynamics at a given moment, making the definition of an equilibrium concept such as MSY challenging.

As such, we do not report MSY based reference points in the model by default, with the exception of the spawning stock biomass at MSY ($SSB_{MSY}$) which tends to be conserved across selectivity patterns as it is primarily a function of the mortality and flow of recruits relative to the maturity ogive.
There are many different strategies for estimating reference points in spatially explicit systems [@kapur2021a].
We leave it to users to define and find relevant reference point as required by their specific needs.

<!-- ### MPA Optimization -->

<!-- Our MPA optimization algorithm seems to create an MPA network that -->

<!-- maximizes an objective function, given by -->

<!-- $$ -->

<!-- obj_p = \Phi \times scale(\sum_f^F \boldsymbol{\phi_{f,p}}) + (1 - \Phi)  scale(\boldsymbol{\sum_s^S\frac{ssb_s}{ssb0_{s,p}}}) -->

<!-- $$ {#eq-obj} -->

<!-- Where *scale* rescales the values to be between 0 and 1 such that the -->

<!-- weight parameter is comparable between vectors. The algorithm works by -->

<!-- first assigning every patch in the system to an MPA. For the first -->

<!-- round, the algorithm then calculates the equilibrium objective function -->

<!-- value resulting from sequentially removing each individual patch from -->

<!-- the MPA network. The cell whose removal from the MPA produces the -->

<!-- highest objective function value is then removed from the MPA network, -->

<!-- and the marginal value of the objective function each of the remaining -->

<!-- cells from the last steps are then stored. For the next steps, we -->

<!-- sampled 25% of the remaining MPA cells with probability proportional to -->

<!-- their marginal contribution to the objective function in the prior -->

<!-- iteration, selecting the cell whose removal produces the highest total -->

<!-- objective function as the cell to remove from the network, and updating -->

<!-- the marginal value of the sampled-but-not-selected cells. This in -->

<!-- essence is a sampling-importance-resampling algorithm. The algorithm -->

<!-- continues until no more cells are left in the MPA. This then provides -->

<!-- for a given objective weighting value $\Phi$ the near-optimal MPA -->

<!-- network at every MPA size from 0 to 100%. We say "near" as they is some -->

<!-- path dependency here, in that for example the network that maximizes the -->

<!-- objective function with 40% of the MPAs does not necessarily contain the -->

<!-- optimal network at 30%. However, the reversed nature of the algorithm, -->

<!-- starting at 100% MPA coverage and working better, produces a more stable -->

<!-- network with regards to this problem than starting at 0% and growing to -->

<!-- 100%. -->

### Case Studies

We include two examples demonstrating how `marlin` can be used to support marine spatial planning.
In the first, we show how `marlin` can be used to compare the total and distributional impacts of MPAs designed in a heavily fished coastal coral reef ecosystem.
In the second, we demonstrate how `marlin` can be used to assess components of climate resilience of alternative MPA design strategies in a pelagic system.

Each of the case studies contains too many parameters and options to be succinctly presented in the text here, and we are concerned that attempting to duplicate all of the parameters of the model within the text would simply introduce errors and confusion.
As such, and since these results are for illustrative purposes only, readers should consult the accompanying code to view the precise details of each simulation.
Targeted tactical applications must carefully consider and document all decisions made around model parameters.

#### MPA Design Strategies

We make use of three potential "rule of thumb" MPA design strategies:

1.  *Rate*: MPAs are placed based on the pre-MPA SSB/SSB0 weighted catch relative to the total catch in a patch.
    So, patches with high rates of catch of depleted species relative to total catch are prioritized.

2.  *Target Fishing*: MPAs locations are prioritized proportional to fishery catches.
    Patches with high total catches are prioritized over patches with low catches.

3.  *Spawning Aggregation*: MPAs are centered on the location of a known spawning aggregation

To our knowledge, MPAs are never randomly placed; their locations are decided based on stakeholder negotiation given available information and objectives.
In theory, the design of MPA networks can be optimized through the use of a modeling framework, and depending on the validity of the model, this process is likely to produce better outcomes than manually-designed strategies [@rassweiler2012; @rassweiler2014].
However, designing optimal MPA networks becomes increasingly difficult as the range of objectives and the complexity of the model increase.
Therefore, while `marlin` provides some support for [optimization](https://danovando.github.io/marlin/articles/optimize-mpas.html), we focus here on the design and performance of these more "rule of thumb" design strategies that may be more accessible to a wider range of users.
We allow all MPAs to be designed in a "mosaic" fashion in these examples, but users can easily extend the analysis to compare outcomes between "contiguous" (MPA is made up of one continuously connected block) and "mosaic" (MPAs can be separated in space) MPA designs [@pons2022] by developing and passing their own MPA design algorithms to the model.

#### Coastal Coral Reef Fishery

In our coastal coral reef example, we model the dynamics of four tropical Pacific species: a grouper (*Epinephelus fuscoguttatus*), a shallow-reef snapper (*Lutjanus malabaricus*), a deep-water snapper (*Pristipomoides filamentosus*), and a reef shark (*Carcharhinus amblyrhynchos*).
The simulated groupers undergo a mass migration to a spawning aggregation once per year, followed by the sharks.
Shallow-reef snappers stay in reefs closer to shore above a steep drop-off year-round, while deep-water snappers stay in the deeper reefs past the drop-off (@fig-coral).

These species are targeted by two different fleets.
Fleet One primarily targets the grouper and near-shore snapper populations, but will land any incidentally captured sharks.
Fleet One has a logistic selectivity pattern for all species, as they retain any fish caught for consumption or sale.
both grouper and shallow-reef snapper due to demand for "plate sized" fish , and logistic selectivity for sharks, as any shark caught is retained.
Fleet One is totally dependent on fishing for their livelihood, meaning the local community takes advantage of every possible opportunity to fish, and as such we model it as a "constant effort" fishery.
Do to having less efficient boats, Fleet One has a higher cost per distance coefficient than Fleet Two.
Fleet One's home port is located on the site of the grouper spawning aggregation.

Fleet Two is a more commercial fleet that primarily targets the deep-water snapper population.
This fleet primarily sells their catch to local restaurants and distributors where plate-sized fish are prized, and so for both snapper and grouper Fleet Two has a dome-shaped selectivity pattern [@kindsvater2017].
While plate-sized deep snapper are the primary target of Fleet Two, we model Fleet Two's selectivity for deep snapper as logistic due to high levels of discard mortality for deep-water snapper resulting from barotrauma.
Fleet Two catches groupers, though less than Fleet One, and receives no price for sharks due to the requirements of a certification program through which they sell their deep-water snapper.
Accidental captures (bycatch) of sharks do occur, which results in mortality.
Fleet Two operates under open-access dynamics, as fishing is not the only means of subsistence for this community; short-term effort expands and contracts in response to profitability of the primarily grouper-driven fishery.
Fleet two's home port is located further north from Fleet One.

We use `marlin` to simulate the outcomes for both food production and conservation for each of the species and both of the fleets as a function of both MPA size and MPA design strategy.
For this exercise, MPAs are placed with perfect information and have no design constraints for continuity (i.e. the selected MPA design can be as patchwork as dictated by the data).
We ran the simulation on a quarterly level ($\Delta_t = 1/4$) and set the area of each patch to be 5km^2^ ($\Delta_p = \sqrt{5}$), using 400 patches for a total area of 2,000 KM^2^.

```{r}
#| label: fig-coral
#| fig-cap: "Distribution of fish spawning biomass (cell color, yellow = higher, blue = lower) each season (columns) under unfished conditions. The x-axis represents longitude, y-axis latitude. Numbers show the approximate location of each fleet's (one or two) port. Columns indicate quarterly seasons that repeat each year."

coral_fauna <- coral_fauna_and_fleets$fauna

coral_fleets <- coral_fauna_and_fleets$fleets

resolution <- sqrt((coral_fauna[[1]]$patches))

grid <- expand_grid(x = 1:resolution, y= 1:resolution) %>%
  mutate(patch = 1:nrow(.))

mpa_locations <- expand_grid(x = 1:resolution, y = 1:resolution) %>%
  mutate(mpa = TRUE)

coral_sim <- simmar(
  fauna = coral_fauna,
  fleets = coral_fleets,
  manager = list(mpas = list(
    locations = mpa_locations,
    mpa_year = 1
  )),
  years = 2
)

patch_biomass <-
  map_df(coral_sim, ~ map_df(.x, ~tibble(biomass = rowSums(.x$ssb_p_a), patch = 1:nrow(.x$ssb_p_a)), .id = "critter"), .id = "step") %>%
  mutate(step = as.numeric(step)) %>%
  left_join(grid, by = "patch")

 patch_biomass %>%
  group_by(critter, step) %>%
  mutate(sbiomass = biomass / max(biomass)) %>%
  ungroup() %>%
  filter(between(step,1,2)) %>%
  mutate(step = fct_recode(as.factor(step), "2" = "1.25", "3" = "1.5", "4" = "1.75")) %>% 
  mutate(Season = step) %>% 
  ggplot()+
  geom_tile(aes(x,y, fill = sbiomass),show.legend = FALSE) +
  geom_text_repel(data = coral_ports, aes(x, y, label = fleet), color = "red") +
  facet_grid(critter~Season, labeller = labeller(critter = titler, Season = label_both)) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_viridis_c() + 
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank())

```

#### Pelagic Systems

```{r}
blue_fauna <- blue_water_fauna_and_fleets$fauna

blue_fleets <- blue_water_fauna_and_fleets$fleets

blue_resolution <- sqrt(blue_fauna[[1]]$patches)

```

We model our pelagic use case loosely on the characteristics of the Western and Central Pacific Ocean (WCPO) tuna fisheries.
We simulate trajectories of `r n_distinct(names(blue_fauna))` species commonly caught in the region, including both the highly abundant skipjack tuna and the heavily depleted oceanic whitetip shark (*Carcharhinus longimanus*) (@fig-blue-ssb).
We use publicly available data on catch-per-unit-effort of each of these species from the WCPO as a very rough proxy for baseline habitat distributions, noting that where possible, fishery-independent abundance indices would be preferable (@fig-blue).

These pelagic species are caught by a longline fleet that primarily targets large adult tunas such as bigeye and yellowfin for high-grade consumption, and a purse seine fishery that primarily targets skipjack tunas for bulk canning.
Selectivities were modeled as logistic for the longline fleet, and dome-shaped for the purse seine fleet, with selectivity values set to broadly match the patterns reported in relevant [stock assessments](https://oceanfish.spc.int/en/ofpsection/sam/sam/210-stock-assessment-input-and-result-files-) for each species.
Both fleets operate under open-access dynamics with an effort cap.
The effort cap was set at the level of effort that resulted in the desired levels of SSB/SSB0 for each species under open-access dynamics (@fig-blue), intended to simulate a scenario where managers step in to prevent further expansion of fishing effort in a fully developed fishery.
For forward-simulation, open-access dynamics can result in effort decreasing in response to profitability, but cannot result in effort beyond the effort cap set for each fleet.

For this exercise, we focused on using `marlin` to assess resilience of the selected Target Fishing and Rate "rule of thumb" MPA design strategies to a climate-driven range shift.
Specifically, we simulate an extreme example where the centroid of each population shifts northward at a rate of 0.33 patches per year, over a 20 year time horizon (@fig-blue-ssb).
We designed MPA networks given the conditions in the starting year, and then held that network constant over the years of the experiment, running one simulation with and another without the climate-drive range shift.
We then compared the effects of this range shift on food production and conservation outcomes from MPA networks designed based on the pre-range shift world.
We ran the pelagic simulation at a quarterly level ($\Delta_t$ = 1/4 year), and set the area of each cell to be roughly 97,000 KM^2^ across 144 patches each with a side length of roughly 311 KM, for a total area of 14,000,000 KM^2^, broadly commensurate with the area of the WCPO.

```{r}
#| label: fig-blue-ssb
#| fig-cap: Spawning biomass divided by unfished spawning biomass in the time period prior to implementation of MPAs for the pelagic case study."

grid <- expand_grid(x = 1:blue_resolution, y= 1:blue_resolution) %>%
  mutate(patch = 1:nrow(.))

mpa_locations <- expand_grid(x = 1:blue_resolution, y = 1:blue_resolution) %>%
  mutate(mpa = TRUE)

blue_sim <- simmar(
  fauna = blue_fauna,
  fleets = blue_fleets,
  years = 50
)

critter_biomass <-
  map_df(blue_sim, ~ map_df(.x, ~tibble(ssb = rowSums(.x$ssb_p_a), patch = 1:nrow(.x$ssb_p_a)), .id = "critter"), .id = "step") %>%
  mutate(step = as.numeric(step)) %>%
  left_join(grid, by = "patch") %>% 
  group_by(critter, step) %>% 
  summarise(ssb = sum(ssb)) %>% 
  group_by(critter) %>% 
  mutate(depletion = ssb / ssb[step == 0])

critter_biomass %>%
  ungroup() %>% 
  left_join(sci_to_com, by = "critter") %>%
  select(-critter) %>%
  rename(critter = common_name) %>%
  filter(step == max(step)) %>% 
  ggplot(aes(reorder(critter, depletion), depletion)) +
  geom_col() +
  scale_y_continuous(limits = c(0, 1), name = "SSB/SSB0", expand = expansion(mult = c(0.0,0.05))) +
  labs(x = "") + 
  coord_flip()

```

```{r}
#| label: fig-blue
#| fig-cap: "Contour plot of fish spawning biomass in space under status quo (blue) and range-shifted (red) conditions for blue-water simulation. Packed together lines indicate higher biomass densities, farther spaced lines lower densities. The x-axis represents longitude, the y-axis latitude. "

blue_climate_sim <- simmar(
  fauna = blue_fauna,
  fleets = blue_fleets,
  habitat = future_habitat,
  years = 20,
  manager = list(mpas = list(
    locations = mpa_locations,
    mpa_year = 1
  ))
)

patch_biomass <-
  map_df(blue_climate_sim, ~ map_df(.x, ~tibble(biomass = rowSums(.x$ssb_p_a), patch = 1:nrow(.x$ssb_p_a)), .id = "critter"), .id = "step") %>%
  mutate(step = as.numeric(step)) %>%
  left_join(grid, by = "patch")

patch_biomass %>%
  left_join(sci_to_com, by = "critter") %>% 
    select(-critter) %>% 
  rename(critter = common_name) %>% 
  filter(step == min(step) | step == max(step)) %>% 
  group_by(critter, step) %>%
  mutate(sbiomass = biomass / max(biomass)) %>%
  ungroup() %>%
  ggplot()+
  geom_contour(aes(x,y, z = sbiomass, color= factor(step)), linewidth = .1) +
  facet_wrap(~critter) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        legend.position = "top",
        axis.line = element_blank()) + 
  scale_color_manual(values = c("#63B8FF","#FF4500"), labels = c("Status Quo","Range Shift"), name = '')

```

# Results

### Coral Reefs

For the coastal coral reef example both the *Spawning Ground* and *Target fishing* MPA design strategies were capable of producing a Pareto-optimal or improving, i.e. potentially win-win, outcome where change in total spawning biomass across all species and yield across all fleets was greater than or equal to zero for some MPA size.
However, producing a total increase in SSB/SSB0 greater than 50% came at a net cost to fisheries yield across both design strategies.
The *Spawning Aggregation* strategy produced more immediate trade-offs between food security and conservation (@fig-coral-results).

These macro-level changes in total yield and conservation outcomes mask interesting internal dynamics.
The *Spawning Ground* strategy increased yield to Fleet One when MPAs were relatively small, but was able to maintain yields near *status quo* levels up to a very large MPA size to the dependence of Fleet One on only a relatively small portion of the total simulation area.
However, *Spawning Ground* MPAs were unable to provide increases in yield to Fleet Two.
The *Spawning Ground* strategy produced the most rapid conservation gains for the snapper population.
Interestingly, due to effort displacement from the spawning grounds onto the fishing areas inhabited by the groupers the rest of the year, closure of the spawning aggregation had limited impact on the grouper population until the MPA size exceeded 50% of the simulation area.
Effort displacement from the spawning grounds also resulted in a small decrease in the deep snapper population for MPA sizes less than 50%.
The *Target Fishing* design strategy has similar yield outcomes to the *Spawning Ground* strategy, but less dramatic negative impacts on Fleet Two, at the expense of slightly fewer benefits for Fleet One.
However, the *Target Fishing* strategy provided more uniform conservation benefits to all species (@fig-coral-results).

MPAs were capable of producing positive outcomes in terms of total biomass and yields in the coastal coral reef example, but all the tested strategies resulted in trade-offs across at least one dimension.
`marlin` could be further used by a community such as the one simulated here to explore alternative design strategies, or to assess options for distributing fishery benefits across communities that may experience very different food provision outcomes from the same MPA network.

```{r}
#| label: fig-coral-results
#| fig-cap: "Percent change in yield per fleet (A) and SSB/SSB0 (spawning biomass divided by unfished spawning biomass) per species (B) as a function of MPA size and MPA design strategy for the coastal coral reef case study. Percent change in total yield across both fleets (y-axis) and total SSB/SSB0 across all species (x-axis) and placement strategies (line color) (C). Text labels in panel C indicate the percent of the simulation area in an MPA when change in yields is maximized, is 0, and is -50%. All results reflect outcomes after 20 years of simulated MPA protection."
#| fig-height: 4
#| fig-width: 6

coral_mpa_experiments$mpas <- map(coral_mpa_experiments$results, "mpa")

coral_mpa_experiments$obj <- map(coral_mpa_experiments$results, "obj")

coral_results <- coral_mpa_experiments %>%
  unnest(cols = obj)

coral_fleet_frontier <- coral_results %>%
  pivot_longer(starts_with("fleet_"), names_to = "fleet", values_to = "fleet_yield") %>%
  group_by(prop_mpa, fleet, placement_strategy) %>%
  summarise(yield = sum(fleet_yield),biodiv = sum(unique(biodiv))) %>%
  ggplot(aes(biodiv, yield, color = placement_strategy)) +
  geom_line() +
  scale_x_continuous(name = "Change in SSB/SSB0",limits = c(0, NA)) +
    scale_y_continuous(name = "Total Yield",limits = c(0, NA)) +
  facet_wrap(~ fleet, scales = "free_y")

tmp <- coral_results %>% 
  group_by(prop_mpa, placement_strategy) %>% 
  summarise(biodiv = sum(biodiv), yield = sum(yield)) %>% 
  group_by(placement_strategy) %>% 
  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1,
         delta_biodiv = biodiv / biodiv[prop_mpa == 0] - 1) %>% 
  ungroup() %>% 
    mutate(placement_strategy= fct_relabel(placement_strategy, titler))

frontier_labels <- tmp %>% 
  mutate(delta_50 = (delta_yield - -0.5)^2,
         delta_0 = delta_yield^2) %>% 
  group_by(placement_strategy) %>% 
  filter(prop_mpa > 0) %>% 
  filter(yield == max(yield) | delta_0 == min(delta_0) | delta_50 == min(delta_50)) %>% 
  ungroup() %>% 
  mutate(pmpa = scales::percent(prop_mpa))

coral_frontier <- tmp %>% 
  ggplot(aes(delta_biodiv, delta_yield)) + 
  geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_line(alpha = 0.8, aes(color = placement_strategy),size = 1.1) + 
  geom_text_repel(data = frontier_labels, aes(label = pmpa, color = placement_strategy), min.segment.length = 0, box.padding = 0.2, show.legend = FALSE) +
  scale_x_continuous(name = "Change in Total SSB/SSB0", labels = scales::label_percent(accuracy = 1), guide = guide_axis(n.dodge = 2)) +
  scale_y_continuous(name = "Change in Yield", labels = scales::label_percent(accuracy = 1)) + 
  scale_color_manual(values = c("tomato", "steelblue"),name = '') +
  guides(fill = guide_legend(order = 1, override.aes = list(size = 4))) + 
  labs(tag = "C")  + 
  theme(
    legend.position = c(0.7, .85),
    legend.background = element_rect(fill = "transparent"),
    legend.text = element_text(size = 8)
  )

coral_fleet_yield <- coral_results %>%
  pivot_longer(starts_with("fleet_"),
               names_to = "fleet",
               values_to = "fleet_yield") %>%
  group_by(prop_mpa, fleet, placement_strategy) %>%
  summarise(yield = sum(fleet_yield), biodiv = sum(unique(biodiv))) %>%
  group_by(fleet, placement_strategy) %>%
  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1) %>%
  ungroup() %>% 
  mutate(fleet= fct_relabel(fleet, titler)) %>% 
  ggplot(aes(prop_mpa, delta_yield, color = fleet)) +
    geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_line() +
  facet_wrap( ~ placement_strategy, labeller = labeller(placement_strategy = titler)) +
  scale_x_continuous(name = "MPA Size",
                     labels = scales::label_percent(accuracy = 1)) +
  scale_y_continuous(name = "Yield",
                     labels = scales::label_percent(accuracy = 1)) + 
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 8))  + 
    scale_color_discrete(name = '') + 
    labs(tag = "A")

coral_critter_bio <- coral_results %>%
  group_by(prop_mpa, critter, placement_strategy) %>%
  summarise(biodiv = sum(unique(biodiv))) %>%
  group_by(critter, placement_strategy) %>%
  mutate(delta_bio = biodiv) %>%
  mutate(critter = fct_relabel(critter, titler)) %>% 
  # filter(prop_mpa < 0.4) %>% 
  ggplot(aes(prop_mpa, pmin(Inf,delta_bio), color = critter)) +
  geom_line() +
  facet_wrap( ~ placement_strategy, labeller = labeller(placement_strategy = titler)) +
  scale_x_continuous(name = "MPA Size",
                     labels = scales::label_percent(accuracy = 1),
                     guide = guide_axis(n.dodge = 2)) +
  scale_y_continuous(name = "SSB/SSB0") + 
  scale_color_brewer(name = '', type = "qual", palette = "Dark2") + 
    labs(tag = "B") + 
  theme(axis.text = element_text(size = 8), axis.text.x = element_text(size = 8))

((coral_fleet_yield / coral_critter_bio) | (coral_frontier + plot_layout(guides = "keep")) ) & theme(strip.text = element_text(size  = 6), plot.margin = margin(.1, .1, .1, .1, "cm"), panel.spacing = unit(0.5, "lines"), legend.text = element_text(size = 6))

```

### Pelagic Systems

The yield impacts of the proposed MPA strategies varied across each of our simulations.
Yields of the purse seine fleet were more resilient to the range shift under the *Rate* design strategy, while the longline fleet had the exact opposite result.
Under the *Rate* strategy, longline yields were relatively stable across a large MPA size under the *status quo* conditions, but declined rapidly under the *range shift* conditions.
For the purse seine fleet, yields were more stable under the *range shift* conditions under the *Target Fishing* design scenario, but declined quickly as a function of MPA size under the *status quo* conditions (@fig-blue-results).
In terms of fisheries yield, the *Target Fishing* strategy would produce better yield outcomes for both fishing fleets under the *range shift* conditions, but the *Rate* strategy would perform best under the *status quo*.

From a conservation perspective, the primary tuna species (bigeye, skipjack, and yellowfin) were most sensitive to design strategy and climate scenario, with the *Target Fishing* strategies producing for example rapid conservation gains for bigeye under the *status quo*, while the *Rate* strategy resulted in net conservation losses until the MPA became extremely large.
However, under the *Range Shift* scenario the MPAs had little impact on many of the tuna populations until the MPA size became extremely large, do to movement of the primary fishing grounds outside of the current hotpots.
The 20 year time horizon simulated here was not enough to produce substantial gains for any of the shark species even with 100% closures except for the more rapidly growing blue shark (@fig-blue-results).

```{r}
#| include: false

blue_water_experiments$mpas <- map(blue_water_experiments$results, "mpa")

blue_water_experiments$obj <- map(blue_water_experiments$results, "obj")

blue_water_climate_experiments$mpas <- map(blue_water_climate_experiments$results, "mpa")

blue_water_climate_experiments$obj <- map(blue_water_climate_experiments$results, "obj")

static_results <- blue_water_experiments %>%
  unnest(cols = obj) %>%
  mutate(climate = FALSE)

climate_results <- blue_water_climate_experiments %>%
  unnest(cols = obj) %>%
  mutate(climate = TRUE)

blue_water_results <-static_results %>%
  bind_rows(climate_results) %>% 
  left_join(sci_to_com, by = "critter") %>% 
  select(-critter) %>% 
  rename(critter = common_name)

fleet_climate_results <- blue_water_results %>%
  select(placement_strategy,
         fleet_model,
         prop_mpa,
         climate,
         longline,
         critter,
         purseseine) %>%
  pivot_longer(c(longline,purseseine), names_to = "fleet", values_to = "yield") %>%
  pivot_wider(names_from = "climate", values_from =  "yield") %>%
  mutate(delta = `TRUE` - `FALSE`) %>%
  group_by(placement_strategy, prop_mpa, fleet) %>%
  summarise(delta = sum(delta))

fleet_climate_impacts <- fleet_climate_results %>%
  ggplot(aes(prop_mpa, delta, color = fleet)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_line() +
  facet_wrap(~placement_strategy) +
  scale_y_continuous(name = "Difference in Yield") +
  theme(axis.text.x = element_blank(), axis.title.x = element_blank())

critter_climate_results <- blue_water_results %>%
  select(placement_strategy,
         fleet_model,
         prop_mpa,
         climate,
         critter,
         biodiv) %>%
  pivot_wider(names_from = "climate", values_from =  "biodiv") %>%
  mutate(delta = `TRUE` - `FALSE`)

critter_climate_impacts <- critter_climate_results %>%
  ggplot(aes(prop_mpa, delta, color = critter)) +
  geom_line() +
  facet_wrap(~placement_strategy) +
  scale_x_continuous(name = "MPA",labels = scales::label_percent()) +
  scale_y_continuous(name = "Difference in SSB/SSB0")

fleet_climate_impacts / critter_climate_impacts

```

```{r}
#| label: fig-blue-results
#| fig-cap: "Change in yield (A) and SSB/SSB0 (spawning biomass divided by unfished spawning biomass) (B) as a function of MPA size and design strategy by fishing fleet and species. Blue lines indicate impacts of MPAs under status quo habitat, red impacts under climate-driven range shift. Results reflect the outcome of 20 years of simulated MPA protection."
#| fig-height: 6
#| fig-width: 6

linesize <- 0.5

# blue_water_results <- blue_water_results %>% 
#   rename(`purse seine` = purseseine)

blue_fleet_frontier <- blue_water_results %>%
  pivot_longer(names(blue_fleets), names_to = "fleet", values_to = "fleet_yield") %>%
  group_by(prop_mpa, fleet, placement_strategy, climate) %>%
  summarise(yield = sum(fleet_yield),biodiv = sum(unique(biodiv))) %>%
  ggplot(aes(biodiv, yield, color = placement_strategy, linetype = climate)) +
  geom_line() +
  scale_x_continuous(name = "Change in Total SSB/SSB0",limits = c(0, NA)) +
    scale_y_continuous(name = "Total Yield",limits = c(0, NA)) +
  facet_wrap(~ fleet, scales = "free_y")

blue_frontier <- blue_water_results %>% 
  filter(climate == FALSE) %>% 
  group_by(prop_mpa, placement_strategy, climate) %>% 
  summarise(biodiv = sum(biodiv), yield = sum(yield)) %>% 
  group_by(placement_strategy) %>% 
  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1,
         delta_biodiv = biodiv / biodiv[prop_mpa == 0] - 1) %>% 
  ggplot(aes(delta_biodiv, delta_yield, color = placement_strategy, linetype = climate)) + 
  geom_vline(aes(xintercept = 0), linetype = 2) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_line(size = linesize) + 
  scale_x_continuous(name = "Change in Conservation", labels = scales::label_percent(accuracy = 1)) +
  scale_y_continuous(name = "Change in Yield", labels = scales::label_percent(accuracy = 1)) + 
  theme(legend.position = "top")

blue_fleet_yield <- blue_water_results %>%
  filter(placement_strategy %in% c("target_fishing", "rate")) %>%
  pivot_longer(names(blue_fleets),
               names_to = "fleet",
               values_to = "fleet_yield") %>%
  group_by(prop_mpa, fleet, placement_strategy, climate) %>%
  summarise(yield = sum(fleet_yield), biodiv = sum(unique(biodiv))) %>%
  group_by(fleet, placement_strategy) %>%
  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1) %>%
  ggplot(aes(prop_mpa, delta_yield, color = climate)) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_line(size = linesize, show.legend = FALSE) +
  facet_grid(fleet ~ placement_strategy, scales = "free_y", labeller = labeller(fleet = titler, placement_strategy = titler)) +
  scale_x_continuous(name = "MPA Size",
                     labels = scales::label_percent(accuracy = 1)) +
  scale_y_continuous(name = "Change in Yield",
                     labels = scales::label_percent(accuracy = 1)) +
  scale_color_manual(
    values = c("#63B8FF", "#FF4500"),
    labels = c("Status Quo", "Range Shift"),
    name = ''
  ) + 
  theme(strip.text = element_text(size = 8)) + 
  labs(tag = "A")

blue_critter_bio <- blue_water_results %>%
    filter(placement_strategy %in% c("target_fishing", "rate")) %>% 
  group_by(prop_mpa, critter, placement_strategy, climate) %>%
  summarise(biodiv = sum(unique(biodiv))) %>%
  group_by(critter, placement_strategy) %>%
  mutate(delta_bio = biodiv) %>%
  ungroup() %>% 
  mutate(placement_strategy = fct_relabel(as.factor(placement_strategy), titler)) %>% 
  ggplot(aes(prop_mpa, pmin(1,delta_bio), color = climate, linetype = placement_strategy)) +
  geom_line(size = linesize) + 
  facet_wrap(~critter) +
  scale_color_manual(values = c("#63B8FF","#FF4500"), labels = c("Status Quo","Range Shift"), name = '') +
  scale_x_continuous(name = "MPA Size",
                     labels = scales::label_percent(accuracy = 1), guide = guide_axis(n.dodge = 2)) +
  scale_y_continuous(name = "SSB/SSB0",limits = c(0, NA)) +
  scale_linetype(name = "MPA Strategy") + 
    labs(tag = "B")+
  theme(strip.text = element_text(size = 7),
        axis.text.y = element_text(size = 6))

# blue_frontier / blue_critter_bio + plot_layout(heights = c(1,2))
(blue_fleet_yield / 
blue_critter_bio) + plot_layout(heights = c(1.5,2))
```

# Discussion

Marine ecosystems are driven by complex social-ecological dynamics.
Communities must often make decisions on how to manage these systems based on limited empirical evidence.
Modeling tools such as `marlin` can help users answer important scientific questions and design marine management policies informed by a better understanding of sensitivities to key uncertainties.
In this paper we have documented the core system of equations underlying `marlin`, and demonstrated how `marlin` can be used to support efforts such as MPA planning.
The variation in both the magnitude and distribution of costs and benefits stemming from alternative design choices and states of nature shown in our case study results demonstrates the value in our method.
For example, our coastal coral reef example illustrates both the potential for MPAs to benefit conservation and food production in these systems, and the potential for the same MPA to benefit some fleets and species while harming others when fishing fleets affected by an MPA do not share the same objectives and species are not uniformly distributed.
Our pelagic case study illustrates how `marlin` can be used to assess the climate resiliency of spatial management strategies, and identify strategies that best meet the needs of both current and future conditions.
Results from this type of work could be used to help prioritize communities and species at particular risk to climate change impacts.

The role of this paper is not to conduct a review of the many tools for spatial-temporal modeling available in the literature, each of which provide useful functionality for specific applications.
However, we highlight here the specific gaps that we feel `marlin` fills in the modeling literature and in the policy support toolbox using some selected publications.
More end-to-end models such as ATLANTIS [@audzijonyte2019] , OSMOSE [@shin2001], Ecopath with Ecosim [@christensen2004], POSEIDON [@bailey2019], DISPLACE [@bastardie2013], or SEAPODYM [@lehodey2008] are capable of representing tremendous amounts of complexity, but can be extremely time consuming to design and run.
`marlin` allows some of the realism of these end-to-end models while being simpler and faster to parameterize and run.

Tools such as virtualspecies [@leroy2016], STEPS [@visintin2020], RangeShifter [@bocedi2014], and [SMS](https://www.aquaveo.com/software/sms-surface-water-modeling-system-introduction) can be efficiently constructed to model the dispersal and distribution of species in space and time as a function of environmental covariates.
However, these typically require specifying a covariate-response curve without explicitly acknowledging how this arises from habitat-specific movement and demography (virtualspecies), or model movement using a dispersal kernel (RangeShifter, STEPS) or a least-cost path algorithm (SMS, STEPS).
By contrast, `marlin` uses a continuous-time Markov chain movement model, which integrates multiple paths (including their path-dependent probability based on intervening habitat types and species preferences) while using scale-free parameters that can be measured experimentally in laboratory or tagging studies.

Simplified bio-economic models such as those used in @hastings2017, @sala2021, and @cabral2019 can be applied at scale and provide analytically tractable results, but as a result must abstract over many bio-economic dynamics that can be important for more tactical applications.
Our results demonstrate how nuances in fleet dynamics and species distributions can dramatically impact MPA outcomes.
`marlin` allows for more realistic representations of spatial social-ecological systems while maintaining processing speed.

In sum, `marlin` complements the existing spatial marine modeling literature by allowing scientists, decision makers, and other stakeholders to efficiently examine the impacts of realistic bio-economic dynamics on academic and applied problems.
For example, scientists could use `marlin` to consider the relative importance of different biological variables in affecting policy outcomes.
Managers could use `marlin` to assess the potential of proposed MPA networks to achieve stated objectives, such as maximizing conservation gains for a threatened species while minimizing or removing negative impacts to affected fishing fleets.

Beyond the current examples, we envision `marlin` being applicable to research on dynamic ocean management, range shifts, management strategy evaluation, policy interactions, and spatial stock assessment.
In addition, the process-based movement model used in `marlin` can directly use empirical estimates of movement dynamics derived in the manner of @thorson2021a, providing a link between empirical and simulation based approaches to marine resource management.
Future iterations of the model may include life history modules for a broader range of species types such as marine mammals, sea birds, and sea turtles.

Even the most complex marine model is a stylized cartoon of the true dynamics at play in the ocean.
However, for all their limitations, models such as `marlin` can help users understand factors that drive the performance of marine management strategies.
The goal of this tool is to empower people to design policies based on evaluation of key uncertainties and trade-offs, and in doing so support more effective and equitable marine resource management.

# Data Availability Statement

All code and materials needed to reproduce this paper is available at (https://github.com/DanOvando/marlin-paper)\[https://github.com/DanOvando/marlin-paper\].

# References

::: {#refs}
:::

# Supporting Information