### 2023-05-23 presentations/impac_slides.qmd
testing slides deploy

---
title: "Assessing the Population-level Impacts of Marine Protected Areas"
subtitle: "IMPAC 2023"
author: "Dan Ovando"
date: "2023/02/04"
format:
  revealjs:
    fig-align: center
    auto-stretch: true
  html: default
  pptx:
    reference-doc: "template.pptx"
execute: 
  eval: true
  message: false
  warning: false
---

```{r}
#| label: setup
#| include: false
library(knitr)

library(tufte)

library(tidyverse)

library(here)

```

## Population-Effects of MPAs

:::: {.columns}
::: {.column width="60%"}
- Thanks to co-authors and PISCO divers
- MPAs a growing part of marine management toolkit
- *Predicting* and *estimating* effects critical for planning and adaptation

**What effects do protected areas have on populations inside and outside of borders?**
:::
::: {.column width="40%"}
![](imgs/paper_cover.png)
:::
::::

## MPAs have Complex Effects {.smaller}

:::: {.columns}
::: {.column width="50%"}

Fishing closures have complex impacts beyond borders

- "spillover" of adult or larval organisms

- Displacement of human activities

- Across multiple species and fleets!

**Creates a challenge for both predicting and estimating effects of spatial policies**

:::

::: {.column width="50%"}
![](imgs/IMG_1679.JPG)
:::
::::

## MPAs - The Evidence {.smaller}

:::: {.columns}
::: {.column width="50%}

:::{.incremental}
- MPAs seem simple: 
  - *Less fishing, more fish*

- Ample evidence of **more everything** inside MPAs than outside (Lester *et al.* 2009)

- Problem solved?

  - Fish and Fishers move

  - What do we do when the **treatment** affects the **control**

  How do we estimate **net effect** inside and out?
:::
:::
::: {.column width="50%"}
![](imgs/lester.jpg)
:::
::::

## Population-Level Effects of MPAs

![](imgs/bio_anim.gif)

::: notes
Relatively sedentary adults, widely dispersed larvae with post-settlement density dependence
:::

## Population-Level Effects of MPAs

![](imgs/fleet_anim.gif)

::: notes
Relatively sedentary adults, widely dispersed larvae with post-settlement density dependence
:::

## The Challenge of Spatial MPA evaluation

::: {.incremental}

- The goal of a control site is to account for **differences and changes caused by things other than the MPA that you do not directly observe**

- The **closer** your control site, the more risk of spillover and fleet movement affecting your "control"

- The **further** your control site, the less likely it will experience common environmental shocks

- Is there an alternative to spatial controls?
:::
## Some useful tips here

![](imgs/hilborn.png)

## Case Study: Channel Islands MPAs

![](imgs/ci.png)

## What Happend in the Channel Islands?

![](imgs/raw_targ_trend.png){fig.align="center"}

## What Happend in the Channel Islands?

![](imgs/raw_trend.png)

## Estimating Population-Level Effects

![](imgs/did.001.jpeg)

## Estimating Population-Level Effects

![](imgs/did.002.jpeg)

---

## Estimating Population-Level Effects

![](imgs/did.003.jpeg)

---

## Estimating Population-Level Effects

### General form

$$(log(D_{MPA=1, T = 1})-log(D_{MPA=0, T = 1}))-(log(D_{MPA = 1,T = 0})-log(D_{MPA =0, T = 0}))$$

### Bayesian Gamma GLM

$$d_{i} \sim Gamma(e^{\beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + \mathbf{B^cX_i} + \mathbf{B^sS_i}},shape)$$

### Hierarchical clustering of sites by island
$$\mathbf{B^s} \sim N(\beta_r,\sigma_r)$$

## Population-Level Effects of MPAs

![](imgs/a_priori_plot.png)

## Population-Level Effects of MPAs {.nostretch}

![](imgs/mpa_effect_plot.png)

## Effects of Protected Areas {.smaller}

::: columns
::: {.column width="50%"}

- Critical to consider effects **inside** and **outside**

-  **Population** effects may be smaller and harder to find than conventionally thought

- Uncertainty does not preclude action

- Communities should have a clear understanding of full range of possible effects of protected areas

:::
::: {.column width="50%"}
![](imgs/IMG_2947.JPG)
:::
:::

## The Next Generation of MPA Models

:::: {.columns}
::: {.column width="60%"}
Empirical strategies should...

- Acknowledge uncertainty
- Have a clear and defensible strategy for causal inference
- Be grounded in theory

:::
::: {.column width="40%"}

Predictive models should...

- Assess where they fall on the "tactical" to "strategic" spectrum
- Acknowledge, explore, and communicate key uncertainties
- Remember that all models are wrong, some are useful
:::
::::

## Reliable Estimation Matters

![](imgs/complexity.png)

## `marlin`

Be on the lookout for a new MPA modeling tool called [`marlin`](https://danovando.github.io/marlin/) in the coming year.

- Full age structure
- 2D spatial structure
- Multi-species & Multi-fleet
- Realistic movement dynamics

**Facilitate simulation testing of MPA design outcomes**

## Questions?

:::: {.columns}
::: {.column width="60%"}

**email:** dovando@iattc.org

**website:** [danovando.com](https://www.weirdfishes.blog/)

**twitter:** [@danovand0](https://twitter.com/DanOvand0)

**Funding:** NOAA Sea Grant Population and Ecosystem Dynamics Fellowship

:::
::: {.column width="40%"}

![](imgs/squats.jpg)
:::
::::

# Extras

## Inside vs. Outside = Effect?

![](imgs/rr1.jpg)

## Inside vs. Outside = Effect?

![](https://raw.githubusercontent.com/DanOvando/weird-fishes/master/static/blog/2021-07-06-pop-effects-of-mpas_files/biased_dr_plot.jpg)

## Details Matter

![](imgs/complexity)

## Details Matter

![](imgs/need.png)

## Effect Size Matters

![](imgs/ERRORS.PNG)

## MPAs - The Evidence

<br>
<br>
<br>

> The jury is in on marine reserves: They work.
> Research has repeatedly shown that fish numbers quickly climb following well-enforced fishing bans, 
> creating tangible benefits for fishers who work the surrounding waters.

[NPR 2018](https://www.npr.org/sections/thesalt/2018/09/14/647441547/could-a-ban-on-fishing-in-international-waters-become-a-reality)