### 2023-05-23 presentations/tunacamp_slides.qmd
embed-resources: true

---
title: "Revealing Benefits, Costs, and Tradeoffs of Spatial Management Arrising from Coupled Economic and Ecological Dynamics in Marine Systems"
author: "Dan Ovando"
format:
  revealjs:
    embed-resources: true
---

## Testing

Hello

## Complexity Matters!

![](imgs/complexity.png)