### 2023-05-23 presentations/tunacamp_slides.qmd
testing equations

---
title: "Revealing Benefits, Costs, and Tradeoffs of Spatial Management Arrising from Coupled Economic and Ecological Dynamics in Marine Systems"
author: "Dan Ovando"
format:
  revealjs:
    embed-resources: true
---

## Testing

Hello

## Equations

$$
\begin{aligned}
n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p,s}}\times{h_s}\times\sum_{p=1}^P{ssb_{t-1,p,s}}}{0.2\times{\sum_{p=1}^Pssb0_{p,s}} \times(1 - h_s)+(h_s - 0.2) \times{\sum_{p=1}^Pssb_{t-1,p,s}}}\right) \times \\ {r0_{p,s} / \sum_p^P{r0_{p,s}}\times \epsilon_{t,s}
}
\end{aligned}
$$ {#eq-dd2}

## Complexity Matters!

![](imgs/complexity.png)