### 2023-05-23 presentations/tunacamp_slides.qmd
pushing with less cmbedding

---
title: "Revealing Benefits, Costs, and Tradeoffs of Spatial Management Arrising from Coupled Economic and Ecological Dynamics in Marine Systems"
author: "Dan Ovando"
format:
  revealjs:
    embed-resources: false
    fig-align: center
    auto-stretch: true
    theme: moon
---

## Acknowledgements

Daniel Ovando^1,2^, Darcy Bradley^3,4,5^, Echelle
Burns^3,4,5^,Lennon Thomas^3,4,5^, James Thorson^6^

Affiliations: ^1^ School of Aquatic and Fishery Sciences, University of
Washington,1122 NE Boat St Box 355020,Seattle, WA, USA ^3^ Marine
Science Institute, University of California, Santa Barbara, CA, USA ^4^
Bren School of Environmental Science & Management, University of
California, Santa Barbara, CA, USA ^5^ Environmental Markets Lab,
University of California, Santa Barbara, CA, USA ^6^ Habitat and
Ecological Processes Research Program, Alaska Fisheries Science Center,
NOAA, NMFS, NOAA, Seattle, WA, USA

## All Models are Wrong

:::: {.columns}

::: {.column width="60%"}

- But some are hopefully useful for something?

-

:::

::: {.column width="40%"}
Right column
:::

::::

## Lots of Spatio-Temporal Models Out There

![](imgs/fulton-mpas.jpg)

::: aside
[Fulton *et al.* 2015](https://royalsocietypublishing.org/doi/10.1098/rstb.2014.0278)
:::

## The marlin package

[`marlin`](https://danovando.github.io/marlin/) is an R package currently available on GitHub

```{r}
#| eval: false

remotes::install_github("danovando/marlin")

```

Core Features

- Full age structure of multiple fish(y) critters (automated life history with [FishLife](https://github.com/James-Thorson-NOAA/FishLife))

- Multiple fleets and *métiers* enabling **technical interactions**
  - Input controls, output controls, bio-economic dynamics

- 2-D spatial structure with scale-free and estimable movement

- Fast! (fractions of a second for 10x10 system over 20 years with quarterly steps)

## The marlin Package

Best suited for *what if* as opposed to *what will* questions

  -

## Fish Model

## Movement Model

## Fleet Model

## Manager Model

## Coral Reef Case Study

![](../03_marlin_paper_files/figure-pdf/fig-coral-1.png)

## Equations

$$
\begin{aligned}
n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p,s}}\times{h_s}\times\sum_{p=1}^P{ssb_{t-1,p,s}}}{0.2\times{\sum_{p=1}^Pssb0_{p,s}} \times(1 - h_s)+(h_s - 0.2) \times{\sum_{p=1}^Pssb_{t-1,p,s}}}\right) \times \\ {r0_{p,s} / \sum_p^P{r0_{p,s}}\times \epsilon_{t,s}
}
\end{aligned}
$$ {#eq-dd2}

## Complexity Matters!

![](imgs/complexity.png)

## Linking to Team Tuna

- Still stuck in the dietary dark ages 
  - Diet data could be used to inform taxis matrix?

- Humans need to be part of the "diet Renaissance"!

- Provides a direct and mechanistic link between tagging studies and simulation work
  - Can be linked to advection and environmental covariates as a bonus!

- Future work adding in more species modules
  - Seabirds?
  - Marine mammals?
  - Turtles?