### 2023-05-23 presentations/tunacamp_slides.qmd
stop tracking slides, not worth it

---
title: "Simulating Benefits, Costs, and Tradeoffs of Spatial Management in Marine Social-Ecological Systems"
subtitle: "73rd Annual Tuna Conference"
author: "Dan Ovando"
format:
  revealjs:
    embed-resources: true
    fig-align: center
    auto-stretch: true
    theme: [moon, custom.scss]
---

## Acknowledgements

**Daniel Ovando^1,2^, Darcy Bradley^3,4,5^, Echelle Burns^3,4,5^,Lennon
Thomas^3,4,5^, James Thorson^6^**

**Funding**: Waitt Foundation

::: aside
Affiliations: ^1^ School of Aquatic and Fishery Sciences, University of
Washington,1122 NE Boat St Box 355020,Seattle, WA, USA ^3^ Marine
Science Institute, University of California, Santa Barbara, CA, USA ^4^
Bren School of Environmental Science & Management, University of
California, Santa Barbara, CA, USA ^5^ Environmental Markets Lab,
University of California, Santa Barbara, CA, USA ^6^ Habitat and
Ecological Processes Research Program, Alaska Fisheries Science Center,
NOAA, NMFS, NOAA, Seattle, WA, USA
:::

## All Models are Wrong

But some are hopefully useful for something?

- At the *strategic* level can help provide general insights around *what kinds of things matter*

- At the *tactical* level can help provide *specific answers* when experiments are off the table

. . .

What is the best model of an elephant?

. . .

Depends on objectives

## Lots of Spatio-Temporal Models Out There

![](imgs/fulton-mpas.jpg)

::: aside
[Fulton *et al.*
2015](https://royalsocietypublishing.org/doi/10.1098/rstb.2014.0278)
:::

## The marlin package {.smaller}

[`marlin`](https://danovando.github.io/marlin/) is an R package
currently available on GitHub

```{r}
#| eval: false

remotes::install_github("danovando/marlin")

```

#### Core Features

-   Full age structure of multiple fish(y) critters (automated life
    history with
    [FishLife](https://github.com/James-Thorson-NOAA/FishLife))

-   Multiple fleets and *métiers* enabling **technical interactions**

    -   Input controls, output controls, bio-economic dynamics

    - Bycatch impacts

    - Distributional impacts of policies on fish and people

-   2-D spatial structure with scale-free and estimable movement

-   Fast! (fractions of a second for 10x10 system over 20 years with
    quarterly steps)

# Help users test the impacts of bio-economic uncertainties on social-ecological outcomes

. . .

Best suited for *what if* as opposed to *what will* questions

## Fish Model {.smaller}

- Flexible time steps

- Simultaneously simulate *N* age-structured fish(ish) populations

- von Bertalanffy growth

- Flexible fecundity at age

- [Lorenzen natural mortality](https://linkinghub.elsevier.com/retrieve/pii/S0165783622002314) by default

- 6 different timing of Beverton-Holt recruitment
  - e.g. pre vs. post settlement
  - With or without potentially auto-correlated recruitment deviates

## Movement Model

Movement simulated using a *continuous-time Markov chain (CTMC)^[[Thorson et al. 2021](https://onlinelibrary.wiley.com/doi/10.1111/faf.12592)]*

- Movement governed by *diffusion* and *taxis* matrices

  - Can be separate for larvae vs. adults

  - See later releases for age-specific and *advection*

  - Can be empirically estimated!

  - Can dodge barriers (e.g. land)

## Movement Model {.smaller}

*Instantaneous* movement matrix is a function of diffusion rate *D*, habitat matrix *H*, and the spatial ($\Delta_d$) and temporal ($\Delta_t$) resolution

$$ 
M_{p1,p2,t,s,a} = \begin{cases}
      = \frac{\Delta_{t}}{\Delta_{d}^2}De^{\frac{\Delta_t(H(p2,t,s,a) - H(p1,t,s,a))}{\Delta_d}} & \text{if p2 and p1 are adjacent}\\
     = -\sum_{p' \neq p1} M_{p1,p2,t,s,a} & \text{if p1 = p2}\\
     = 0 & \text{otherwise.}
\end{cases}
$$ {#eq-diffusion}

. . .

Numbers at age by patch then distributed each time step via matrix exponential

$$
\pmb{n}_{t+1,s,a} = \pmb{n}_{t,s,a}e^{\pmb{M}_{t,s,a}}
$$ {#eq-movement}

. . .

Scale-free parameterization means parameters can be preserved and compared across models with different spatio-temporal resolution

## {background-image=imgs/ctmc.gif background-size=contain}

## Fleet Model

- *N* fleets defined by *M* *metiers*

- Fleets have
  - Cost per unit effort
  - Travel costs from ports

- *metier* defined by
  - Species
  - Price
  - Selectivity at age
  - Catchability in space

## Fleet Model {.smaller}

Catch per species calculated by Baranov equation

$$
Y_{t,p,s,a} = \frac{u_{t,p,s,a}}{z_{t,p,s,a}} \times b_{t,p,s,a} \times (1 - e^{-z_{t,p,s,a}})
$$ {#eq-yield}

. . . 
And then revenues divided to each fleet proportionally

$$
 R_{t,p,f} = \sum_s^{N_s} \sum_a^{N_a} \Pi_{s,f}\frac{\alpha_{a,s,f}q_{s,f}E_{t,p,f}}{u_{t,p,s,a}} \times Y_{t,p,s,a}
$$ {#eq-revenue}

. . .

Fishers then allocate effort in space based on total profits in space in last time step (conditional on management...)^[Allows for economic MPAs!]

## Fleet Model

Absent management fleet can either be *constant effort*^[with or without MPA attrition] or *open access*

When *open access*, fishing effort expands and contracts in response to total profits

. . .

$$
E_{t+1,f} = E_{t,f} \times e^{\theta_f log(R_{t,f} / C_{t,f})} 
$$ {#eq-oa}

## Manager Model

Managers can set arbitrary combinations of

- Size / slot limits

- Fishing seasons by species

- Effort limits by species

- Catch quotas by species

- MPAs

- "Bycatch" penalties

## Coral Reef Case Study

![](imgs/03_marlin_paper_files/figure-pdf/fig-coral-1.png){.absolute left=50 height="600"}

## Coral Reef Case Study

![](imgs/03_marlin_paper_files/figure-pdf/fig-coral-results-1.png){.absolute left=50 height="600"}

## Pelagic Case Study

![](imgs/03_marlin_paper_files/figure-pdf/fig-blue-1.png){.absolute left=50 height="600"}

## Pelagic Case Study

![](imgs/03_marlin_paper_files/figure-pdf/fig-blue-results-1.png){.absolute top=100 left=200 width="650" height="550"}

## Complexity Matters!

. . .

![](imgs/complexity.png)

## Linking to Team Tuna {.smaller}

-   Many models still stuck in the dietary dark ages

    -   Diet data could be used to inform taxis matrix?

-   Humans need to be part of the "diet Renaissance"!

-   `marlin` provides a mechanistic link between tagging studies and
    simulation work

    -   Can be linked to advection and environmental covariates as a
        bonus!

-   Future work adding in more species modules

    -   Seabirds?
    -   Marine mammals?
    -   Turtles?

## Of MICE and `marlin`

::: columns
::: {.column width="60%"}

- Spatial stock assessment and management increasingly needed

- BBNJ and 30x30 means that "tunas" are now firmly in the great MPA debate

- Lack of consistent / transparent assumptions can produce inconsistent outcomes

- MICE models like `marlin` can help illuminate drivers of social-ecological outcomes in fisheries

:::
::: {.column width="40%"}

![](imgs/bbnj.png)

![](imgs/medoff.png)

![](imgs/hampton.png)

:::

:::
## Questions?

::: columns
::: {.column width="65%"}
-   Be on the lookout for `marlin` paper sometime in 2023 (fingers
    crossed)

-   Check out the model at <br> [https://danovando.github.io/marlin/](https://danovando.github.io/marlin/)

-   Open source! Contributions welcome

-   email: dovando@iattc.org

-   github: danovando

-   twitter: @danovand0
:::

::: {.column width="35%"}
![](imgs/IMG_1679.JPG)
:::
:::