### 2023-06-05 presentations/tunacamp_slides.qmd
updating lockfile

---
title: "Simulating Benefits, Costs, and Tradeoffs of Spatial Management in Marine Social-Ecological Systems"
subtitle: "73rd Annual Tuna Conference"
author: "Dan Ovando"
format:
  revealjs:
    embed-resources: true
    fig-align: center
    auto-stretch: true
    theme: [moon, custom.scss]
---

## Time for Models! {background-image="https://c.tenor.com/FtuByrhdiwIAAAAC/monty-python.gif" background-size=contain}

## Acknowledgements

**Daniel Ovando^1,2^, Darcy Bradley^3,4,5^, Echelle Burns^3,4,5^,Lennon
Thomas^3,4,5^, James Thorson^6^**

**Funding**: Waitt Foundation

::: aside
Affiliations: ^1^ School of Aquatic and Fishery Sciences, University of
Washington,1122 NE Boat St Box 355020,Seattle, WA, USA ^3^ Marine
Science Institute, University of California, Santa Barbara, CA, USA ^4^
Bren School of Environmental Science & Management, University of
California, Santa Barbara, CA, USA ^5^ Environmental Markets Lab,
University of California, Santa Barbara, CA, USA ^6^ Habitat and
Ecological Processes Research Program, Alaska Fisheries Science Center,
NOAA, NMFS, NOAA, Seattle, WA, USA
:::

## All Models are Wrong

But some are hopefully useful for something?

- At the *strategic* level can help provide general insights around *what kinds of things matter*

- At the *tactical* level can help provide *specific answers* when experiments are off the table

. . .

What is the best model of an elephant?

. . .

Depends on objectives

## Lots of Spatio-Temporal Models Out There

![](imgs/fulton-mpas.jpg)

::: aside
[Fulton *et al.*
2015](https://royalsocietypublishing.org/doi/10.1098/rstb.2014.0278)
:::

## The marlin package {.smaller}

[`marlin`](https://danovando.github.io/marlin/) is an R package
currently available on GitHub

```{r}
#| eval: false

remotes::install_github("danovando/marlin")

```

#### Core Features

-   Full age structure of multiple fish(y) critters (automated life
    history with
    [FishLife](https://github.com/James-Thorson-NOAA/FishLife))

-   Multiple fleets and *métiers* enabling **technical interactions**

    -   Input controls, output controls, bio-economic dynamics

    - Bycatch impacts

    - Distributional impacts of policies on fish and people

-   2-D spatial structure with scale-free and estimable movement

-   Fast! (fractions of a second for 10x10 system over 20 years with
    quarterly steps)
      - R in the front, C++ in the back

- Recursive (you can pass it to itself) for customized applications (e.g. MSE)

# Help users test the impacts of bio-economic uncertainties on social-ecological outcomes

. . .

Best suited for *what if* as opposed to *what will* questions

## Fish Model {.smaller}

::: columns
::: {.column width="60%"}

- Flexible time steps

- Simultaneously simulate *N* age-structured fish(ish) populations

- von Bertalanffy growth

- Flexible fecundity at age

- [Lorenzen natural mortality](https://linkinghub.elsevier.com/retrieve/pii/S0165783622002314) by default

- 6 different timing of Beverton-Holt recruitment
  - e.g. pre vs. post settlement
  - With or without potentially auto-correlated recruitment deviates

- Flexible spawning seasons

- We'll get to movement...
:::
::: {.column width="40%"}

![](imgs/fishy.jpeg)

:::
:::

## Movement Model

Movement simulated using a *continuous-time Markov chain (CTMC)^[[Thorson et al. 2021](https://onlinelibrary.wiley.com/doi/10.1111/faf.12592)]*

- Movement governed by *diffusion* and *taxis* matrices

  - Can be separate for larvae vs. adults

  - See later releases for age-specific and *advection* (aka oceanography!)

  - Can be empirically estimated!

  - Can dodge barriers (e.g. land)

## Movement Model {.smaller}

*Instantaneous* movement matrix is a function of diffusion rate *D*, habitat matrix *H*, and the spatial ($\Delta_d$) and temporal ($\Delta_t$) resolution

$$ 
M_{p1,p2,t,s,a} = \begin{cases}
      = \frac{\Delta_{t}}{\Delta_{d}^2}De^{\frac{\Delta_t(H(p2,t,s,a) - H(p1,t,s,a))}{\Delta_d}} & \text{if p2 and p1 are adjacent}\\
     = -\sum_{p' \neq p1} M_{p1,p2,t,s,a} & \text{if p1 = p2}\\
     = 0 & \text{otherwise.}
\end{cases}
$$ {#eq-diffusion}

. . .

This is the important part

$$ 
\frac{\Delta_{t}}{\Delta_{d}^2}De^{\frac{\Delta_t(H(p2,t,s,a) - H(p1,t,s,a))}{\Delta_d}} 
$$ {#eq-diffusion2}

## Movement Model

Numbers at age by patch then distributed each time step via matrix exponential

$$
\pmb{n}_{t+1,s,a} = \pmb{n}_{t,s,a}e^{\pmb{M}_{t,s,a}}
$$ {#eq-movement}

. . .

Scale-free parameterization means parameters can be preserved and compared across models with different spatio-temporal resolution

## {background-image=imgs/ctmc.gif background-size=contain}

## Fleet Model

- *N* fleets defined by *M* *metiers*

- Fleets have
  - Cost per unit effort
  - Travel costs from ports

- *metier* defined by
  - Species
  - Price
  - Selectivity at age
  - Catchability in space

## Fleet Model {.smaller}

Catch per species calculated by Baranov equation

$$
Y_{t,p,s,a} = \frac{u_{t,p,s,a}}{z_{t,p,s,a}} \times b_{t,p,s,a} \times (1 - e^{-z_{t,p,s,a}})
$$ {#eq-yield}

. . . 
And then revenues divided to each fleet proportionally

$$
 R_{t,p,f} = \sum_s^{N_s} \sum_a^{N_a} \Pi_{s,f}\frac{\alpha_{a,s,f}q_{s,f}E_{t,p,f}}{u_{t,p,s,a}} \times Y_{t,p,s,a}
$$ {#eq-revenue}

. . .

Fishers then allocate effort in space based on total profits in space in last time step (conditional on management...)^[Allows for economic MPAs!]

## Fleet Model

Conditional on management management fleet can either be *constant effort*^[with or without MPA attrition] or *open access*

When *open access*, fishing effort expands and contracts in response to total profits

. . .

$$
E_{t+1,f} = E_{t,f} \times e^{\theta_f log(R_{t,f} / C_{t,f})} 
$$ {#eq-oa}

## Manager Model

Managers can set arbitrary combinations of

- Size / slot limits

- Fishing seasons by species

- Effort limits by species

- Catch quotas by species

- MPAs

- "Bycatch" penalties

## Coral Reef Case Study

![](imgs/03_marlin_paper_files/figure-pdf/fig-coral-1.png){.absolute left=50 height="600"}

::: {.notes}

Fleet One primarily targets the grouper and near-shore snapper populations, but will land
any incidentally captured sharks. Fleet One has a logistic selectivity
pattern for all species, as they retain any fish caught for consumption
or sale. both grouper and shallow-reef snapper due to demand for "plate
sized" fish , and logistic selectivity for sharks, as any shark caught
is retained. Fleet One is totally dependent on fishing for their
livelihood, meaning the local community takes advantage of every
possible opportunity to fish, and as such we model it as a "constant
effort" fishery. Do to having less efficient boats, Fleet One has a
higher cost per distance coefficient than Fleet Two. Fleet One's home
port is located on the site of the grouper spawning aggregation.

Fleet Two is a more commercial fleet that primarily targets the
deep-water snapper population. This fleet primarily sells their catch to
local restaurants and distributors where plate-sized fish are prized,
and so for both snapper and grouper Fleet Two has a dome-shaped
selectivity pattern [@kindsvater2017]. While plate-sized deep snapper
are the primary target of Fleet Two, we model Fleet Two's selectivity
for deep snapper as logistic due to high levels of discard mortality for
deep-water snapper resulting from barotrauma. Fleet Two catches
groupers, though less than Fleet One, and receives no price for sharks
due to the requirements of a certification program through which they
sell their deep-water snapper. Accidental captures (bycatch) of sharks
do occur, which results in mortality. Fleet Two operates under
open-access dynamics, as fishing is not the only means of subsistence
for this community; short-term effort expands and contracts in response
to profitability of the primarily grouper-driven fishery. Fleet two's
home port is located further north from Fleet One.

:::

## Coral Reef Case Study

![](imgs/03_marlin_paper_files/figure-pdf/fig-coral-results-1.png){.absolute left=50 height="600"}

## Pelagic Case Study

![](imgs/03_marlin_paper_files/figure-pdf/fig-blue-1.png){.absolute left=50 height="600"}

## Pelagic Case Study

![](imgs/03_marlin_paper_files/figure-pdf/fig-blue-results-1.png){.absolute top=100 left=200 width="750" height="600"}

## Complexity Matters! (but it's costly)

. . .

![](imgs/complexity.png)

## Linking to Team Tuna {.smaller}

-   Many models still stuck in the dietary dark ages

    -   Diet data could be used to inform taxis matrix?

-   Humans need to be part of the "diet Renaissance"!

-   `marlin` provides a mechanistic link between tagging studies and
    simulation work

    -   Can be linked to advection and environmental covariates as a
        bonus!

-   Future work adding in more species modules

    -   Seabirds?
    -   Marine mammals?
    -   Turtles?

## Of MICE and `marlin` {.smaller}

::: columns
::: {.column width="60%"}

- Spatial stock assessment and management increasingly needed

- BBNJ and 30x30 means that "tunas" are now firmly in the great MPA debate

- Lack of consistent / transparent assumptions can produce inconsistent outcomes

- [MICE(ish)](https://research.csiro.au/mice/) models like `marlin` can help illuminate drivers of social-ecological outcomes in fisheries

:::
::: {.column width="40%"}

![](imgs/bbnj.png)

![](imgs/medoff.png)

![](imgs/hampton.png)

:::

:::
## Questions?

::: columns
::: {.column width="65%"}
-   Be on the lookout for `marlin` paper sometime in 2023 (fingers
    crossed)

-   Check out the model at <br> [https://danovando.github.io/marlin/](https://danovando.github.io/marlin/)

-   Open source! Contributions welcome

-   email: dovando@iattc.org

-   github: danovando

-   twitter: @danovand0

- website: [danovando.com](https://www.weirdfishes.blog/)
:::

::: {.column width="35%"}
![](imgs/IMG_1679.JPG)
:::
:::